---
title: "metaG_CN"
author: "Colin G. Finlay"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---
# Setup

```{r setup, include=FALSE}
#use to set global options for chunks e.g., echo and warning options will be applied to all chunks:
knitr::opts_chunk$set(echo = TRUE)

# Clear environment:
rm(list=ls())

# Set working directory:
setwd("~/GitHub/WetlandMesocosm_GHG_Timberlake/analyses")

# Standard error (se) and confidence interval (ci):
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code dependencies:
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("reshape"); require("ggplot2"): require("tidyr"); require("dplyr"); require("dbstats"); require("reshape2"); require("ggpubr"); require("glue"); require("stringr")
```

# Load files

Import KO function files from JGI IMG/MER
```{r Import Files}
#Create vector with desired row order:
sample_order <- c("03NFC", "03PFC", "03Pre", "06NFC", "06PFC", "06Pre", "08NFC", "08PFC", "08Pre", "09NFC", "09PFC", "09Pre", "10NFC", "10PFC", "10Pre", "11NFC", "11PFC", "11Pre", "12NFC", "12PFC", "12Pre", "16NFC", "16PFC", "16Pre")
length(sample_order) #Should be 24

#Denitrification KEGG module:
denit <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Denit_CF.csv")
#Reorder rows:
denit <- denit[match(sample_order, denit$Sample), ]

#Methanogen Signature KEGG module:
meth <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Meth_CF.csv")
meth <- meth[match(sample_order, meth$Sample), ]

#All Central Carbohydrate Metabolism KEGG modules:
carb_metab <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Central_carb_metab_CF.csv")
carb_metab <- carb_metab[match(sample_order, carb_metab$Sample), ]

#Cytochrome C Oxidase KEGG module:
cytoCox <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/CytCOx_CF.csv")
cytoCox <- cytoCox[match(sample_order, cytoCox$Sample), ]

#Design file:
design <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/design.csv")
design2 <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/tmg_design.csv")
#Reorder rows of design:
design <- design[match(sample_order, design$Sample), ]

#Remove mock, field, and box samples from design2. 
rownames(design2) <- design2$sample
design2 <- design2[!(rownames(design2) %in% "MOCK"),]
design2 <- design2[!(rownames(design2) %in% "TL1"),]
design2 <- design2[!(rownames(design2) %in% "TL2"),]
design2 <- design2[!(rownames(design2) %in% "TL3"),]
design2 <- design2[!(rownames(design2) %in% "TL4"),]
design2 <- design2[!(rownames(design2) %in% "TL5"),]
design2 <- design2[!(rownames(design2) %in% "TL6"),]
design2 <- design2[!(design2$source %in% "b"),]
design2 <- design2[!(design2$source %in% "s"),]

#Greenhouse gas data:
#GHG <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/TMG_GHG-Flux_T1_Pub_2-2020.csv")
GHG <- read.csv("../data/TMG_GHG-Flux_Pub_3.csv")
```

# Data Wrangling

Convert gene counts to relative abundance:
```{r gene counts to relative abundance}
#Make sample IDs the rownames:
denitNumerical <- denit
rownames(denitNumerical) <- denitNumerical$Sample

methNumerical <- meth
rownames(methNumerical) <- methNumerical$Sample

carb_metabNumerical <- carb_metab
rownames(carb_metabNumerical) <- carb_metabNumerical$Sample

cytoCoxNumerical <- cytoCox
rownames(cytoCoxNumerical) <- cytoCoxNumerical$Sample

# Drop all non-numeric variables:
denitNumerical <- denitNumerical[,3:12]
methNumerical <- methNumerical[,3:46]
carb_metabNumerical <- carb_metabNumerical[,3:122]
cytoCoxNumerical <- cytoCoxNumerical[,3:7]

#Convert denitrification KO counts to relative abundance:
denitRA <- denitNumerical
for(i in 1:dim(denitNumerical)[1]){
  denitRA[i,] <- denitNumerical[i,]/sum(denitNumerical[i,])
}

#Convert methanogen KO counts to relative abundance:
methRA <- methNumerical
for(i in 1:dim(methNumerical)[1]){
  methRA[i,] <- methNumerical[i,]/sum(methNumerical[i,])
}

#Convert central carbohydrate metabolism KO counts to relative abundance:
carb_metabRA <- carb_metabNumerical
for(i in 1:dim(carb_metabNumerical)[1]){
  carb_metabRA[i,] <- carb_metabNumerical[i,]/sum(carb_metabNumerical[i,])
}

#Convert cytochrome C oxidase KO counts to relative abundance:
cytoCoxRA <- cytoCoxNumerical
for(i in 1:dim(cytoCoxNumerical)[1]){
  cytoCoxRA[i,] <- cytoCoxNumerical[i,]/sum(cytoCoxNumerical[i,])
}

#Subset start and finish samples from all gene sets (for respective start and finish GHG analysis):
denitRA_start <- denitRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
denitRA_finish <- denitRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

methRA_start <- methRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
methRA_finish <- methRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

carb_metabRA_start <- carb_metabRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
carb_metabRA_finish <- carb_metabRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

cytoCoxRA_start <- cytoCoxRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
cytoCoxRA_finish <- cytoCoxRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

#Check dimensions of each relative abundance (RA) matrix:
dim(denitRA) #24 10
dim(methRA) #24 44
dim(carb_metabRA) #24 120
dim(cytoCoxRA) #24 5

dim(denitRA_start) #8 10
dim(methRA_start) #8 44
dim(carb_metabRA_start) #8 120
dim(cytoCoxRA_start) #8 5

dim(denitRA_finish) #16 10
dim(methRA_finish) #16 44
dim(carb_metabRA_finish) #16 120
dim(cytoCoxRA_finish) #16 5

#Make metaG sample ID's the rownames on 'design'. Then drop design$Sample:
rownames(design) <- design$Sample
design <- design[,2:4]

# Make data frames from gene set relative abundance (RA) files and design
denitRA_design <- cbind(design, denitRA)
methRA_design <- cbind(design, methRA)
carb_metabRA_design <- cbind(design, carb_metabRA)
cytoCoxRA_design <- cbind(design, cytoCoxRA)
```

## Bray Curtis dissimilarity

Calculate Bray Curtis dissimilarity:
```{r relative abundance to Bray Curtis dissimilarity}
#Bray Curtis calculation for denitrification relative abundance data:
denitBC <- vegdist(denitRA, method = "bray")
denitBC_start <- vegdist(denitRA_start, method = "bray")
denitBC_finish <- vegdist(denitRA_finish, method = "bray")
#print(denitBC)

#Bray Curtis calculation for methanogen relative abundance data:
methBC <- vegdist(methRA, method = "bray")
methBC_start <- vegdist(methRA_start, method = "bray")
methBC_finish <- vegdist(methRA_finish, method = "bray")
#print(methBC)

#Bray Curtis calculation for central carbohydrate metabolism relative abundance data:
carb_metabBC <- vegdist(carb_metabRA, method = "bray")
carb_metabBC_start <- vegdist(carb_metabRA_start, method = "bray")
carb_metabBC_finish <- vegdist(carb_metabRA_finish, method = "bray")
#print(carb_metabBC)

#Bray Curtis calculation for cytochrome C oxidase relative abundance data:
cytoCoxBC <- vegdist(cytoCoxRA, method = "bray")
cytoCoxBC_start <- vegdist(cytoCoxRA_start, method = "bray")
cytoCoxBC_finish <- vegdist(cytoCoxRA_finish, method = "bray")
#print(cytoCoxBC)
```

## GHG wrangling

Wrangling greenhouse gas (GHG) data:
```{r wrangling GHG data}
# Fix 'c2o' typo:
colnames(GHG)[18] <- "co2_ppm"

#Create a new GHG data frame with less columns:
#GHG_less_columns <- cbind(GHG[,1:4], GHG[24:32])
GHG_less_columns <- cbind(GHG[,1:4], GHG[17:19])

# Select only T1 rows, and July 24 (night sampling):
#GHG_less_columns <- GHG_less_columns %>% drop_na()
GHG_less_columns <- GHG_less_columns %>%
  dplyr::filter(timepoint == "T1" & date !='7/24/2016')

#Change column name "chamber" to "box" to match with design2:
colnames(GHG_less_columns)[2] <- "box"

#Convert GHG_less_columns$plant to lowercase to match with design2$plant:
GHG_less_columns$plant <- tolower(GHG_less_columns$plant)

#Subset GHG_less_columns for 'start' time point (6/13/2016) and 'finish' time point (8/11/2016):
GHG_start <- dplyr::filter(GHG_less_columns, date == "6/13/2016")
GHG_finish <- dplyr::filter(GHG_less_columns, date == "8/11/2016")

#Add design2 into GHG 'start' and 'finish':
GHG_start <- left_join(GHG_start, design2, by = c("box", "plant"))
GHG_finish <- left_join(GHG_finish, design2, by = c("box", "plant"))

#Subset GHG_finish and GHG_start to only have metaG sample numbers:
GHG_start_metaG <- rbind(GHG_start[5:6,], GHG_start[11:12,], GHG_start[15:24,], GHG_start[31:32,], make.row.names = FALSE)
GHG_finish_metaG <- rbind(GHG_finish[5:6,], GHG_finish[11:12,], GHG_finish[15:24,], GHG_finish[31:32,], make.row.names = FALSE)

#To align "pre" metaG samples with a single, initial GHG value (from the start GHG measurements on 6/28/206), average together GHG samples in the same box number, then transpose and set row name to corresponding metaG Sample ID:
GHG_start_03 <- data.frame(colMeans(GHG_start_metaG[1:2, 5:7]))
GHG_start_03 <- as.data.frame(t(GHG_start_03))
row.names(GHG_start_03) <- "03Pre"

GHG_start_06 <- data.frame(colMeans(GHG_start_metaG[3:4, 5:7]))
GHG_start_06 <- as.data.frame(t(GHG_start_06))
row.names(GHG_start_06) <- "06Pre"

GHG_start_08 <- data.frame(colMeans(GHG_start_metaG[5:6, 5:7]))
GHG_start_08 <- as.data.frame(t(GHG_start_08))
row.names(GHG_start_08) <- "08Pre"

GHG_start_09 <- data.frame(colMeans(GHG_start_metaG[7:8, 5:7]))
GHG_start_09 <- as.data.frame(t(GHG_start_09))
row.names(GHG_start_09) <- "09Pre"

GHG_start_10 <- data.frame(colMeans(GHG_start_metaG[9:10, 5:7]))
GHG_start_10 <- as.data.frame(t(GHG_start_10))
row.names(GHG_start_10) <- "10Pre"

GHG_start_11 <- data.frame(colMeans(GHG_start_metaG[11:12, 5:7]))
GHG_start_11 <- as.data.frame(t(GHG_start_11))
row.names(GHG_start_11) <- "11Pre"

GHG_start_12 <- data.frame(colMeans(GHG_start_metaG[13:14, 5:7]))
GHG_start_12 <- as.data.frame(t(GHG_start_12))
row.names(GHG_start_12) <- "12Pre"

GHG_start_16 <- data.frame(colMeans(GHG_start_metaG[15:16, 5:7]))
GHG_start_16 <- as.data.frame(t(GHG_start_16))
row.names(GHG_start_16) <- "16Pre"

#Then rbind() all the dfs together
GHG_start_metaG_avg <- rbind(GHG_start_03, GHG_start_06, GHG_start_08, GHG_start_09, GHG_start_10, GHG_start_11, GHG_start_12, GHG_start_16)

#Set the row names of GHG data to the metaG sample names, so that stats functions can compare the GHG and metaG data sets:
row.names(GHG_finish_metaG) <- c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC","09PFC", "10NFC","10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC")

#Combine GHG_start_metaG_avg and GHG_finish_metaG[,5:7] to have all 24 samples:
GHG_metaG <- rbind(GHG_start_metaG_avg, GHG_finish_metaG[,5:7])
GHG_metaG <- GHG_metaG[match(sample_order, rownames(GHG_metaG)), ]
```

# Statistics

## dbplsr

Distance-based partial least squares regression:
```{r distance-based partial least squares regression}
start.denit.dbplsr <- dbplsr(GHG_start_metaG_avg$n2o_ppm ~ as.matrix(denitBC_start), ncomp = 7, method = "GCV")
summary(start.denit.dbplsr)
#plot(start.denit.dbplsr)

finish.denit.dbplsr <- dbplsr(GHG_finish_metaG$n2o_ppm ~ as.matrix(denitBC_finish), ncomp = 15, method = "GCV")
summary(finish.denit.dbplsr)
#plot(finish.denit.dbplsr)

start.meth.dbplsr <- dbplsr(GHG_start_metaG_avg$ch4_ppm ~ as.matrix(methBC_start), ncomp = 7, method = "GCV")
summary(start.meth.dbplsr)
#plot(start.meth.dbplsr)

finish.meth.dbplsr <- dbplsr(GHG_finish_metaG$ch4_ppm ~ as.matrix(methBC_finish), ncomp = 15, method = "GCV")
summary(finish.meth.dbplsr)
#plot(finish.meth.dbplsr)

start.carb_metab.dbplsr <- dbplsr(GHG_start_metaG_avg$co2_ppm ~ as.matrix(carb_metabBC_start), ncomp = 7, method = "GCV")
summary(start.carb_metab.dbplsr)
#plot(start.carb_metab.dbplsr)

finish.carb_metab.dbplsr <- dbplsr(GHG_finish_metaG$co2_ppm ~ as.matrix(carb_metabBC_finish), ncomp = 15, method = "GCV")
summary(finish.carb_metab.dbplsr)
#plot(finish.carb_metab.dbplsr)

start.cytoCox.dbplsr <- dbplsr(GHG_start_metaG_avg$co2_ppm ~ as.matrix(cytoCoxBC_start), ncomp = 7, method = "GCV")
summary(start.cytoCox.dbplsr)
#plot(start.cytoCox.dbplsr)

finish.cytoCox.dbplsr <- dbplsr(GHG_finish_metaG$co2_ppm ~ as.matrix(cytoCoxBC_finish), ncomp = 15, method = "GCV")
summary(finish.cytoCox.dbplsr)
#plot(finish.cytoCox.dbplsr)

# Because of envfit() results showing significant relationship between denitrification composition and [CO2], running dbplsrs for denit and CO2:
start.denit.dbplsr.co2 <- dbplsr(GHG_start_metaG_avg$co2_ppm ~ as.matrix(denitBC_start), ncomp = 7, method = "GCV")
summary(start.denit.dbplsr.co2)
#plot(start.denit.dbplsr.co2)

finish.denit.dbplsr.co2 <- dbplsr(GHG_finish_metaG$co2_ppm ~ as.matrix(denitBC_finish), ncomp = 15, method = "GCV")
summary(finish.denit.dbplsr.co2)
#plot(finish.denit.dbplsr.co2)
```

## PERMANOVA

```{r set factors}
# Treatment (Treat):
denitRA_design$Treat <- as.factor(denitRA_design$Treat)
methRA_design$Treat <- as.factor(methRA_design$Treat)
carb_metabRA_design$Treat <- as.factor(carb_metabRA_design$Treat)
cytoCoxRA_design$Treat <- as.factor(cytoCoxRA_design$Treat)

# History (Hist):
denitRA_design$Hist <- as.factor(denitRA_design$Hist)
methRA_design$Hist <- as.factor(methRA_design$Hist)
carb_metabRA_design$Hist <- as.factor(carb_metabRA_design$Hist)
cytoCoxRA_design$Hist <- as.factor(cytoCoxRA_design$Hist)

# Plant presence/absence (Plant):
denitRA_design$Plant <- as.factor(denitRA_design$Plant)
methRA_design$Plant <- as.factor(methRA_design$Plant)
carb_metabRA_design$Plant <- as.factor(carb_metabRA_design$Plant)
cytoCoxRA_design$Plant <- as.factor(cytoCoxRA_design$Plant)
```

```{r PERMANOVA}
#PERMANOVA without History:
adonis_denit = adonis2(denitRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = denitRA_design, perm=10000, set.seed=42)
adonis_denit

adonis_meth = adonis2(methRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = methRA_design, perm=10000, set.seed=42)
adonis_meth

adonis_carb_metab = adonis2(carb_metabRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = carb_metabRA_design, perm=10000, set.seed=42)
adonis_carb_metab

adonis_cytoCox = adonis2(cytoCoxRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = cytoCoxRA_design, perm=10000, set.seed=42)
adonis_cytoCox

#PERMANOVA with History:
adonis_denit_H = adonis2(denitRA_design[,-c(1:3)]~Plant*Hist + Plant*Treat, method = "bray", data = denitRA_design, perm=10000, set.seed=42)
adonis_denit_H

adonis_meth_H = adonis2(methRA_design[,-c(1:3)]~Plant*Hist + Plant*Treat, method = "bray", data = methRA_design, perm=10000, set.seed=42)
adonis_meth_H

adonis_carb_metab_H = adonis2(carb_metabRA_design[,-c(1:3)]~Plant*Hist + Plant*Treat, method = "bray", data = carb_metabRA_design, perm=10000, set.seed=42)
adonis_carb_metab_H

adonis_cytoCox_H = adonis2(cytoCoxRA_design[,-c(1:3)]~Plant*Hist + Plant*Treat, method = "bray", data = cytoCoxRA_design, perm=10000, set.seed=42)
adonis_cytoCox_H
```
## PCoA

Principal Coordinates Analysis (PCoA) denitrification:
```{r PCoA: denitrification, start&finish}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_denit <- cmdscale(denitBC, k=2, eig=TRUE, add=TRUE) # changing k to 2 because we only want 2 dimensions anyway
  # Now trying add = TRUE because of https://www.youtube.com/watch?v=G5Qckqq5Erw

explainvar1b_denit <- round(pcoa_denit$eig[1] / sum(pcoa_denit$eig), 3) * 100
explainvar2b_denit <- round(pcoa_denit$eig[2] / sum(pcoa_denit$eig), 3) * 100
sum.eigb <- sum(explainvar1b_denit, explainvar2b_denit)

explainvar1b_denit # 27.1
explainvar2b_denit # 16.3

pcoa_denit2.groups <- paste(denitRA_design$Treat, denitRA_design$Plant, denitRA_design$Hist, sep = "_")
pcoa_denit2.points <- data.frame(pcoa_denit$points, group = pcoa_denit2.groups)

# Calculate Centroids (mean and SE)
pcoa_denit2.L.centroids <- reshape2::melt(pcoa_denit2.points, id="group", measure.vars = c("X1", "X2"))
pcoa_denit2.centroids <- acast(pcoa_denit2.L.centroids, variable ~ group, mean)
pcoa_denit2.centroids.se <- acast(pcoa_denit2.L.centroids, variable ~ group, se)
pcoa_denit2.centroids.sd <- acast(pcoa_denit2.L.centroids, variable ~ group, sd)

# Combine
pcoa_denit2.cent.dataframe <- cbind(t(pcoa_denit2.centroids), t(pcoa_denit2.centroids.se))
colnames(pcoa_denit2.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

# Adding columns with the factors. H: hydrologic history
trt_H <- c("Baseline", "Baseline", "Dry","Dry", "Dry","Dry","Wet","Wet", "Wet","Wet") 
plant_H <- c("Baseline", "Baseline", "No Plant","No Plant","Plant", "Plant", "No Plant","No Plant","Plant", "Plant")
history <- c("Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet")
history_plant_combined <- c(paste(history, plant_H, sep = "_"))

pcoa_denit2.cent.dataframe.trts <- as.data.frame(pcoa_denit2.cent.dataframe) 
pcoa_denit2.cent.dataframe.trts$trt <- as.factor(trt_H)
pcoa_denit2.cent.dataframe.trts$plant <- as.factor(plant_H)
pcoa_denit2.cent.dataframe.trts$history <- as.factor(history)
pcoa_denit2.cent.dataframe.trts$history_plant<- as.factor(history_plant_combined)
pcoa_denit2.cent.dataframe.trts$centroid_or_sample <- "centroid"

# Add in points to data frame:
pcoa_denit2.points_ForPlot <- pcoa_denit2.points
colnames(pcoa_denit2.points_ForPlot) <- c("V1", "V2", "group")
# Add labels to match ...cent.dataframe.trts
pcoa_denit2.points_ForPlot$trt <- denitRA_design$Treat %>% str_replace_all(c("w" = "Wet", "_pre" = "Baseline", "d" = "Dry"))
pcoa_denit2.points_ForPlot$plant <- denitRA_design$Plant %>% str_replace_all(c("n" = "No Plant", "y" = "Plant", "_pre" = "Baseline"))
pcoa_denit2.points_ForPlot$history <- denitRA_design$Hist %>% str_replace_all(c("d" = "Dry", "w" = "Wet"))
pcoa_denit2.points_ForPlot$history_plant <- c(paste(pcoa_denit2.points_ForPlot$history, pcoa_denit2.points_ForPlot$plant, sep = "_"))
#remove "group" column
pcoa_denit2.points_ForPlot <- pcoa_denit2.points_ForPlot[, -3]
#add V1e and V2e as zeros:
pcoa_denit2.points_ForPlot$V1e <- numeric(24)
pcoa_denit2.points_ForPlot$V2e <- numeric(24)
# Add column to distinguish sample points from centroids, and later use to set alpha in plot
pcoa_denit2.points_ForPlot$centroid_or_sample <- "sample"

#Merge:
denit_centroids_points_merged <- rbind(pcoa_denit2.cent.dataframe.trts, pcoa_denit2.points_ForPlot)
```
```{r envfit with GHGs ONLY}
#Run envfit to see which [GHG]s (start or finish) have good fit with ordination:
fit_denit <- envfit(pcoa_denit$points, GHG_metaG, permutations = 10000)
fit_denit

A_denit <-as.list(fit_denit$vectors)
vec_denit <- as.data.frame(fit_denit$vectors$arrows*sqrt(fit_denit$vectors$r)*0.15)
p_denit <- as.data.frame(A_denit$pvals)
vec_denit <- cbind(vec_denit, p_denit)
vec_denit <- subset(vec_denit, A_denit$pvals<=0.1)
```


```{r plot denit PCoA}
#Plot using ggplot2
df2a <- as.data.frame(denit_centroids_points_merged)

plot2a <- ggplot(df2a, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.003), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.003), colour="black") +   
  geom_point(aes(shape = history_plant, colour=trt, alpha = centroid_or_sample, size = centroid_or_sample), stroke=2) +  
  theme_bw() +
  scale_size_manual(values = c(6, 3), name = "Centroid or Sample")+
  scale_alpha_manual(values = c(1, 0.3), name = "Centroid or Sample")+
  scale_color_manual(values=c("black", "#997700", "#004488"), name="Treatment") +
  scale_shape_manual(values = c(7, 0, 15, 13, 1, 19), name="History_Plant") +
  xlab(glue("PCoA 1 ({explainvar1b_denit}%)")) + ylab(glue("PCoA 2 ({explainvar2b_denit}%)")) + 
  geom_segment(data=vec_denit, aes(x=0,xend=Dim1,y=0,yend=Dim2), linewidth=1, arrow = arrow(length = unit(0.2, "cm")),colour="black", alpha = 0.4, inherit.aes=F)+ 
 geom_text(data=vec_denit, aes(x=Dim1+0.025, y=Dim2), label= expression(paste("CO"[2]," ppm")), size=3, inherit.aes=F)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14), 
          axis.text.x = element_text(size=14),  axis.text.y = element_text(size=14),
          panel.border = element_rect(colour = "black", linewidth=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=14)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(shape = guide_legend(override.aes = list(size = 4), order = 1), 
           colour = guide_legend(override.aes = list(pch=16, size = 4), order = 2)) +
    ggtitle("Denitrification")

plot2a
```

PCoA Methanogenesis, all factors, HISTORY AS SHAPE:
```{r PCoA: Methanogenesis; all factors}
pcoa_meth <- cmdscale(methBC, k=2, eig=TRUE, add=TRUE)

explainvar1b_meth <- round(pcoa_meth$eig[1] / sum(pcoa_meth$eig), 3) * 100
explainvar2b_meth <- round(pcoa_meth$eig[2] / sum(pcoa_meth$eig), 3) * 100
sum.eigb <- sum(explainvar1b_meth, explainvar2b_meth)

explainvar1b_meth # 29.7
explainvar2b_meth # 20.8

pcoa_meth2.groups <- paste(methRA_design$Treat, methRA_design$Plant, methRA_design$Hist, sep = "_")
pcoa_meth2.points <- data.frame(pcoa_meth$points, group = pcoa_meth2.groups)

# Calculate Centroids (mean and SE)
pcoa_meth2.L.centroids <- reshape2::melt(pcoa_meth2.points, id="group", measure.vars = c("X1", "X2"))
pcoa_meth2.centroids <- acast(pcoa_meth2.L.centroids, variable ~ group, mean)
pcoa_meth2.centroids.se <- acast(pcoa_meth2.L.centroids, variable ~ group, se)
pcoa_meth2.centroids.sd <- acast(pcoa_meth2.L.centroids, variable ~ group, sd)

# Combine
pcoa_meth2.cent.dataframe <- cbind(t(pcoa_meth2.centroids), t(pcoa_meth2.centroids.se))
colnames(pcoa_meth2.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

# No need to run again if these lists have already been declared:
#trt_H <- c("Baseline", "Baseline", "Dry","Dry", "Dry","Dry","Wet","Wet", "Wet","Wet") 
#plant_H <- c("Baseline", "Baseline", "No Plant","No Plant","Plant", "Plant", "No Plant","No Plant","Plant", "Plant")
#history <- c("Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet")
#trt_plant_combined <- c(paste(trt, plant, sep = "_"))

pcoa_meth2.cent.dataframe.trts <- as.data.frame(pcoa_meth2.cent.dataframe) 
pcoa_meth2.cent.dataframe.trts$trt <- as.factor(trt_H)
pcoa_meth2.cent.dataframe.trts$plant <- as.factor(plant_H)
pcoa_meth2.cent.dataframe.trts$history <- as.factor(history)
pcoa_meth2.cent.dataframe.trts$history_plant<- as.factor(history_plant_combined)
pcoa_meth2.cent.dataframe.trts$centroid_or_sample <- "centroid"

# Add in points to data frame:
pcoa_meth2.points_ForPlot <- pcoa_meth2.points
colnames(pcoa_meth2.points_ForPlot) <- c("V1", "V2", "group")
# Add labels to match ...cent.dataframe.trts
pcoa_meth2.points_ForPlot$trt <- methRA_design$Treat %>% str_replace_all(c("w" = "Wet", "_pre" = "Baseline", "d" = "Dry"))
pcoa_meth2.points_ForPlot$plant <- methRA_design$Plant %>% str_replace_all(c("n" = "No Plant", "y" = "Plant", "_pre" = "Baseline"))
pcoa_meth2.points_ForPlot$history <- methRA_design$Hist %>% str_replace_all(c("d" = "Dry", "w" = "Wet"))
pcoa_meth2.points_ForPlot$history_plant <- c(paste(pcoa_meth2.points_ForPlot$history, pcoa_meth2.points_ForPlot$plant, sep = "_"))
#remove "group" column
pcoa_meth2.points_ForPlot <- pcoa_meth2.points_ForPlot[, -3]
#add V1e and V2e as zeros:
pcoa_meth2.points_ForPlot$V1e <- numeric(24)
pcoa_meth2.points_ForPlot$V2e <- numeric(24)
# Add column to distinguish sample points from centroids, and later use to set alpha in plot
pcoa_meth2.points_ForPlot$centroid_or_sample <- "sample"

#Merge:
meth_centroids_points_merged <- rbind(pcoa_meth2.cent.dataframe.trts, pcoa_meth2.points_ForPlot)
```
```{r envfit with GHGs ONLY}
#Run envfit to see which [GHG]s (start or finish) have good fit with ordination:
fit_meth <- envfit(pcoa_meth$points, GHG_metaG, permutations = 10000)
fit_meth

A_meth <-as.list(fit_meth$vectors)
vec_meth <- as.data.frame(fit_meth$vectors$arrows*sqrt(fit_meth$vectors$r)*0.15)
p_meth <- as.data.frame(A_meth$pvals)
vec_meth <- cbind(vec_meth, p_meth)
vec_meth <- subset(vec_meth, A_meth$pvals<=0.1)
```

```{r plot meth PCoA}
#Plot using ggplot2
df2b <- as.data.frame(meth_centroids_points_merged)

plot2b <- ggplot(df2b, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.003), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.003), colour="black") +   
  geom_point(aes(shape = history_plant, colour=trt, alpha = centroid_or_sample, size = centroid_or_sample), stroke=2) +  
  theme_bw() +
  scale_size_manual(values = c(6, 3), name = "Centroid or Sample")+
  scale_alpha_manual(values = c(1, 0.3), name = "Centroid or Sample")+
  scale_color_manual(values=c("black", "#997700", "#004488"), name="Treatment") +
  scale_shape_manual(values = c(7, 0, 15, 13, 1, 19), name="History_Plant") +
  xlab(glue("PCoA 1 ({explainvar1b_meth}%)")) + ylab(glue("PCoA 2 ({explainvar2b_meth}%)")) + 
  geom_segment(data=vec_meth, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, arrow = arrow(length = unit(0.2, "cm")),colour="black", alpha=0.4, inherit.aes=F)+ 
 geom_text(data=vec_meth, aes(x=Dim1, y=Dim2+0.004), label= expression(paste("CH"[4]," ppm")), size=3, inherit.aes=F)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14), 
          axis.text.x = element_text(size=14),  axis.text.y = element_text(size=14),
          panel.border = element_rect(colour = "black", linewidth=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=14)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(shape = guide_legend(override.aes = list(size = 4), order = 1), 
           colour = guide_legend(override.aes = list(pch=16, size = 4), order = 2)) +
    ggtitle("Methanogenesis")

plot2b
```



PCoA central carbohydrate metabolism, all factors, HISTORY AS SHAPE:
```{r PCoA: central carbohydrate metabolism; all factors}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_carb_metab <- cmdscale(carb_metabBC, k=2, eig=TRUE, add=TRUE)

explainvar1b_carb_metab <- format(round(pcoa_carb_metab$eig[1] / sum(pcoa_carb_metab$eig), 3) * 100, nsmall = 1) # Using format() to get .0
explainvar2b_carb_metab <- round(pcoa_carb_metab$eig[2] / sum(pcoa_carb_metab$eig), 3) * 100
#sum.eigc <- sum(explainvar1b_carb_metab, explainvar2b_carb_metab) #Doesn't work when using format()

explainvar1b_carb_metab # 31.0
explainvar2b_carb_metab # 14.9

pcoa_carb_metab2.groups <- paste(carb_metabRA_design$Treat, carb_metabRA_design$Plant, carb_metabRA_design$Hist, sep = "_")
pcoa_carb_metab2.points <- data.frame(pcoa_carb_metab$points, group = pcoa_carb_metab2.groups)

# Calculate Centroids (mean and SE)
pcoa_carb_metab2.L.centroids <- reshape2::melt(pcoa_carb_metab2.points, id="group", measure.vars = c("X1", "X2"))
pcoa_carb_metab2.centroids <- acast(pcoa_carb_metab2.L.centroids, variable ~ group, mean)
pcoa_carb_metab2.centroids.se <- acast(pcoa_carb_metab2.L.centroids, variable ~ group, se)
pcoa_carb_metab2.centroids.sd <- acast(pcoa_carb_metab2.L.centroids, variable ~ group, sd)

# Combine
pcoa_carb_metab2.cent.dataframe <- cbind(t(pcoa_carb_metab2.centroids), t(pcoa_carb_metab2.centroids.se))
colnames(pcoa_carb_metab2.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

# No need to run again if these lists have already been declared:
#trt_H <- c("Baseline", "Baseline", "Dry","Dry", "Dry","Dry","Wet","Wet", "Wet","Wet") 
#plant_H <- c("Baseline", "Baseline", "No Plant","No Plant","Plant", "Plant", "No Plant","No Plant","Plant", "Plant")
#history <- c("Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet")
#trt_plant_combined <- c(paste(trt, plant, sep = "_"))

pcoa_carb_metab2.cent.dataframe.trts <- as.data.frame(pcoa_carb_metab2.cent.dataframe) 
pcoa_carb_metab2.cent.dataframe.trts$trt <- as.factor(trt_H)
pcoa_carb_metab2.cent.dataframe.trts$plant <- as.factor(plant_H)
pcoa_carb_metab2.cent.dataframe.trts$history <- as.factor(history)
pcoa_carb_metab2.cent.dataframe.trts$history_plant<- as.factor(history_plant_combined)
pcoa_carb_metab2.cent.dataframe.trts$centroid_or_sample <- "centroid"

# Add in points to data frame:
pcoa_carb_metab2.points_ForPlot <- pcoa_carb_metab2.points
colnames(pcoa_carb_metab2.points_ForPlot) <- c("V1", "V2", "group")
# Add labels to match ...cent.dataframe.trts
pcoa_carb_metab2.points_ForPlot$trt <- carb_metabRA_design$Treat %>% str_replace_all(c("w" = "Wet", "_pre" = "Baseline", "d" = "Dry"))
pcoa_carb_metab2.points_ForPlot$plant <- carb_metabRA_design$Plant %>% str_replace_all(c("n" = "No Plant", "y" = "Plant", "_pre" = "Baseline"))
pcoa_carb_metab2.points_ForPlot$history <- carb_metabRA_design$Hist %>% str_replace_all(c("d" = "Dry", "w" = "Wet"))
pcoa_carb_metab2.points_ForPlot$history_plant <- c(paste(pcoa_carb_metab2.points_ForPlot$history, pcoa_carb_metab2.points_ForPlot$plant, sep = "_"))
#remove "group" column
pcoa_carb_metab2.points_ForPlot <- pcoa_carb_metab2.points_ForPlot[, -3]
#add V1e and V2e as zeros:
pcoa_carb_metab2.points_ForPlot$V1e <- numeric(24)
pcoa_carb_metab2.points_ForPlot$V2e <- numeric(24)
# Add column to distinguish sample points from centroids, and later use to set alpha in plot
pcoa_carb_metab2.points_ForPlot$centroid_or_sample <- "sample"

#Merge:
carb_metab_centroids_points_merged <- rbind(pcoa_carb_metab2.cent.dataframe.trts, pcoa_carb_metab2.points_ForPlot)
```
```{r envfit with GHGs ONLY}
#Run envfit to see which [GHG]s (start or finish) have good fit with ordination:
fit_carb_metab <- envfit(pcoa_carb_metab$points, GHG_metaG, permutations = 10000)
fit_carb_metab

A_carb_metab <-as.list(fit_carb_metab$vectors)
vec_carb_metab <- as.data.frame(fit_carb_metab$vectors$arrows*sqrt(fit_carb_metab$vectors$r)*0.15)
p_carb_metab <- as.data.frame(A_carb_metab$pvals)
vec_carb_metab <- cbind(vec_carb_metab, p_carb_metab)
vec_carb_metab <- subset(vec_carb_metab, A_carb_metab$pvals<=0.1)
```
```{r carb_metab SSF envfit and PCoA}
#Plot using ggplot2
df2c <- as.data.frame(carb_metab_centroids_points_merged)

plot2c <- ggplot(df2c, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.003), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.003), colour="black") +   
  geom_point(aes(shape = history_plant, colour=trt, alpha = centroid_or_sample, size = centroid_or_sample), stroke=2) +  
  theme_bw() +
  scale_size_manual(values = c(6, 3), name = "Centroid or Sample")+
  scale_alpha_manual(values = c(1, 0.3), name = "Centroid or Sample")+
  scale_color_manual(values=c("black", "#997700", "#004488"), name="Treatment") +
  scale_shape_manual(values = c(7, 0, 15, 13, 1, 19), name="History_Plant") +
  xlab(glue("PCoA 1 ({explainvar1b_carb_metab}%)")) + ylab(glue("PCoA 2 ({explainvar2b_carb_metab}%)")) + 
  geom_segment(data=vec_carb_metab, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, arrow = arrow(length = unit(0.2, "cm")),colour="black", alpha = 0.4, inherit.aes=F)+ 
 geom_text(data=vec_carb_metab, aes(x=Dim1+0.015, y=Dim2-0.005), label= expression(paste("CO"[2]," ppm")), size=3, inherit.aes=F)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14), 
          axis.text.x = element_text(size=14),  axis.text.y = element_text(size=14),
          panel.border = element_rect(colour = "black", linewidth=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=14)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(shape = guide_legend(override.aes = list(size = 4), order = 1), 
           colour = guide_legend(override.aes = list(pch=16, size = 4), order = 2)) +
    ggtitle("Central Carbohydrate Metabolism")

plot2c
```


PCoA cytochrome C oxidase, all factors, HISTORY AS SHAPE:
```{r PCoA: cytochrome C oxidase; all factors}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_cytoCox <- cmdscale(cytoCoxBC, k=2, eig=TRUE, add=TRUE)

explainvar1b_cytoCox <- round(pcoa_cytoCox$eig[1] / sum(pcoa_cytoCox$eig), 3) * 100
explainvar2b_cytoCox <- round(pcoa_cytoCox$eig[2] / sum(pcoa_cytoCox$eig), 3) * 100
sum.eigc <- sum(explainvar1b_cytoCox, explainvar2b_cytoCox)

explainvar1b_cytoCox # 29.2
explainvar2b_cytoCox # 22.2

pcoa_cytoCox2.groups <- paste(cytoCoxRA_design$Treat, cytoCoxRA_design$Plant, cytoCoxRA_design$Hist, sep = "_")
pcoa_cytoCox2.points <- data.frame(pcoa_cytoCox$points, group = pcoa_cytoCox2.groups)

# Calculate Centroids (mean and SE)
pcoa_cytoCox2.L.centroids <- reshape2::melt(pcoa_cytoCox2.points, id="group", measure.vars = c("X1", "X2"))
pcoa_cytoCox2.centroids <- acast(pcoa_cytoCox2.L.centroids, variable ~ group, mean)
pcoa_cytoCox2.centroids.se <- acast(pcoa_cytoCox2.L.centroids, variable ~ group, se)
pcoa_cytoCox2.centroids.sd <- acast(pcoa_cytoCox2.L.centroids, variable ~ group, sd)

# Combine
pcoa_cytoCox2.cent.dataframe <- cbind(t(pcoa_cytoCox2.centroids), t(pcoa_cytoCox2.centroids.se))
colnames(pcoa_cytoCox2.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

# No need to run again if these lists have already been declared:
#trt_H <- c("Baseline", "Baseline", "Dry","Dry", "Dry","Dry","Wet","Wet", "Wet","Wet") 
#plant_H <- c("Baseline", "Baseline", "No Plant","No Plant","Plant", "Plant", "No Plant","No Plant","Plant", "Plant")
#history <- c("Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet", "Dry", "Wet")
#trt_plant_combined <- c(paste(trt, plant, sep = "_"))

pcoa_cytoCox2.cent.dataframe.trts <- as.data.frame(pcoa_cytoCox2.cent.dataframe) 
pcoa_cytoCox2.cent.dataframe.trts$trt <- as.factor(trt_H)
pcoa_cytoCox2.cent.dataframe.trts$plant <- as.factor(plant_H)
pcoa_cytoCox2.cent.dataframe.trts$history <- as.factor(history)
pcoa_cytoCox2.cent.dataframe.trts$history_plant<- as.factor(history_plant_combined)
pcoa_cytoCox2.cent.dataframe.trts$centroid_or_sample <- "centroid"

# Add in points to data frame:
pcoa_cytoCox2.points_ForPlot <- pcoa_cytoCox2.points
colnames(pcoa_cytoCox2.points_ForPlot) <- c("V1", "V2", "group")
# Add labels to match ...cent.dataframe.trts
pcoa_cytoCox2.points_ForPlot$trt <- cytoCoxRA_design$Treat %>% str_replace_all(c("w" = "Wet", "_pre" = "Baseline", "d" = "Dry"))
pcoa_cytoCox2.points_ForPlot$plant <- cytoCoxRA_design$Plant %>% str_replace_all(c("n" = "No Plant", "y" = "Plant", "_pre" = "Baseline"))
pcoa_cytoCox2.points_ForPlot$history <- cytoCoxRA_design$Hist %>% str_replace_all(c("d" = "Dry", "w" = "Wet"))
pcoa_cytoCox2.points_ForPlot$history_plant <- c(paste(pcoa_cytoCox2.points_ForPlot$history, pcoa_cytoCox2.points_ForPlot$plant, sep = "_"))
#remove "group" column
pcoa_cytoCox2.points_ForPlot <- pcoa_cytoCox2.points_ForPlot[, -3]
#add V1e and V2e as zeros:
pcoa_cytoCox2.points_ForPlot$V1e <- numeric(24)
pcoa_cytoCox2.points_ForPlot$V2e <- numeric(24)
# Add column to distinguish sample points from centroids, and later use to set alpha in plot
pcoa_cytoCox2.points_ForPlot$centroid_or_sample <- "sample"

#Merge:
cytoCox_centroids_points_merged <- rbind(pcoa_cytoCox2.cent.dataframe.trts, pcoa_cytoCox2.points_ForPlot)
```
```{r envfit with GHGs ONLY}
#Run envfit to see which [GHG]s (start or finish) have good fit with ordination:
fit_cytoCox <- envfit(pcoa_cytoCox$points, GHG_metaG, permutations = 10000)
fit_cytoCox

A_cytoCox <-as.list(fit_cytoCox$vectors)
vec_cytoCox <- as.data.frame(fit_cytoCox$vectors$arrows*sqrt(fit_cytoCox$vectors$r)*0.15)
p_cytoCox <- as.data.frame(A_cytoCox$pvals)
vec_cytoCox <- cbind(vec_cytoCox, p_cytoCox)
vec_cytoCox <- subset(vec_cytoCox, A_cytoCox$pvals<=0.1)
```
```{r cytoCox SSF envfit and PCoA}
#Plot using ggplot2
df2d <- as.data.frame(cytoCox_centroids_points_merged)

plot2d <- ggplot(df2d, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.003), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.003), colour="black") +   
  geom_point(aes(shape = history_plant, colour=trt, alpha = centroid_or_sample, size = centroid_or_sample), stroke=2) +  
  theme_bw() +
  scale_size_manual(values = c(6, 3), name = "Centroid or Sample")+
  scale_alpha_manual(values = c(1, 0.3), name = "Centroid or Sample")+
  scale_color_manual(values=c("black", "#997700", "#004488"), name="Treatment") +
  scale_shape_manual(values = c(7, 0, 15, 13, 1, 19), name="History_Plant") +
  xlab(glue("PCoA 1 ({explainvar1b_cytoCox}%)")) + ylab(glue("PCoA 2 ({explainvar2b_cytoCox}%)")) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14), 
          axis.text.x = element_text(size=14),  axis.text.y = element_text(size=14),
          panel.border = element_rect(colour = "black", linewidth=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=14)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(shape = guide_legend(override.aes = list(size = 4), order = 1), 
           colour = guide_legend(override.aes = list(pch=16, size = 4), order = 2)) +
    ggtitle("Cytochrome C Oxidase")

plot2d
```


Create 4-panel figure of the 4 PCoA plots above:
```{r 4-panel PCoA plot}
#require(ggpubr) #load in library for multi-panel figures (if not already loaded)

# 4 panel metaG PCoA with History, start and finish sample (SSF)
MetaG_pcoa_4panel_SSF <- ggarrange(plot2a, plot2b, plot2c, plot2d, labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2, legend = "right", common.legend = T)

MetaG_pcoa_4panel_SSF

#ggsave("../figures/pub/metaG_pcoa4panel.tiff", plot = MetaG_pcoa_4panel_SSF, device="tiff", path=NULL, scale=1, height = 7, width = 10, limitsize = T, bg = "white", dpi = 300)

#ggsave("../figures/Thesis/metaG_pcoa4panel_SSF.jpg", plot = MetaG_pcoa_4panel_SSF, device=NULL, path=NULL, scale=1, height = 7, width = 10, limitsize = TRUE, dpi = 600)

# 4 panel metaG PCoA with History, Finish samples only (SFO)

MetaG_pcoa_4panel_SFO <- ggarrange(plot3a, plot3b, plot3c, plot3d, labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2, legend = "right", common.legend = T)

MetaG_pcoa_4panel_SFO

#ggsave("../figures/Thesis/metaG_pcoa4panel_SFO.jpg", plot = MetaG_pcoa_4panel_SFO, device=NULL, path=NULL, scale=1, height = 7, width = 10, limitsize = TRUE, dpi = 600)
```