---
title: "metaG"
author: "Colin Finlay"
date: "6/20/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#use to set global options for chunks e.g., echo and warning options will be applied to all chunks:
knitr::opts_chunk$set(echo = TRUE)

# Clear environment:
rm(list=ls())

# Set working directory:
setwd("~/GitHub/WetlandMesocosm_GHG_Timberlake/analyses")

# Standard error (se) and confidence interval (ci):
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code dependencies:
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("reshape"); require("ggplot2"): require("tidyr"); require("dplyr"); require("dbstats"); require("reshape2")
```

Import KO function files from JGI IMG/MER
```{r Import Files}
#Create vector with desired row order:
sample_order <- c("03NFC", "03PFC", "03Pre", "06NFC", "06PFC", "06Pre", "08NFC", "08PFC", "08Pre", "09NFC", "09PFC", "09Pre", "10NFC", "10PFC", "10Pre", "11NFC", "11PFC", "11Pre", "12NFC", "12PFC", "12Pre", "16NFC", "16PFC", "16Pre")
length(sample_order) #Should be 24

#Denitrification KEGG module:
denit <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Denit_CF.csv")
#Reorder rows:
denit <- denit[match(sample_order, denit$Sample), ]

#Methanogen Signature KEGG module:
meth <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Meth_CF.csv")
meth <- meth[match(sample_order, meth$Sample), ]

#All Central Carbohydrate Metabolism KEGG modules:
carb_metab <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/Central_carb_metab_CF.csv")
carb_metab <- carb_metab[match(sample_order, carb_metab$Sample), ]

#Cytochrome C Oxidase KEGG module:
cytoCox <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/Redownloaded/CSV/CytCOx_CF.csv")
cytoCox <- cytoCox[match(sample_order, cytoCox$Sample), ]

#Design file:
design <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/metaG/design.csv")
design2 <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/tmg_design.csv")
#Reorder rows of design:
design <- design[match(sample_order, design$Sample), ]

#Remove mock, field, and box samples from design2. 
rownames(design2) <- design2$sample
design2 <- design2[!(rownames(design2) %in% "MOCK"),]
design2 <- design2[!(rownames(design2) %in% "TL1"),]
design2 <- design2[!(rownames(design2) %in% "TL2"),]
design2 <- design2[!(rownames(design2) %in% "TL3"),]
design2 <- design2[!(rownames(design2) %in% "TL4"),]
design2 <- design2[!(rownames(design2) %in% "TL5"),]
design2 <- design2[!(rownames(design2) %in% "TL6"),]
design2 <- design2[!(design2$source %in% "b"),]
design2 <- design2[!(design2$source %in% "s"),]

#Greenhouse gas data:
GHG <- read.csv("~/GitHub/WetlandMesocosm_GHG_Timberlake/data/TMG_GHG-Flux_T1_Pub_2-2020.csv")

```

Convert gene counts to relative abundance:
```{r gene counts to relative abundance}
#Make sample IDs the rownames:
denitNumerical <- denit
rownames(denitNumerical) <- denitNumerical$Sample

methNumerical <- meth
rownames(methNumerical) <- methNumerical$Sample

carb_metabNumerical <- carb_metab
rownames(carb_metabNumerical) <- carb_metabNumerical$Sample

cytoCoxNumerical <- cytoCox
rownames(cytoCoxNumerical) <- cytoCoxNumerical$Sample

# Drop all non-numeric variables:
denitNumerical <- denitNumerical[,3:12]
methNumerical <- methNumerical[,3:46]
carb_metabNumerical <- carb_metabNumerical[,3:122]
cytoCoxNumerical <- cytoCoxNumerical[,3:7]

#Convert denitrification KO counts to relative abundance:
denitRA <- denitNumerical
for(i in 1:dim(denitNumerical)[1]){
  denitRA[i,] <- denitNumerical[i,]/sum(denitNumerical[i,])
}

#Convert methanogen KO counts to relative abundance:
methRA <- methNumerical
for(i in 1:dim(methNumerical)[1]){
  methRA[i,] <- methNumerical[i,]/sum(methNumerical[i,])
}

#Convert central carbohydrate metabolism KO counts to relative abundance:
carb_metabRA <- carb_metabNumerical
for(i in 1:dim(carb_metabNumerical)[1]){
  carb_metabRA[i,] <- carb_metabNumerical[i,]/sum(carb_metabNumerical[i,])
}

#Convert cytochrome c oxidase KO counts to relative abundance:
cytoCoxRA <- cytoCoxNumerical
for(i in 1:dim(cytoCoxNumerical)[1]){
  cytoCoxRA[i,] <- cytoCoxNumerical[i,]/sum(cytoCoxNumerical[i,])
}

#Subset start and finish samples from all gene sets (for respective start and finish GHG analysis):
denitRA_start <- denitRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
denitRA_finish <- denitRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

methRA_start <- methRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
methRA_finish <- methRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

carb_metabRA_start <- carb_metabRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
carb_metabRA_finish <- carb_metabRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]

cytoCoxRA_start <- cytoCoxRA[c("03Pre", "06Pre", "08Pre", "09Pre", "10Pre", "11Pre", "12Pre", "16Pre"), ]
cytoCoxRA_finish <- cytoCoxRA[c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC", "09PFC", "10NFC", "10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC"), ]
#Check dimensions of each relative abundance (RA) matrix:
dim(denitRA) #24 10
dim(methRA) #24 44
dim(carb_metabRA) #24 120
dim(cytoCoxRA) #24 5

dim(denitRA_start) #8 10
dim(methRA_start) #8 44
dim(carb_metabRA_start) #8 120
dim(cytoCoxRA_start) #8 5

dim(denitRA_finish) #16 10
dim(methRA_finish) #16 44
dim(carb_metabRA_finish) #16 120
dim(cytoCoxRA_finish) #16 5
```

Calculate Bray Curtis dissimilarity:
```{r relative abundance to Bray Curtis dissimilarity}
#Bray Curtis calculation for denitrification relative abundance data:
denitBC <- vegdist(denitRA, method = "bray")
denitBC_start <- vegdist(denitRA_start, method = "bray")
denitBC_finish <- vegdist(denitRA_finish, method = "bray")
#print(denitBC)

#Bray Curtis calculation for methanogen relative abundance data:
methBC <- vegdist(methRA, method = "bray")
methBC_start <- vegdist(methRA_start, method = "bray")
methBC_finish <- vegdist(methRA_finish, method = "bray")
#print(methBC)

#Bray Curtis calculation for central carbohydrate relative abundance data:
carb_metabBC <- vegdist(carb_metabRA, method = "bray")
carb_metabBC_start <- vegdist(carb_metabRA_start, method = "bray")
carb_metabBC_finish <- vegdist(carb_metabRA_finish, method = "bray")
#print(carb_metabBC)

#Bray Curtis calculation for cytochrome C oxidase relative abundance data:
cytoCoxBC <- vegdist(cytoCoxRA, method = "bray")
cytoCoxBC_start <- vegdist(cytoCoxRA_start, method = "bray")
cytoCoxBC_finish <- vegdist(cytoCoxRA_finish, method = "bray")
#print(cytoCoxBC)
```

Wrangling GHG data:
```{r wrangling GHG data}
#Create a new GHG data frame with less columns:
GHG_less_columns <- cbind(GHG[,1:4], GHG[24:32])

#Drop rows with NAs (non-T1 rows)
GHG_less_columns <- GHG_less_columns %>% drop_na()

#Change column name "chamber" to "box" to match with design2
colnames(GHG_less_columns)[2] <- "box"

#Convert GHG_less_columns$plant to lowercase to match with design2$plant:
GHG_less_columns$plant <- tolower(GHG_less_columns$plant)

#Subset GHG_less_columns for 'start' time point (6/13/2016) and 'finish' time point (8/11/2016):
GHG_start <- GHG_less_columns[1:36,]
GHG_finish <- GHG_less_columns[157:192,]

#Add design2 into GHG 'start' and 'finish':
GHG_start <- left_join(GHG_start, design2, by = c("box", "plant"))
GHG_finish <- left_join(GHG_finish, design2, by = c("box", "plant"))

#Subset GHG_finish and GHG_start to only have metaG sample numbers:
GHG_start_metaG <- rbind(GHG_start[5:6,], GHG_start[11:12,], GHG_start[15:24,], GHG_start[31:32,], make.row.names = FALSE)
GHG_finish_metaG <- rbind(GHG_finish[5:6,], GHG_finish[11:12,], GHG_finish[15:24,], GHG_finish[31:32,], make.row.names = FALSE)

#To align "pre" metaG samples with a single, initial GHG values (from the start GHG measurements on 6/13/206), average together GHG samples in the same box number, then transpose and set row name to corresponding metaG Sample ID:
GHG_start_03 <- data.frame(colMeans(GHG_start_metaG[1:2, 5:13]))
GHG_start_03 <- as.data.frame(t(GHG_start_03))
row.names(GHG_start_03) <- "03Pre"

GHG_start_06 <- data.frame(colMeans(GHG_start_metaG[3:4, 5:13]))
GHG_start_06 <- as.data.frame(t(GHG_start_06))
row.names(GHG_start_06) <- "06Pre"

GHG_start_08 <- data.frame(colMeans(GHG_start_metaG[5:6, 5:13]))
GHG_start_08 <- as.data.frame(t(GHG_start_08))
row.names(GHG_start_08) <- "08Pre"

GHG_start_09 <- data.frame(colMeans(GHG_start_metaG[7:8, 5:13]))
GHG_start_09 <- as.data.frame(t(GHG_start_09))
row.names(GHG_start_09) <- "09Pre"

GHG_start_10 <- data.frame(colMeans(GHG_start_metaG[9:10, 5:13]))
GHG_start_10 <- as.data.frame(t(GHG_start_10))
row.names(GHG_start_10) <- "10Pre"

GHG_start_11 <- data.frame(colMeans(GHG_start_metaG[11:12, 5:13]))
GHG_start_11 <- as.data.frame(t(GHG_start_11))
row.names(GHG_start_11) <- "11Pre"

GHG_start_12 <- data.frame(colMeans(GHG_start_metaG[13:14, 5:13]))
GHG_start_12 <- as.data.frame(t(GHG_start_12))
row.names(GHG_start_12) <- "12Pre"

GHG_start_16 <- data.frame(colMeans(GHG_start_metaG[15:16, 5:13]))
GHG_start_16 <- as.data.frame(t(GHG_start_16))
row.names(GHG_start_16) <- "16Pre"

#Then rbind() all the dfs together
GHG_start_metaG_avg <- rbind(GHG_start_03, GHG_start_06, GHG_start_08, GHG_start_09, GHG_start_10, GHG_start_11, GHG_start_12, GHG_start_16)

#Set the row names of GHG data to the metaG sample names, so that stats functions can compare the GHG and metaG data sets:
row.names(GHG_finish_metaG) <- c("03NFC", "03PFC", "06NFC", "06PFC", "08NFC", "08PFC", "09NFC","09PFC", "10NFC","10PFC", "11NFC", "11PFC", "12NFC", "12PFC", "16NFC", "16PFC")

#Combine GHG_start_metaG_avg and GHG_finish_metaG[,5:13] to have all 24 samples:
GHG_metaG <- rbind(GHG_start_metaG_avg, GHG_finish_metaG[,5:13])
GHG_metaG <- GHG_metaG[match(sample_order, rownames(GHG_metaG)), ]
GHG_metaG <- GHG_metaG[,c(2,5,8)] #Subset to only have units in mg
```

Distance-based partial least squares regression:
```{r distance-based partial least squares regression}
library(dbstats)

start.denit.dbplsr <- dbplsr(GHG_start_metaG_avg$mg.N2O.N.m.2 ~ as.matrix(denitBC_start), ncomp = 7, method = "GCV")
summary(start.denit.dbplsr)
#plot(start.denit.dbplsr)

finish.denit.dbplsr <- dbplsr(GHG_finish_metaG$mg.N2O.N.m.2 ~ as.matrix(denitBC_finish), ncomp = 15, method = "GCV")
summary(finish.denit.dbplsr)
#plot(finish.denit.dbplsr)

start.meth.dbplsr <- dbplsr(GHG_start_metaG_avg$mg.CH4.N.m.2 ~ as.matrix(methBC_start), ncomp = 7, method = "GCV")
summary(start.meth.dbplsr)
#plot(start.meth.dbplsr)

finish.meth.dbplsr <- dbplsr(GHG_finish_metaG$mg.CH4.N.m.2 ~ as.matrix(methBC_finish), ncomp = 15, method = "GCV")
summary(finish.meth.dbplsr)
#plot(finish.meth.dbplsr)

start.carb_metab.dbplsr <- dbplsr(GHG_start_metaG_avg$mg.CO2.N.m.2 ~ as.matrix(carb_metabBC_start), ncomp = 7, method = "GCV")
summary(start.carb_metab.dbplsr)
#plot(start.carb_metab.dbplsr)

finish.carb_metab.dbplsr <- dbplsr(GHG_finish_metaG$mg.CO2.N.m.2 ~ as.matrix(carb_metabBC_finish), ncomp = 15, method = "GCV")
summary(finish.carb_metab.dbplsr)
#plot(finish.carb_metab.dbplsr)

start.cytoCox.dbplsr <- dbplsr(GHG_start_metaG_avg$mg.CO2.N.m.2 ~ as.matrix(cytoCoxBC_start), ncomp = 7, method = "GCV")
summary(start.cytoCox.dbplsr)
#plot(start.cytoCox.dbplsr)

finish.cytoCox.dbplsr <- dbplsr(GHG_finish_metaG$mg.CO2.N.m.2 ~ as.matrix(cytoCoxBC_finish), ncomp = 15, method = "GCV")
summary(finish.cytoCox.dbplsr)
#plot(finish.cytoCox.dbplsr)

# Because of envfit() results showing significant relationship between denitrification composition and [CO2], running dbplsrs for denit and CO2:
start.denit.dbplsr.co2 <- dbplsr(GHG_start_metaG_avg$mg.CO2.N.m.2 ~ as.matrix(denitBC_start), ncomp = 7, method = "GCV")
summary(start.denit.dbplsr.co2)
#plot(start.denit.dbplsr.co2)

finish.denit.dbplsr.co2 <- dbplsr(GHG_finish_metaG$mg.CO2.N.m.2 ~ as.matrix(denitBC_finish), ncomp = 15, method = "GCV")
summary(finish.denit.dbplsr.co2)
#plot(finish.denit.dbplsr.co2)
```

PERMANOVA:
```{r PERMANOVA}
#Make metaG sample ID's the rownames on 'design'. Then drop design$Sample:
rownames(design) <- design$Sample
design <- design[,2:4]

# Make data frames from gene set relative abundance (RA) files and design
denitRA_design <- cbind(design, denitRA)
methRA_design <- cbind(design, methRA)
carb_metabRA_design <- cbind(design, carb_metabRA)
cytoCoxRA_design <- cbind(design, cytoCoxRA)

#PERMANOVA:
adonis_denit = adonis(denitRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = denitRA_design, perm=10000)
adonis_denit

adonis_meth = adonis(methRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = methRA_design, perm=10000)
adonis_meth

adonis_carb_metab = adonis(carb_metabRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = carb_metabRA_design, perm=10000)
adonis_carb_metab

adonis_cytoCox = adonis(cytoCoxRA_design[,-c(1:3)]~Plant*Treat, method = "bray", data = cytoCoxRA_design, perm=10000)
adonis_cytoCox
```

Principal Coordinates Analysis (PCoA) denitrification:
```{r PCoA:denitrification}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_denit <- cmdscale(denitBC, k=3, eig=TRUE, add=FALSE)

explainvar1b_denit <- round(pcoa_denit$eig[1] / sum(pcoa_denit$eig), 3) * 100
explainvar2b_denit <- round(pcoa_denit$eig[2] / sum(pcoa_denit$eig), 3) * 100
sum.eigb <- sum(explainvar1b_denit, explainvar2b_denit)

explainvar1b_denit # 44.6
explainvar2b_denit # 25.1

pcoa_denit.groups <- paste(denitRA_design$Treat, denitRA_design$Plant, sep = "_")
pcoa_denit.points <- data.frame(pcoa_denit$points, group = pcoa_denit.groups)

# Calculate Centroids (mean and SE)
pcoa_denit.L.centroids <- reshape2::melt(pcoa_denit.points, id="group", measure.vars = c("X1", "X2"))
pcoa_denit.centroids <- acast(pcoa_denit.L.centroids, variable ~ group, mean)
pcoa_denit.centroids.se <- acast(pcoa_denit.L.centroids, variable ~ group, se)
pcoa_denit.centroids.sd <- acast(pcoa_denit.L.centroids, variable ~ group, sd)

# Combine
pcoa_denit.cent.dataframe <- cbind(t(pcoa_denit.centroids), t(pcoa_denit.centroids.se))
colnames(pcoa_denit.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

trt <- c("baseline","dry","dry","wet","wet") #this should match the treatments from pcoa_denit.cent.dataframe
plant <- c("baseline","no plant","plant","no plant","plant") #this should match the treatments from pcoa_denit.cent.dataframe

pcoa_denit.cent.dataframe.trts <- as.data.frame(pcoa_denit.cent.dataframe) 
pcoa_denit.cent.dataframe.trts$trt <- as.factor(trt) # now you add the new columns that you made above to the pcoa.cent.dataframe.trts df
pcoa_denit.cent.dataframe.trts$plant <- as.factor(plant)

#Run envfit to see which [GHG]s (start or finish) are explained by ordination:
fit_denit <- envfit(pcoa_denit$points, GHG_metaG, permutations = 10000)
fit_denit

A_denit <-as.list(fit_denit$vectors)
vec_denit <- as.data.frame(fit_denit$vectors$arrows*sqrt(fit_denit$vectors$r)*0.2) #This calculation is adjusting the length of the arrow that ends up on the plot.
p_denit <- as.data.frame(A_denit$pvals)
vec_denit <- cbind(vec_denit, p_denit)
vec_denit <- subset(vec_denit, A_denit$pvals<=0.075)

#Plot using ggplot2
df1a <- as.data.frame(pcoa_denit.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt, shape = plant)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_segment(data=vec_denit, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, 
      arrow = arrow(length = unit(0.2, "cm")),colour="black",inherit.aes=F)+
  geom_text(data=vec_denit, aes(x=0.0125,y=-0.075), label= expression(paste("mg CO"[2],"-C m"^{-2})), size=5, inherit.aes=F)+
  geom_text(x = 0.03, y = -0.031, label = expression(paste("plant: R"^{2}," = 0.09, P = 0.48")), show.legend = FALSE, colour="black")+
  geom_text(x=0.03, y=-0.038, label=expression(paste("treatment: R"^{2}," = 0.03, P = 0.56")), show.legend = FALSE, colour="black")+
  geom_text(x=0.028, y=-0.045, label=expression(paste("plant:treatment : R"^{2}," = 0.02, P = 0.86")), show.legend = FALSE, colour="black")+
  scale_colour_manual(labels = c("baseline","dry","wet"), 
                      values = c("gray70", "cyan", "darkblue")) +
  scale_shape_manual(labels = c("baseline","no plant","plant"), values = c(16,22,15)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (44.6%)") + ylab("PCoA 2 (25.1%)") +
        labs(colour = "Hydrologic Treatment", shape = "Plant") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Denitrifier Community Composition")

ggsave("../figures/MetaG_PCoAs/DNFcomposition1a.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=8, height=5, dpi=300, limitsize=TRUE)
```

Principal Coordinates Analysis (PCoA) methanogenesis:
```{r PCoA:methanogenesis}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_meth <- cmdscale(methBC, k=3, eig=TRUE, add=FALSE)

explainvar1b_meth <- round(pcoa_meth$eig[1] / sum(pcoa_meth$eig), 3) * 100
explainvar2b_meth <- round(pcoa_meth$eig[2] / sum(pcoa_meth$eig), 3) * 100
sum.eigb <- sum(explainvar1b_meth, explainvar2b_meth)

explainvar1b_meth # 41.3
explainvar2b_meth # 27.8

pcoa_meth.groups <- paste(methRA_design$Treat, methRA_design$Plant, sep = "_")
pcoa_meth.points <- data.frame(pcoa_meth$points, group = pcoa_meth.groups)

# Calculate Centroids (mean and SE)
pcoa_meth.L.centroids <- reshape2::melt(pcoa_meth.points, id="group", measure.vars = c("X1", "X2"))
pcoa_meth.centroids <- acast(pcoa_meth.L.centroids, variable ~ group, mean)
pcoa_meth.centroids.se <- acast(pcoa_meth.L.centroids, variable ~ group, se)
pcoa_meth.centroids.sd <- acast(pcoa_meth.L.centroids, variable ~ group, sd)

# Combine
pcoa_meth.cent.dataframe <- cbind(t(pcoa_meth.centroids), t(pcoa_meth.centroids.se))
colnames(pcoa_meth.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

trt <- c("baseline","dry","dry","wet","wet") 
plant <- c("baseline","no plant","plant","no plant","plant")

pcoa_meth.cent.dataframe.trts <- as.data.frame(pcoa_meth.cent.dataframe) 
pcoa_meth.cent.dataframe.trts$trt <- as.factor(trt)
pcoa_meth.cent.dataframe.trts$plant <- as.factor(plant)

#Run envfit to see which [GHG]s (start or finish) are explained by ordination:
fit_meth <- envfit(pcoa_meth$points, GHG_metaG, permutations = 1000)
fit_meth

A_meth <-as.list(fit_meth$vectors)
vec_meth <- as.data.frame(fit_meth$vectors$arrows*sqrt(fit_meth$vectors$r)*0.15) #Adjust arrow length
p_meth <- as.data.frame(A_meth$pvals)
vec_meth <- cbind(vec_meth, p_meth)
vec_meth <- subset(vec_meth, A_meth$pvals<=0.1)

#Plot using ggplot2
df2a <- as.data.frame(pcoa_meth.cent.dataframe.trts)
plot2a <- ggplot(df2a, aes(x=V1, y=V2, colour=trt, shape = plant)) + theme_bw() 
plot2a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_segment(data=vec_meth, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, 
      arrow = arrow(length = unit(0.2, "cm")),colour="black",inherit.aes=F)+
  geom_text(data=vec_meth, aes(x=-0.05,y=0.015), angle=-26, label= expression(paste("mg CH"[4],"-C m"^{-2})), size=5, inherit.aes=F)+
  geom_text(x = -0.04, y = 0.044, label = expression(paste("plant: R"^{2}," = 0.08, P = 0.47")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=-0.035, y=0.040, label=expression(paste("treatment: R"^{2}," = 0.05, P = 0.36")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=-0.03, y=0.036, label=expression(paste("plant:treatment : R"^{2}," = 0.02, P = 0.74")), show.legend = FALSE, colour="black", inherit.aes=F)+
  scale_colour_manual(labels = c("baseline","dry","wet"), 
                      values = c("gray70", "cyan", "darkblue")) +
  scale_shape_manual(labels = c("baseline","no plant","plant"), values = c(16,22,15)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (41.3%)") + ylab("PCoA 2 (27.8%)") +
        labs(colour = "Hydrologic Treatment", shape = "Plant") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Methanogen Community Composition")

ggsave("../figures/MetaG_PCoAs/METHcomposition2a.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=8, height=5, dpi=300, limitsize=TRUE)
```

Principal Coordinates Analysis (PCoA) central carbohydrate metabolism:
```{r PCoA:central carbohydrate metabolism}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_carb_metab <- cmdscale(carb_metabBC, k=3, eig=TRUE, add=FALSE)

explainvar1b_carb_metab <- round(pcoa_carb_metab$eig[1] / sum(pcoa_carb_metab$eig), 3) * 100
explainvar2b_carb_metab <- round(pcoa_carb_metab$eig[2] / sum(pcoa_carb_metab$eig), 3) * 100
sum.eigb <- sum(explainvar1b_carb_metab, explainvar2b_carb_metab)

explainvar1b_carb_metab # 33.7
explainvar2b_carb_metab # 15.9

pcoa_carb_metab.groups <- paste(carb_metabRA_design$Treat, carb_metabRA_design$Plant, sep = "_")
pcoa_carb_metab.points <- data.frame(pcoa_carb_metab$points, group = pcoa_carb_metab.groups)

# Calculate Centroids (mean and SE)
pcoa_carb_metab.L.centroids <- reshape2::melt(pcoa_carb_metab.points, id="group", measure.vars = c("X1", "X2"))
pcoa_carb_metab.centroids <- acast(pcoa_carb_metab.L.centroids, variable ~ group, mean)
pcoa_carb_metab.centroids.se <- acast(pcoa_carb_metab.L.centroids, variable ~ group, se)
pcoa_carb_metab.centroids.sd <- acast(pcoa_carb_metab.L.centroids, variable ~ group, sd)

# Combine
pcoa_carb_metab.cent.dataframe <- cbind(t(pcoa_carb_metab.centroids), t(pcoa_carb_metab.centroids.se))
colnames(pcoa_carb_metab.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

trt <- c("baseline","dry","dry","wet","wet")
plant <- c("baseline","no plant","plant","no plant","plant")

pcoa_carb_metab.cent.dataframe.trts <- as.data.frame(pcoa_carb_metab.cent.dataframe) 
pcoa_carb_metab.cent.dataframe.trts$trt <- as.factor(trt)
pcoa_carb_metab.cent.dataframe.trts$plant <- as.factor(plant)

#Run envfit to see which [GHG]s (start or finish) are explained by ordination:
fit_carb_metab <- envfit(pcoa_carb_metab$points, GHG_metaG, permutations = 10000)
fit_carb_metab

A_carb_metab <-as.list(fit_carb_metab$vectors)
vec_carb_metab <- as.data.frame(fit_carb_metab$vectors$arrows*sqrt(fit_carb_metab$vectors$r)*0.1) #Adjust arrow length
p_carb_metab <- as.data.frame(A_carb_metab$pvals)
vec_carb_metab <- cbind(vec_carb_metab, p_carb_metab)
vec_carb_metab <- subset(vec_carb_metab, A_carb_metab$pvals<=0.075)

#Plot using ggplot2
df3a <- as.data.frame(pcoa_carb_metab.cent.dataframe.trts)
plot3a <- ggplot(df3a, aes(x=V1, y=V2, colour=trt, shape = plant)) + theme_bw() 
plot3a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.001), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_segment(data=vec_carb_metab, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, 
      arrow = arrow(length = unit(0.2, "cm")),colour="black",inherit.aes=F)+
  geom_text(data=vec_carb_metab, aes(x=-0.02,y=0.03), angle=-42, label= expression(paste("mg CO"[2],"-C m"^{-2})), size=5, inherit.aes=F)+
  geom_text(x=0.0, y=0.04, label = expression(paste("plant: R"^{2}," = 0.07, P = 0.69")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=0.0, y=0.036, label=expression(paste("treatment: R"^{2}," = 0.09, P = 0.04*")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=0.0, y=0.032, label=expression(paste("plant:treatment : R"^{2}," = 0.04, P = 0.54")), show.legend = FALSE, colour="black", inherit.aes=F)+
  scale_colour_manual(labels = c("baseline","dry","wet"), 
                      values = c("gray70", "cyan", "darkblue")) +
  scale_shape_manual(labels = c("baseline","no plant","plant"), values = c(16,22,15)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (33.7%)") + ylab("PCoA 2 (15.9%)") +
        labs(colour = "Hydrologic Treatment", shape = "Plant") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Central Carbohydrate Metabolism Community Composition")

ggsave("../figures/MetaG_PCoAs/CARBcomposition3a.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=8, height=5, dpi=300, limitsize=TRUE)
```

Principal Coordinates Analysis (PCoA) cytochrome C oxidase:
```{r PCoA:cytochrome C oxidase}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
pcoa_cytoCox <- cmdscale(cytoCoxBC, k=3, eig=TRUE, add=FALSE)

explainvar1b_cytoCox <- round(pcoa_cytoCox$eig[1] / sum(pcoa_cytoCox$eig), 3) * 100
explainvar2b_cytoCox <- round(pcoa_cytoCox$eig[2] / sum(pcoa_cytoCox$eig), 3) * 100
sum.eigb <- sum(explainvar1b_cytoCox, explainvar2b_cytoCox)

explainvar1b_cytoCox # 50.7
explainvar2b_cytoCox # 35.6

pcoa_cytoCox.groups <- paste(cytoCoxRA_design$Treat, cytoCoxRA_design$Plant, sep = "_")
pcoa_cytoCox.points <- data.frame(pcoa_cytoCox$points, group = pcoa_cytoCox.groups)

# Calculate Centroids (mean and SE)
pcoa_cytoCox.L.centroids <- reshape2::melt(pcoa_cytoCox.points, id="group", measure.vars = c("X1", "X2"))
pcoa_cytoCox.centroids <- acast(pcoa_cytoCox.L.centroids, variable ~ group, mean)
pcoa_cytoCox.centroids.se <- acast(pcoa_cytoCox.L.centroids, variable ~ group, se)
pcoa_cytoCox.centroids.sd <- acast(pcoa_cytoCox.L.centroids, variable ~ group, sd)

# Combine
pcoa_cytoCox.cent.dataframe <- cbind(t(pcoa_cytoCox.centroids), t(pcoa_cytoCox.centroids.se))
colnames(pcoa_cytoCox.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

trt <- c("baseline","dry","dry","wet","wet")
plant <- c("baseline","no plant","plant","no plant","plant")

pcoa_cytoCox.cent.dataframe.trts <- as.data.frame(pcoa_cytoCox.cent.dataframe) 
pcoa_cytoCox.cent.dataframe.trts$trt <- as.factor(trt)
pcoa_cytoCox.cent.dataframe.trts$plant <- as.factor(plant)

#Run envfit to see which [GHG]s (start or finish) are explained by ordination:
fit_cytoCox <- envfit(pcoa_cytoCox$points, GHG_metaG, permutations = 10000)
fit_cytoCox #All P-values are large (>0.075). So, stopping envfit for Cytochrome C Oxidase here.

#Plot using ggplot2
df4a <- as.data.frame(pcoa_cytoCox.cent.dataframe.trts)
plot4a <- ggplot(df4a, aes(x=V1, y=V2, colour=trt, shape = plant)) + theme_bw() 
plot4a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_text(x=-0.015, y=0.013, label = expression(paste("plant: R"^{2}," = 0.08, P = 0.51")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=-0.015, y=0.011, label=expression(paste("treatment: R"^{2}," = 0.03, P = 0.55")), show.legend = FALSE, colour="black", inherit.aes=F)+
  geom_text(x=-0.014, y=0.009, label=expression(paste("plant:treatment : R"^{2}," = 0.08, P = 0.17")), show.legend = FALSE, colour="black", inherit.aes=F)+
  scale_colour_manual(labels = c("baseline","dry","wet"), 
                      values = c("gray70", "cyan", "darkblue")) +
  scale_shape_manual(labels = c("baseline","no plant","plant"), values = c(16,22,15)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (50.7%)") + ylab("PCoA 2 (35.6%)") +
        labs(colour = "Hydrologic Treatment", shape = "Plant") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Cytochrome C Oxidase Community Composition")

ggsave("../figures/MetaG_PCoAs/CcOcomposition4a.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=8, height=5, dpi=300, limitsize=TRUE)
```

Creating an ordination of 16S data that doesn't include 'interim' but does test envfit vectors for GHG data. This will be done by modifying the code in 'timberlake_publicationV2.Rmd', lines 0 to 310:
```{r setup for 16S data}
#Assign file names to variables
sharedfile = "../data/tmg2016.with.archaeaopti_mcc.shared"
metafile = "../data/tmg_design.csv"

#Read in OTU file
otu <- read.otu(sharedfile)
rownames(otu)
#remove mock
#otu.2 <- otu.2[!(rownames(otu.2) %in% "MOCK"),]
#remove rogue sample... looks like there was a naming error at some point resulting in the additional sample below
otu.2 <- otu[!(rownames(otu) %in% "13PN"),]

#Removing microbial samples from "box" soils and keeping samples collected from within the chamber area since these should more directly relate to GHG flux data
c <- grep("*B", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#REMOVING STARTING SAMPLES
c <- grep("*S", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#Remove Field samples. 
otu.2.nofield <- otu.2[!(rownames(otu.2) %in% "TL1"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL2"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL3"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL4"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL5"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL6"),]

otu.2 <- otu.2.nofield

#Remove OTUs with less than 10 occurrences across all sites
b <- ncol(otu.2) #67355 OTUs,
s<- sum(colSums(otu.2)) #1570135
otu.2 <- otu.2[, which(colSums(otu.2) >= 10)]
a <- ncol(otu.2) #7026 OTUs
sf<- sum(colSums(otu.2)) #1505797

#What sample has lowest read count? 
otu2<- colSums(t(otu.2))
otu2[which.min(colSums(t(otu.2)))]
otu2[which.max(colSums(t(otu.2)))]
#otu2

#Lowest sample is 8NS 20370, highest 16NS 154034
#Most sample fall between 20370 and ~60000 with only a few higher than that so not removing the lowest sample 

#graph of read counts, need to work on display
d<-as.data.frame(otu.2)
d$colsum <- colSums(t(d))
d$group <- rownames(d)
p<-ggplot(d, aes(x=group,y=colsum))+geom_bar(stat="identity")+coord_flip()
p

#Rarefy
#Set min sample number
min.N <- min(rowSums(otu.2))
#rrarefy returns a randomly rarefied community data frame or vector of selected size
#This is the otu set that should be used for beta diversity metrics
otu.2 <- rrarefy(otu.2, min.N)

#Read in design file
meta <- read.csv(metafile)
#Combine meta data with relative abundance data
#Remove mock, field, and box samples. 
rownames(meta) <- meta$sample
meta <- meta[!(rownames(meta) %in% "MOCK"),]
meta <- meta[!(rownames(meta) %in% "TL1"),]
meta <- meta[!(rownames(meta) %in% "TL2"),]
meta <- meta[!(rownames(meta) %in% "TL3"),]
meta <- meta[!(rownames(meta) %in% "TL4"),]
meta <- meta[!(rownames(meta) %in% "TL5"),]
meta <- meta[!(rownames(meta) %in% "TL6"),]
meta <- meta[!(meta$source %in% "b"),]
meta <- meta[!(meta$source %in% "s"),]
```

```{r 16S read counts to relative abundance and 16S PERMANOVA}
#transforms read counts to relative abundance
otu.2.rel <- otu.2
for(i in 1:dim(otu.2)[1]){
  otu.2.rel[i,] <- otu.2[i,]/sum(otu.2[i,])
}

rownames(meta) <- rownames(otu.2.rel)
all.equal(rownames(otu.2.rel), rownames(meta))
metaotu <- cbind(meta, otu.2.rel)

#PERMANOVAs:
metaotu.ad <- adonis(metaotu[,-c(1:9)] ~ plant * history + plant * treatment, method = "bray", data = metaotu, perm=1000, set.seed=42)
metaotu.ad 

metaotu.ad.plant.trt <- adonis(metaotu[,-c(1:9)] ~ plant * treatment, method = "bray", data = metaotu, permutations = 10000)
metaotu.ad.plant.trt
```


```{r PCoA: Bray-Curtis matrix of 16S data}
otu.2.rel.dist<-vegdist(otu.2.rel, method="bray") #Calculate Bray-Curtis dissimilarity of 16S data

# Principal Coordinates Analysis
otu.2.rel.dist.pcoa <- cmdscale(otu.2.rel.dist, k=3, eig=TRUE, add=FALSE)
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1a <- round(otu.2.rel.dist.pcoa$eig[1] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
explainvar2a <- round(otu.2.rel.dist.pcoa$eig[2] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
sum.eiga <- sum(explainvar1a, explainvar2a)

explainvar1a #24.8%
explainvar2a #10.6%

all.equal(rownames(meta), rownames(otu.2.rel))

pcoa_16S.groups <- paste(meta$treatment, meta$plant, sep = "_")

pcoa_16S.points <- data.frame(otu.2.rel.dist.pcoa$points, group = pcoa_16S.groups)

# Calculate Centroids (mean and SE)
pcoa_16S.L.centroids <- melt(pcoa_16S.points, id="group", measure.vars = c("X1", "X2"))
pcoa_16S.centroids <- acast(pcoa_16S.L.centroids, variable ~ group, mean)
pcoa_16S.centroids.se <- acast(pcoa_16S.L.centroids, variable ~ group, se)
pcoa_16S.centroids.sd <- acast(pcoa_16S.L.centroids, variable ~ group, sd)

# Combine
pcoa_16S.cent.dataframe <- cbind(t(pcoa_16S.centroids), t(pcoa_16S.centroids.se))
colnames(pcoa_16S.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

#Lining up treatments:
pcoa_16S.cent.dataframe.trts <- as.data.frame(pcoa_16S.cent.dataframe)
#Subset to remove Interim (I) samples:
pcoa_16S.cent.dataframe.trts <- pcoa_16S.cent.dataframe.trts[c(1,2,5,6), ]

trt_16S <- c("dry", "dry", "wet", "wet")
plant_16S <- c("no plant","plant","no plant","plant")

pcoa_16S.cent.dataframe.trts$trt <- as.factor(trt_16S)
pcoa_16S.cent.dataframe.trts$plant <- as.factor(plant_16S)

#trying to line up GHG_finish with 16S samples:
GHG_16S <- GHG_finish
sixteenS_rows <- row.names(otu.2.rel)
row.names(GHG_16S) <- sixteenS_rows
#Subset to only have mg of GHGs:
GHG_16S <- GHG_16S[, c(6,9,12)]

#Run envfit to see which [GHG]s (start or finish) are explained by ordination:
fit_16S <- envfit(otu.2.rel.dist.pcoa$points, GHG_16S, permutations = 1000, na.rm = T) #I would need to make a new GHG_16S file that lines up with 16S samples.
fit_16S # No significant P-values, so not proceeding with envfit plotting for now

#Create a string with 16S PERMANOVA R2 and p-values:
stats_label_16S <- expression(paste("R"^{2}," = 0.02, P = 0.98",
                                    "R"^{2}," = 0.12, P < 0.01**",
                                    "R"^{2}," =0.02, P = 1.00", sep = "/n"))

print(stats_label_16S)
#expression(paste("mg N"[2],"O m"^{2}))

#Plotting:
df5a <- as.data.frame(pcoa_16S.cent.dataframe.trts)
plot5a <- ggplot(df5a, aes(x=V1, y=V2, colour=trt, shape = plant)) + theme_bw() 
plot5a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.005), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.005), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_text(x =-0.06, y = -0.03, label = expression(paste("plant: R"^{2}," = 0.02, P = 0.98")), show.legend = FALSE, colour="black")+
  geom_text(x=-0.055, y=-0.036, label=expression(paste("treatment: R"^{2}," = 0.12, P = 0.01**")), show.legend = FALSE, colour="black")+
  geom_text(x=-0.05, y=-0.042, label=expression(paste("plant:treatment : R"^{2}," = 0.02, P = 1.00")), show.legend = FALSE, colour="black")+
  scale_colour_manual(labels = c("dry","wet"), 
                      values = c("cyan", "darkblue")) +
  scale_shape_manual(labels = c("no plant","plant"), values = c(22,15)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (24.8%)") + ylab("PCoA 2 (10.6%)") +
        labs(colour = "Hydrologic Treatment", shape = "Plant") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("16S rRNA Community Composition")

ggsave("../figures/16SrRNA_timberlake_end_NO_INTERIM.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=8, height=5, dpi=300, limitsize=TRUE)
```
