---
title: "TOWeR GHG microcosm experiment - Summer 2016"
author: "Regina B. Bledsoe, Colin G. Finlay, Ariane L. Peralta"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Setup

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#use to set working directory 
knitr::opts_knit$set(root.dir="~/GitHub/WetlandMesocosm_GHG_Timberlake/analyses")
```

# Load packages and functions

```{r load packages and functions, include FALSE }
#Set source R tools
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")

#load required packages
require("vegan")
require("tidyverse")
require("nlme")
require("reshape2")
require("ecodist")
require("MASS")
require("phyloseq")
require("MuMIn")
require("AICcmodavg")
require("emmeans")
require("lme4")
require("car")
require("dbstats")
require("plyr")
require("ggpubr")
require("labdsv")
require("grid")
require("multcomp")
require("multcompView")

```

```{r Phyloseq install and load}
# If phyloseq not installed, un-comment and run:
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("phyloseq")

require("phyloseq")
```

```{r Standard error and confidence intervals}
#Set Std Err and Conf Int
se <- function(x, ...) {
  sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
  1.96 * sd(x, na.rm = TRUE)
}
```
# Load files

```{r load files paths}
#Load data files
#Assign file names to variables
sharedfile = "../data/tmg2016.with.archaeaopti_mcc.shared"
taxfile = "../data/tmg2016.with.archaea.opti_mcc.0.03.cons.taxonomy"
metafile = "../data/tmg_design.csv"
```

```{r read in otu file and trim}
#Read in OTU file
otu <- read.otu(sharedfile)
rownames(otu)

#remove rogue sample... looks like there was a naming error at some point resulting in the additional sample below
otu.2 <- otu[!(rownames(otu) %in% "13PN"),]

#Removing microbial samples from "box" soils and keeping samples collected from within the chamber area since these should more directly relate to GHG flux data
c <- grep("*B", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#REMOVING STARTING SAMPLES
c <- grep("*S", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#Remove Field samples. 
otu.2.nofield <- otu.2[!(rownames(otu.2) %in% "TL1"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL2"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL3"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL4"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL5"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL6"),]

otu.2 <- otu.2.nofield

#Remove OTUs with less than 10 occurrences across all sites
b <- ncol(otu.2) #67355 OTUs,

# !=0 35,897
# =0 31,458
# =1 15,582
# =2  5,536
# =3  2,580
# =4  1,613
# =5  1,029
# =6    853
# =7    669
# =8    535
# =9    474
#=10    402

s<- sum(colSums(otu.2)) #1570135
otu.2 <- otu.2[, which(colSums(otu.2) >= 10)]
a <- ncol(otu.2) #7026 OTUs
sf<- sum(colSums(otu.2)) #1505797

#What sample has lowest read count? 
otu2<- colSums(t(otu.2))
otu2[which.min(colSums(t(otu.2)))]
otu2[which.max(colSums(t(otu.2)))]

#Lowest sample is 8NS 20370, highest 16NS 154034
#Most sample fall between 20370 and ~60000 with only a few higher than that so not removing the lowest sample 

#graph of read counts
d<-as.data.frame(otu.2)
d$colsum <- colSums(t(d))
d$group <- rownames(d)
p<-ggplot(d, aes(x=group,y=colsum))+geom_bar(stat="identity")+coord_flip()
p

#Rarefy
#Set min sample number
min.N <- min(rowSums(otu.2))
#rrarefy returns a randomly rarefied community data frame or vector of selected size
#This is the otu set that should be used for beta diversity metrics
otu.2 <- rrarefy(otu.2, min.N)
```

```{r read in meta data}
#Read in design file
meta <- read.csv(metafile)

#Remove mock, field, and box samples. 
rownames(meta) <- meta$sample
meta <- meta[!(rownames(meta) %in% "MOCK"),]
meta <- meta[!(rownames(meta) %in% "TL1"),]
meta <- meta[!(rownames(meta) %in% "TL2"),]
meta <- meta[!(rownames(meta) %in% "TL3"),]
meta <- meta[!(rownames(meta) %in% "TL4"),]
meta <- meta[!(rownames(meta) %in% "TL5"),]
meta <- meta[!(rownames(meta) %in% "TL6"),]
meta <- meta[!(meta$source %in% "b"),]
meta <- meta[!(meta$source %in% "s"),]
```

```{r soil}
soil <- read.csv("../data/Design_and_soil.csv",  header=T)

soil <- soil %>%
  filter(time!="field")%>%
  filter(time!="start")

by_time_plant_treatment <- soil %>%
  column_to_rownames("chamber")%>%
  dplyr::select(-history, -time) %>%
  group_by(plant, treatment) %>%
  dplyr::summarise_all(list(mean,sd, min, max), na.rm=T)

#write.csv(x = by_time_plant_treatment, file = "../figures/pub/soil_chem.csv")
```

# PERMANOVA

```{r PERMANOVA with just end samples}
#transforms read counts to relative abundance
otu.2.rel <- otu.2
for(i in 1:dim(otu.2)[1]){
  otu.2.rel[i,] <- otu.2[i,]/sum(otu.2[i,])
}

rownames(meta) <- rownames(otu.2.rel)
all.equal(rownames(otu.2.rel), rownames(meta))
metaotu <- cbind(meta, otu.2.rel)

# PERMANOVA:
metaotu.ad <- adonis2(metaotu[,-c(1:9)] ~ plant * history + plant * treatment, method = "bray", data = metaotu, perm=10000, set.seed=42)
metaotu.ad 
```
# PCoA

```{r Calculate distance matrix and PCoA components without field, end samples only}
otu.2.rel.dist<-vegdist(otu.2.rel, method="bray")

# Principal Coordinates Analysis
otu.2.rel.dist.pcoa <- cmdscale(otu.2.rel.dist, k=3, eig=TRUE, add=FALSE)
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1a <- round(otu.2.rel.dist.pcoa$eig[1] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
explainvar2a <- round(otu.2.rel.dist.pcoa$eig[2] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
sum.eiga <- sum(explainvar1a, explainvar2a)

explainvar1a #25% (rounding)
explainvar2a #11% (rounding)
```

```{r plot PCoA without field, end samples only}
all.equal(rownames(meta), rownames(otu.2.rel))
# Set treatments
treat1 <- as.factor(meta$history)
droplevels(treat1)
treat2 <- as.factor(meta$treatment)
droplevels(treat2)
#levels(treat2) <- c("I","W","D")
treat3 <- as.factor(meta$plant)
droplevels(treat3)

pcoa.groups <- paste(meta$history, meta$treatment, meta$plant, sep = "_")

pcoa.points <- data.frame(otu.2.rel.dist.pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)

pcoa.plant <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 3))  #plant
pcoa.trt <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 2)) #treatment
pcoa.history <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 1))  #history 

pcoa.trt <- ordered(pcoa.trt, levels=c("D", "I", "W"))
pcoa.trt <-revalue(pcoa.trt, c("D"="Dry", "I"="Interim","W"="Wet"))

pcoa.history <- ordered(pcoa.history, levels=c("D", "I", "W"))
pcoa.history <-revalue(pcoa.history, c("D"="Dry", "I"="Interim","W"="Wet"))

pcoa.plant <- ordered(pcoa.plant, levels=c("n", "p"))
pcoa.plant <-revalue(pcoa.plant, c("n"="No Plant","p"="Plant"))

soil.finish <- soil %>% 
dplyr::select(chamber, time, treatment, moisture.percent, pH, nh4.mg.l, no2.no3.mg.l, c.percent, n.percent, P.ppm, K.ppm, Mg.ppm, S.ppm, Fe.ppm, Mn.ppm, HM, CEC)

c<-cor(soil.finish[, -c(1:3,7)]) #design fields and no3 because many missing values
c<-cor(soil.finish[, -c(1:3,7,9,12,14,17)]) # %N Mg Fe CEC

# envfit:
meta.env <- soil.finish[, -c(1:3,7,9,12,14,17)]
fit<-envfit(otu.2.rel.dist.pcoa$points, meta.env, perm=1000,na.rm=T, set.seed(42))
A <-as.list(fit$vectors)
vec<-as.data.frame(fit$vectors$arrows*sqrt(fit$vectors$r)*0.25) #Scale vector by correlation
p<- as.data.frame(A$pvals)

vec <- cbind(vec, p) 
vec <-  subset(vec, A$pvals<=0.05)

vec$parm<-rownames(vec)
colnames(vec)
rownames(vec)
vec$parm<-c("pH", "%C", "Sulfur", "Manganese") # Check that these readable labels match the originals

# To use open and closed shapes, must create a combined category for History and Plant:
history_plant_combined <- c(paste(pcoa.history, pcoa.plant, sep = "_"))

#Make new PCoA data frame, then append experimental factors (plant, treatment, history) as new columns:
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$history <- as.factor(pcoa.history)
pcoa.cent.dataframe.trts$trt <- as.factor(pcoa.trt)
pcoa.cent.dataframe.trts$plant <- as.factor(pcoa.plant)
pcoa.cent.dataframe.trts$history_plant <- as.factor(history_plant_combined)

#Plotting:
plot16S_PCoA <- ggplot(pcoa.cent.dataframe.trts, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +   
  geom_point(aes(shape = history_plant, color=pcoa.trt), size=5, stroke=2) +  
  theme_bw() +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  scale_shape_manual(labels = c("Dry_No Plant", "Dry_Plant", "Interim_No Plant", "Interim_Plant", "Wet_No Plant", "Wet_Plant"), values = c(0, 15, 2, 17, 1, 19), name="History_Plant") +
  xlab("PCoA 1 (25%)") + ylab("PCoA 2 (11%)") + 
  geom_segment(data=vec, aes(x=0,xend=Dim1,y=0,yend=Dim2), size=1, arrow = arrow(length = unit(0.2, "cm")),colour="black",inherit.aes=F)+ 
  #Label soil vectors 
  geom_text(data=vec,aes(x=Dim1-0.014,y=Dim2+0.011,label=parm),size=5, inherit.aes=F)+
  ggtitle("") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=18), axis.text=element_text(size=18), 
          axis.text.x = element_text(size=18),  axis.text.y = element_text(size=18),
          panel.border = element_rect(colour = "black", size=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=18)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(fill = guide_legend(override.aes=list(shape=21)))

plot16S_PCoA

#ggsave("../figures/pub/16S_PCoA.tiff", plot=plot16S_PCoA, device="tiff", path=NULL, scale=1, width=9.5, height=5, dpi=900, limitsize=TRUE)

#ggsave("../figures/pub/16S_PCoA.png", plot=plot16S_PCoA, device="png", path=NULL, scale=1, width=9.5, height=5, dpi=300, limitsize=TRUE)

#ggsave("../figures/pub/16SrRNA_timberlake_end_vectors.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=9.5, height=5, dpi=900, limitsize=TRUE)
```
# Indicator species

```{r read and clean taxonomy file}
tax <- import_mothur(mothur_constaxonomy_file =  taxfile)
tax.df <- as.data.frame(tax)

tax.df <- tax.df %>% 
  rownames_to_column(var = "otu")

colnames(tax.df)=c("otu","domain","Phylum","Class", "Order","Family","Genus")

#Clean up taxonomy file. Remove underscores and unclassified
tax.df <- tax.df %>%
    mutate(Family = str_replace(Family, "_family_incertae_sedis", ""), Genus = str_replace(Genus, "_genera_incertae_sedis", "")) %>%
  mutate(Phylum = str_replace_all(Phylum,"_unclassified", ""), Order = str_replace_all(Order,"_unclassified", ""),Class = str_replace_all(Class,"_unclassified", ""), Family = str_replace_all(Family,"_unclassified", ""), Genus = str_replace_all(Genus,"_unclassified", "")) %>%
  mutate(Phylum = str_replace_all(Phylum,"_", " "), Order = str_replace_all(Order,"_", " "),Class = str_replace_all(Class,"_", " "), Family = str_replace_all(Family,"_", " "), Genus = str_replace_all(Genus,"_", " ")) 
```

```{r Indicator Species}
#Final sampling, no field
new.data <-metaotu
new.data$history <- as.factor(new.data$history)
new.data$treatment <- as.factor(new.data$treatment)

#group = interaction(new.data$plant, new.data$treatment, new.data$history) #Setup to analyze all three factors
# Only History and Treatment had significant results in PERMANOVA (i.e., adonis) (see above)
# Therefore, only History and Treatment will be used to group bacterial indicator species analysis:
group.hist <- new.data$history
group.trt <- new.data$treatment

dataREL <- new.data[,-c(1:9)] # Only OTU columns (i.e., relative [REL] abundance data), no metadata columns

dataREL <- dataREL[, colSums(dataREL) > 0.025] # Drop very rare OTUs from the analysis; only analyze most abundant 2.5%

bac.ind_hist <- indval(dataREL, group.hist, numitr = 10000, set.seed(42))
bac.ind_trt <- indval(dataREL, group.trt, numitr = 10000, set.seed(42))

summary(bac.ind_hist)
summary(bac.ind_trt)

# p-value cutoff:
inds_hist <- which(bac.ind_hist$pval <= 0.01)
inds_trt <- which(bac.ind_trt$pval <= 0.01)

# Make data frame skeleton:
bac.indicators_hist <- as.data.frame(matrix(NA, nrow = length(inds_hist), ncol = 4))
bac.indicators_trt <- as.data.frame(matrix(NA, nrow = length(inds_trt), ncol = 4))

# Set column names:
colnames(bac.indicators_hist) <- c("OTU", "Cluster", "IndVal", "Prob")
colnames(bac.indicators_trt) <- c("OTU", "Cluster", "IndVal", "Prob")

# Poplulate OTU column:
bac.indicators_hist$OTU <- names(inds_hist)
bac.indicators_trt$OTU <- names(inds_trt)

# Populate Cluster column:
bac.indicators_hist$Cluster <- bac.ind_hist$maxcls[inds_hist]
bac.indicators_trt$Cluster <- bac.ind_trt$maxcls[inds_trt]

# Populate IndVal column:
bac.indicators_hist$IndVal <- bac.ind_hist$indcls[inds_hist]
bac.indicators_trt$IndVal <- bac.ind_trt$indcls[inds_trt]

# Populate p-value column:
bac.indicators_hist$Prob <- bac.ind_hist$pval[inds_hist]
bac.indicators_trt$Prob <- bac.ind_trt$pval[inds_trt]

otu.tax <- as.data.frame(tax.df)

ind.tax_hist <- otu.tax[which(as.character(otu.tax$otu) %in% bac.indicators_hist$OTU), ]
ind.tax_trt <- otu.tax[which(as.character(otu.tax$otu) %in% bac.indicators_trt$OTU), ]

ind.tax_hist <- ind.tax_hist[match(ind.tax_hist$otu, bac.indicators_hist$OTU), ]
ind.tax_trt <- ind.tax_trt[match(ind.tax_trt$otu, bac.indicators_trt$OTU), ]

# Add taxonomic information:
indicator.bac_hist <- cbind(bac.indicators_hist, ind.tax_hist[, -c(1)])
indicator.bac_trt <- cbind(bac.indicators_trt, ind.tax_trt[, -c(1)])

# Order rows by Cluster:
indicator.bac_hist <- indicator.bac_hist[order(as.numeric(indicator.bac_hist$Cluster)), ]
indicator.bac_trt <- indicator.bac_trt[order(as.numeric(indicator.bac_trt$Cluster)), ]

# Print information from tables:
table(indicator.bac_hist$Cluster)
table(indicator.bac_trt$Cluster)

table(indicator.bac_hist$Phylum)
table(indicator.bac_trt$Phylum)

# Identify what the Cluster numbers mean:
levels(group.hist)
levels(group.trt)

#Replace cluster numbers with factor names:
indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 1] <- "Dry"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 1] <- "Dry"

indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 2] <- "Interim"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 2] <- "Interim"

indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 3] <- "Wet"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 3] <- "Wet"

#Find overlapping indicator tax in hist and trt:
common_indicators <- generics::intersect(indicator.bac_hist$OTU, indicator.bac_trt$OTU)
common_indicators <- as.data.frame(common_indicators) #Make it a data frame
colnames(common_indicators) <- "OTU" #Rename column header to "OTU" to match with indicator.bac_hist and indicator.bac_trt

common_indicators <- left_join(common_indicators, indicator.bac_hist, by = "OTU") #Add data from indicator.bac_hist
colnames(common_indicators)[2:4] <- c("Cluster_hist", "IndVal_hist", "Prob_hist") #Rename to not confuse with treatment indicators, added next:
  #Add data from indicator.bac_trt:
common_indicators <- left_join(common_indicators, indicator.bac_trt, by = c("OTU", "domain", "Phylum", "Class", "Order", "Family", "Genus")) 
colnames(common_indicators)[11:13] <- c("Cluster_treatment", "IndVal_treatment", "Prob_treatment")

# Export Bacteria Indicator Tables:
#write.table(indicator.bac_hist, "../figures/pub/BacterialIndicators_History.txt",
           #sep="\t", row.names = F, quote = F)

#write.table(indicator.bac_trt, "../figures/pub/BacterialIndicators_Treatment.txt",
            #sep="\t", row.names = F, quote = F)
```

# Soil parameters

```{r soil parameters}
soil <- read.csv("../data/Design_and_soil.csv",  header=T)

# Remove field samples:
soil_nofield <- soil %>%
  filter(soil$time!="field")

# For axes titles, change abbreviations (D, I, W) to full words (Dry, Interim, Wet):
soil_nofield$treatment <-revalue(soil_nofield$treatment, c("D" = "Dry", "I" = "Interim","W" = "Wet"))

# For axes titles, change abbreviations (N, P) to full words (No Plant, Plant):
soil_nofield$plant <- revalue(soil_nofield$plant, c("N" = "No Plant", "P" = "Plant"))

# Make soil_nofield$time uppercase, first letter only:
soil_nofield$time <- str_to_title(soil_nofield$time)

write.csv(soil_nofield, file="../figures/soil.csv")

#Subset Start and Finish for plotting:
soil_nofield.Start <- subset(soil_nofield, time == "Start")
soil_nofield.Finish <- subset(soil_nofield, time == "Finish")

#Plotting:
#Start:
iris_tubes.Start <- ggplot(soil_nofield.Start, aes(x=treatment, y=(redox.percent.paint.removed*100))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(~plant) +
  geom_point(aes(color=factor(treatment)), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  theme_bw() +
  #Remove plot grid lines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = "IRIS Percent", title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

iris_tubes.Start

#Finish:
iris_tubes.Finish <- ggplot(soil_nofield.Finish, aes(x=treatment, y=(redox.percent.paint.removed*100))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(~plant) +
  geom_point(aes(color=factor(treatment)), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  theme_bw() +
  #Remove plot grid lines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "Treatment", y = "IRIS Percent", title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

iris_tubes.Finish

#Panel together using ggarrange{ggpubr}:
iris_tubes.Combined <- ggarrange(iris_tubes.Start, iris_tubes.Finish, labels = c("Start (Week 0)", "Finish (Week 8)"), ncol = 1, nrow = 2, legend = "right", common.legend = T)

iris_tubes.Combined

#ggsave("../figures/pub/iris_tubes.tiff", plot=iris_tubes.Combined, device="tiff", path=NULL, scale=1, width=6, height=5, dpi=900, limitsize=TRUE, bg="white")

#ggsave("../figures/pub/iris_tubes.jpg", plot=iris_tubes.Combined, device=NULL, path=NULL, scale=1, width=6, height=5, dpi=900, limitsize=TRUE ,bg="transparent")

#ggsave("../figures/pub/iris_tubes.pdf", plot=iris_tubes.Combined, device=NULL, path=NULL, scale=1, width=6, height=5, dpi=900, limitsize=TRUE ,bg="transparent")
```

```{r soil moisture boxplot}
soil_moisture <- ggplot(soil_nofield, aes(x=treatment, y=moisture.percent, fill=plant)) + 
  geom_boxplot() +
  facet_grid(~ fct_rev(time)) +
 scale_fill_manual(values=c("gray", "darkgreen"), labels=c("No Plant", "Plant")) +
#Set plot elements
theme_bw()  +
#Remove plot gridlines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right") + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = "Moisture Percent", title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

soil_moisture

#ggsave("../figure/pub/soil_moisture.tiff", plot=soil_moisture, device="tiff", path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE, bg="transparent")

#ggsave("../figures/pub/soil_moisture.jpg", plot=soil_moisture, device=NULL, path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE , bg="transparent")

#ggsave("../figures/pub/soil_moisture.pdf", plot=soil_moisture, device=NULL, path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE , bg="transparent")
```

```{r soil plots}
ammonium <- ggplot(soil_nofield, aes(x=treatment, y=nh4.mg.l)) + 
  geom_boxplot(aes(x=treatment, y=nh4.mg.l, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment")+
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Ammonium (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

ammonium

nitrate <- ggplot(soil_nofield, aes(x=treatment, y=no2.no3.mg.l)) + 
  geom_boxplot(aes(x=treatment, y=no2.no3.mg.l, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment")+
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Nitrate (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

nitrate

soil_Fe <- ggplot(soil_nofield, aes(x=treatment, y=Fe.ppm)) + 
  geom_boxplot(aes(x=treatment, y=Fe.ppm, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment", labels=c("Dry", "Interim", "Wet"))+
  scale_x_discrete(labels=c("Dry", "Interim", "Wet"), name="") +
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Iron (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

soil_Fe
```

# Greenhouse gases (GHGs)

```{r GHG T1 concentration analysis}
#read in gas data
ghg <- read.csv("../data/TMG_GHG-Flux_T1_Pub_2-2020.csv", header=TRUE)

# Version of ghg with labels for T1 through T4 analysis:
ghg2 <- read.csv("../data/TMG_GHG-Flux_Pub_3.csv", header=TRUE)

# FIX c2o TYPO IN ghg2, then remove line 600
colnames(ghg2)[18] <- "co2_ppm"

ghg <- ghg %>%
  dplyr::select('date', 'chamber', 'timepoint', 'plant', 'mg.CH4.N.m.2', 'mg.CO2.N.m.2', 'mg.N2O.N.m.2')%>%
  dplyr::filter(timepoint=='T1' & date !='7/24/2016' & date != '6/13/2016')

# ghg2, select T1 through T4, and all ppm measurements (typo: c2o should be co2)
ghg_ppm <- ghg2 %>%
  dplyr::select('date', 'chamber', 'timepoint', 'plant', 'ch4_ppm', 'co2_ppm', 'n2o_ppm')%>%
  dplyr::filter(date !='7/24/2016' & date != '6/13/2016')

# Set factors:
ghg$chamber <- as.factor(ghg$chamber)
ghg$date <- as.factor(ghg$date)

ghg_ppm$chamber <- as.factor(ghg_ppm$chamber)
ghg_ppm$date <- as.factor(ghg_ppm$date)

# Change design$box to design$chamber in order to join with ghg$chamber
colnames(meta)[2] <- "chamber"

ghg <- ghg %>%
  left_join(meta[,-c(3, 4, 7, 8, 9)], by="chamber")

ghg_ppm <- ghg_ppm %>%
  left_join(meta[,-c(3, 4, 7, 8, 9)], by="chamber")

# Add new date column, to shorten:
ghg$date_short <- ghg$date
ghg_ppm$date_short <- ghg_ppm$date
# remove "/2016" from date_short:
ghg$date_short <- str_remove(ghg$date_short, "/2016")
ghg_ppm$date_short <- str_remove(ghg_ppm$date_short, "/2016")

# Separate data frames for "No Plant" and "Plant", for plotting:
ghg.noplant <- ghg %>%
  filter(plant=="N")

ghg.plant <- ghg %>%
  filter(plant=="P")

ghg_ppm.noplant <- ghg_ppm %>%
  filter(plant=="N")

ghg_ppm.plant <- ghg_ppm %>%
  filter(plant=="P")
```

```{r add T1-T4 averages to ghg}
# Should be only line needed now. Delete rest of this chunk after confirming rest of code works.
ghg_ppm_filtered <- ghg_ppm %>%
  dplyr::filter(timepoint == "T1")

#install.packages("slider")
#library(slider)

# Average CH4 ppm:
#ghg_ppm$ch4_ppm_avg <- slide(
 # ghg_ppm$ch4_ppm,
#  ~mean(.x, na.rm = T),
#  .after = 3,
#  .step = 4,
#  .complete = F
#)

# Average CO2 ppm:
#ghg_ppm$co2_ppm_avg <- slide(
#  ghg_ppm$co2_ppm,
#  ~mean(.x, na.rm = T),
#  .after = 3,
#  .step = 4,
#  .complete = F
#)

# Average N2O ppm:
#ghg_ppm$n2o_ppm_avg <- slide(
 # ghg_ppm$n2o_ppm,
  #~mean(.x, na.rm = T),
  #.after = 3,
  #.step = 4,
  #.complete = F
#)

# Convert one of the averaged column to character, then use it to filter:
#ghg_ppm$ch4_ppm_avg <- as.character(ghg_ppm$ch4_ppm_avg)

#ghg_ppm_filtered <- ghg_ppm %>%
 # dplyr::filter(ch4_ppm_avg != "NULL")



# After filtering, convert all averaged columns to numeric:
#ghg_ppm_filtered$ch4_ppm_avg <- as.numeric(ghg_ppm_filtered$ch4_ppm_avg)
#ghg_ppm_filtered$co2_ppm_avg <- as.numeric(ghg_ppm_filtered$co2_ppm_avg)
#ghg_ppm_filtered$n2o_ppm_avg <- as.numeric(ghg_ppm_filtered$n2o_ppm_avg)
```

```{r GHG flux wrangling}
# Format a dataframe for 'gasfluxes' package:
ghg_flux <- ghg2[,c(1:2, 12, 15:19)]

# Remove 7/24/2016 and 6/13/2016 from data frame, as per Gina's advice
ghg_flux <- ghg_flux %>%
  dplyr::filter(date !='7/24/2016' & date != '6/13/2016')

# Combine 'date' and 'chamber' strings to create a unique ID:
ghg_flux$ID <- paste(ghg_flux$date, ghg_flux$chamber)

# convert Volume (cm^3) to Volume (m^3):
ghg_flux$V <- (ghg_flux$Vol..cm3. * (1/1e+06))

# Copy 'Area..m2.' to a new column named 'A':
ghg_flux$A <- ghg_flux$Area..m2.

# Convert time in minutes to time in hours:
ghg_flux$time <- (ghg_flux$time_min * (1/60))

# Convert ppm to mg/m^3, based on https://www.cdc.gov/niosh/docs/2004-101/calc.html (based on 25°C and 1 atm):
# y mg/m3 = (x ppm)(molar mass)/24.45
  # 24.45 is the volume (liters) of a mole (gram molecular weight) of a gas or vapor when the pressure is at 1 atmosphere (760 torr or 760 mm Hg) and at 25°C

# CH4 mg/m3
ghg_flux$C_ch4 <- ((ghg_flux$ch4_ppm * 16.04)/24.45) # C is the package 'gasfluxes' abbreviation for concentration. 16.04 g/mol = molar mass of CH4

# CO2 mg/m3
ghg_flux$C_co2 <- ((ghg_flux$co2_ppm * 44.01)/24.45) # molar mass of CO2 = 44.01 g/mol

# N2O mg/m3
ghg_flux$C_n2o <-  ((ghg_flux$n2o_ppm * 44.013)/24.45) # molar mass of N2O = 44.013 g/mol

# Split off data frames for each greenhouse gas:
ch4_flux_input <- ghg_flux[,9:13]

co2_flux_input <- ghg_flux[,c(9:12, 14)]

n2o_flux_input <- ghg_flux[,c(9:12, 15)]

# Rename C_ghg columns to 'C':
colnames(ch4_flux_input)[5] <- "C"
colnames(co2_flux_input)[5] <- "C"
colnames(n2o_flux_input)[5] <- "C"

# Must omit NAs before using gasfluxes():
ch4_flux_input <- na.omit(ch4_flux_input)
co2_flux_input <- na.omit(co2_flux_input)

# n2o_flux_input has negative values, will need to be converted to NAs, then all NAs omitted
n2o_flux_input$C[n2o_flux_input$C<0] <- NA
n2o_flux_input <- na.omit(n2o_flux_input)
```

```{r GHG flux calculation}
#install.packages("gasfluxes")
require("gasfluxes")

ch4_flux.results <- gasfluxes(ch4_flux_input, method = c("linear","robust linear", "HMR"), plot = T)

co2_flux.results <- gasfluxes(co2_flux_input, method = c("linear","robust linear", "HMR"), plot = T)

n2o_flux.results <- gasfluxes(n2o_flux_input, method = c("linear","robust linear", "HMR"), plot = T)
```

```{r GHG flux results wrangling}
# If HMR is available, use HMR
# If HMR is not available, use robust linear
# If robust linear is not available use linear
# New DF with all three f0s:
ch4_fluxes_only <- ch4_flux.results[,c(1:2, 11, 17)]

co2_fluxes_only <- co2_flux.results[,c(1:2, 11, 17)]

n2o_fluxes_only <- n2o_flux.results[,c(1:2, 11, 17)]

# New column for final flux selection:
ch4_fluxes_only$f0_selected <- ch4_fluxes_only$HMR.f0

co2_fluxes_only$f0_selected <- co2_fluxes_only$HMR.f0

n2o_fluxes_only$f0_selected <- n2o_flux.results$HMR.f0

# Where HMR = NA, replace with the robust linear value:
ch4_fluxes_only$f0_selected[na.action(na.omit(ch4_fluxes_only$HMR.f0))] <- ch4_fluxes_only$robust.linear.f0[na.action(na.omit(ch4_fluxes_only$HMR.f0))]

co2_fluxes_only$f0_selected[na.action(na.omit(co2_fluxes_only$HMR.f0))] <- co2_fluxes_only$robust.linear.f0[na.action(na.omit(co2_fluxes_only$HMR.f0))]

n2o_fluxes_only$f0_selected[na.action(na.omit(n2o_fluxes_only$HMR.f0))] <- n2o_fluxes_only$robust.linear.f0[na.action(na.omit(n2o_fluxes_only$HMR.f0))]

# Where f0_selected still has NAs, replace with the linear.f0:
ch4_fluxes_only$f0_selected[na.action(na.omit(ch4_fluxes_only$f0_selected))] <- ch4_fluxes_only$linear.f0[na.action(na.omit(ch4_fluxes_only$f0_selected))]

co2_fluxes_only$f0_selected[na.action(na.omit(co2_fluxes_only$f0_selected))] <- co2_fluxes_only$linear.f0[na.action(na.omit(co2_fluxes_only$f0_selected))]

n2o_fluxes_only$f0_selected[na.action(na.omit(n2o_fluxes_only$f0_selected))] <- n2o_fluxes_only$linear.f0[na.action(na.omit(n2o_fluxes_only$f0_selected))]

# Combine separate DFs back into one DF for statistics
select_fluxes_combined <- ch4_fluxes_only[,c(1,5)] # start with CH4 IDs and f0 for a base dataframe

colnames(select_fluxes_combined)[2] <- "f0_CH4" # rename to distinguish between GHGs

select_fluxes_combined$f0_CO2 <- co2_fluxes_only$f0_selected # add on CO2 fluxes

select_fluxes_combined$f0_N2O <- n2o_fluxes_only$f0_selected # add on N2O fluxes

# Add date_short back onto DFs for plotting
select_fluxes_combined$date_short <- ghg_ppm_filtered$date_short

# Add on 'plant'
select_fluxes_combined$plant <- ghg_ppm_filtered$plant

# Add on 'treatment'
select_fluxes_combined$treatment <- ghg_ppm_filtered$treatment

# Add on 'history'
select_fluxes_combined$history <- ghg_ppm_filtered$history
```


```{r ANOVAS prep}
# https://www.datanovia.com/en/lessons/repeated-measures-anova-in-r/

library(rstatix)

# Make sure factors are factors: 

  # ghg_ppm_filtered:
ghg_ppm_filtered$chamber <- as.factor(ghg_ppm_filtered$chamber)
ghg_ppm_filtered$date_short <- as.factor(ghg_ppm_filtered$date_short)
ghg_ppm_filtered$plant <- as.factor(ghg_ppm_filtered$plant)
ghg_ppm_filtered$treatment <- as.factor(ghg_ppm_filtered$treatment)

  # ghg
ghg$chamber <- as.factor(ghg$chamber)
ghg$date_short <- as.factor(ghg$date_short)
ghg$plant <- as.factor(ghg$plant)
ghg$treatment <- as.factor(ghg$treatment)

  # select_fluxes_combined
    # Add on chamber to be consistent with ghg_ppm_filtered and ghg:
select_fluxes_combined$chamber <- ghg$chamber

    # setting other factors, as above
select_fluxes_combined$date_short <- as.factor(select_fluxes_combined$date_short)
select_fluxes_combined$plant <- as.factor(select_fluxes_combined$plant)
select_fluxes_combined$treatment <- as.factor(select_fluxes_combined$treatment)
```

```{r ANOVA assumption testing: outliers}
# CH4 PPM
Outliers_ch4_ppm <- ghg_ppm_filtered %>%
  group_by(date_short) %>%
  identify_outliers(ch4_ppm)
  # There are extreme outliers. Could run with and without outliers.

# Filter extreme outliers only:
Extreme_Outliers_ch4_ppm <- Outliers_ch4_ppm %>%
  dplyr::filter(is.extreme == TRUE)

# Create unique key for filtering:
Extreme_Outliers_ch4_ppm$chamber_date <-paste(Extreme_Outliers_ch4_ppm$chamber, Extreme_Outliers_ch4_ppm$date_short)

# New CH4 dataframe with extreme outliers dropped:
ch4_ppm_no_outliers <- ghg_ppm_filtered
ch4_ppm_no_outliers$chamber_date <- paste(ch4_ppm_no_outliers$chamber, ch4_ppm_no_outliers$date_short) 

# If values appear in Extreme_Outliers, drop them from 'no_outliers' dataframe:
ch4_ppm_no_outliers <- ch4_ppm_no_outliers[!(ch4_ppm_no_outliers$chamber_date %in% Extreme_Outliers_ch4_ppm$chamber_date),]

# Test for extreme outliers again:
ch4_ppm_no_outliers %>%
  group_by(date_short) %>%
  identify_outliers(ch4_ppm)
  # still extreme outliers, but different ones now
```
  
```{r log transform}  
  # log transform
ghg_ppm_filtered$log10_CH4 <- log(ghg_ppm_filtered$ch4_ppm, 10)

ghg_ppm_filtered %>%
  group_by(date_short) %>%
  identify_outliers(log10_CH4)
  # Reduces number of extreme outliers by 2, but still not great



# CO2 PPM
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  identify_outliers(co2_ppm)
  # There are extreme outliers. Could run with and without outliers.

# N2O PPM
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  identify_outliers(n2o_ppm)
  # There are extreme outliers. Could run with and without outliers.
```

```{r ANOVA assumption testing: normality}
# CH4 PPM
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  shapiro_test(ch4_ppm)
  # Not normally distributed

  # QQ plot:
ggqqplot(ghg_ppm_filtered, "ch4_ppm", facet.by = "date_short")


  # log transform:
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  shapiro_test(log10_CH4)
  # Not normally distributed

ggqqplot(ghg_ppm_filtered, "log10_CH4", facet.by = "date_short")
  # seems to be an improvement, but not good enough

  # outliers dropped:
ch4_ppm_no_outliers %>%
  group_by(date_short) %>%
  shapiro_test(ch4_ppm)
  # Not normally distributed

ggqqplot(ch4_ppm_no_outliers, "ch4_ppm", facet.by = "date_short")
  # seems to be an improvement, but not good enough

# CO2 PPM
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  shapiro_test(co2_ppm)
  # Not normally distributed

  # QQ plot:
ggqqplot(ghg_ppm_filtered, "co2_ppm", facet.by = "date_short")

# N2O PPM
ghg_ppm_filtered %>%
  group_by(date_short) %>%
  shapiro_test(n2o_ppm)
  # Not normally distributed

  # QQ plot:
ggqqplot(ghg_ppm_filtered, "n2o_ppm", facet.by = "date_short")
```



```{r one-way repeated measures ANOVA}
# CH4:
res.aov.CH4 <- anova_test(data = ghg_ppm_filtered, dv = ch4_ppm, wid = chamber, within = date_short)
get_anova_table(res.aov.CH4)

# CO2
res.aov.CO2 <- anova_test(data = ghg_ppm_filtered, dv = co2_ppm, wid = chamber, within = date_short)
get_anova_table(res.aov.CO2) # Sig. different within time points

# Pair-wise comparisons (PWC):
pwc_CO2 <- ghg_ppm_filtered %>%
  pairwise_t_test(
    co2_ppm ~ date_short, paired = TRUE,
    p.adjust.method = "fdr"
    )
pwc_CO2

# N2O:
res.aov.N2O <- anova_test(data = ghg_ppm_filtered, dv = n2o_ppm, wid = chamber, within = date_short)
get_anova_table(res.aov.N2O)
```

```{r Mixed ANOVA}
# CH4: 
    # PPM with date:
MixANOVA_CH4_ppm <- anova_test(
  data = ghg_ppm_filtered, dv = ch4_ppm, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_CH4_ppm)

  # Flux with date
MixANOVA_CH4_Flux <- anova_test(
  data = select_fluxes_combined, dv = f0_CH4, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_CH4_Flux)

# CO2:
    # PPM with date:
MixANOVA_CO2_ppm <- anova_test(
  data = ghg_ppm_filtered, dv = co2_ppm, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_CO2_ppm)

  # Flux with date
MixANOVA_CO2_Flux <- anova_test(
  data = select_fluxes_combined, dv = f0_CO2, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_CO2_Flux)

# N2O:
    # PPM with date:
MixANOVA_N2O_ppm <- anova_test(
  data = ghg_ppm_filtered, dv = n2o_ppm, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_N2O_ppm)

  # Flux with date
MixANOVA_N2O_Flux <- anova_test(
  data = select_fluxes_combined, dv = f0_N2O, wid = chamber,
  between = c(plant, treatment),
  within = date_short)
get_anova_table(MixANOVA_N2O_Flux)
```

```{r Mixed ANOVA, limit factors}
# CH4: 
    # PPM with date, TREATMENT:
MixANOVA_CH4_ppm_treatment <- anova_test(
  data = ghg_ppm_filtered, dv = ch4_ppm, wid = chamber,
  between = treatment,
  within = date_short)
get_anova_table(MixANOVA_CH4_ppm_treatment)

# CO2:
    # PPM with date, TREATMENT:
MixANOVA_CO2_ppm_treatment <- anova_test(
  data = ghg_ppm_filtered, dv = co2_ppm, wid = chamber,
  between = treatment,
  within = date_short)
get_anova_table(MixANOVA_CO2_ppm_treatment)
  # Do a Tukey on treatment:
    # Tukey on treatment
tukey_AOV_CH4_treatment <- tukey_hsd(ghg_ppm_filtered, co2_ppm ~ treatment)
tukey_AOV_CH4_treatment

# N2O:
    # PPM with date, TREATMENT:
MixANOVA_N2O_ppm_treatment <- anova_test(
  data = ghg_ppm_filtered, dv = n2o_ppm, wid = chamber,
  between = treatment,
  within = date_short)
get_anova_table(MixANOVA_N2O_ppm_treatment)
```

```{r Wilcoxon Rank Sum for Plant No Plant}
# CH4
wilcox_CH4_date <- wilcox.test(data = ghg_ppm_filtered, ch4_ppm ~ plant, paired = T, subset = date_short)
wilcox_CH4_date

wilcox_CH4_6.28 <- wilcox.test(data = ghg_ppm_filtered, ch4_ppm ~ plant, paired = T, subset = date_short == "6/28")
wilcox_CH4_6.28

wilcox_CH4_7.11 <- wilcox.test(data = ghg_ppm_filtered, ch4_ppm ~ plant, paired = T, subset = date_short == "7/11")
wilcox_CH4_7.11

wilcox_CH4_7.25 <- wilcox.test(data = ghg_ppm_filtered, ch4_ppm ~ plant, paired = T, subset = date_short == "7/25")
wilcox_CH4_7.25

wilcox_CH4_8.11 <- wilcox.test(data = ghg_ppm_filtered, ch4_ppm ~ plant, paired = T, subset = date_short == "8/11")
wilcox_CH4_8.11

# CO2
wilcox_CO2_date <- wilcox.test(data = ghg_ppm_filtered, co2_ppm ~ plant, paired = T, subset = date_short)
wilcox_CO2_date

wilcox_CO2_6.28 <- wilcox.test(data = ghg_ppm_filtered, co2_ppm ~ plant, paired = T, subset = date_short == "6/28")
wilcox_CO2_6.28

wilcox_CO2_7.11 <- wilcox.test(data = ghg_ppm_filtered, co2_ppm ~ plant, paired = T, subset = date_short == "7/11")
wilcox_CO2_7.11

wilcox_CO2_7.25 <- wilcox.test(data = ghg_ppm_filtered, co2_ppm ~ plant, paired = T, subset = date_short == "7/25")
wilcox_CO2_7.25

wilcox_CO2_8.11 <- wilcox.test(data = ghg_ppm_filtered, co2_ppm ~ plant, paired = T, subset = date_short == "8/11")
wilcox_CO2_8.11

# N2O
wilcox_N2O_date <- wilcox.test(data = ghg_ppm_filtered, n2o_ppm ~ plant, paired = T, subset = date_short)
wilcox_N2O_date

wilcox_N2O_6.28 <- wilcox.test(data = ghg_ppm_filtered, n2o_ppm ~ plant, paired = T, subset = date_short == "6/28")
wilcox_N2O_6.28

wilcox_N2O_7.11 <- wilcox.test(data = ghg_ppm_filtered, n2o_ppm ~ plant, paired = T, subset = date_short == "7/11")
wilcox_N2O_7.11

wilcox_N2O_7.25 <- wilcox.test(data = ghg_ppm_filtered, n2o_ppm ~ plant, paired = T, subset = date_short == "7/25")
wilcox_N2O_7.25

wilcox_N2O_8.11 <- wilcox.test(data = ghg_ppm_filtered, n2o_ppm ~ plant, paired = T, subset = date_short == "8/11")
wilcox_N2O_8.11
```

```{r Kruskal-Wallis Rank Sum Test on treatment}
# Will use DWASS-Steel-Critchlow-Fligner test for multiple comparisons, in package NSM3
#install.packages("NSM3")

require("NSM3")


# CH4
kruskal_CH4_6.28 <- kruskal.test(data = ghg_ppm_filtered, ch4_ppm ~ treatment, subset = date_short == "6/28")
kruskal_CH4_6.28

  # Pairwise comparisons:
    # Must subset 6/28:
ghg_ppm_filtered_6.28 <- ghg_ppm_filtered %>% filter(date_short == "6/28")

SDCFlig_CH4_6.28 <- pSDCFlig(x = ghg_ppm_filtered_6.28$ch4_ppm, g = ghg_ppm_filtered_6.28$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_6.28

kruskal_CH4_7.11 <- kruskal.test(data = ghg_ppm_filtered, ch4_ppm ~ treatment, subset = date_short == "7/11")
kruskal_CH4_7.11

  # Pairwise comparisons:
    # Must subset 7/11:
ghg_ppm_filtered_7.11 <- ghg_ppm_filtered %>% filter(date_short == "7/11")

SDCFlig_CH4_7.11 <- pSDCFlig(x = ghg_ppm_filtered_7.11$ch4_ppm, g = ghg_ppm_filtered_7.11$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_7.11

kruskal_CH4_7.25 <- kruskal.test(data = ghg_ppm_filtered, ch4_ppm ~ treatment, subset = date_short == "7/25")
kruskal_CH4_7.25

  # Pairwise comparisons:
    # Must subset 7/25:
ghg_ppm_filtered_7.25 <- ghg_ppm_filtered %>% filter(date_short == "7/25")

SDCFlig_CH4_7.25 <- pSDCFlig(x = ghg_ppm_filtered_7.25$ch4_ppm, g = ghg_ppm_filtered_7.25$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_7.25

kruskal_CH4_8.11 <- kruskal.test(data = ghg_ppm_filtered, ch4_ppm ~ treatment, subset = date_short == "8/11")
kruskal_CH4_8.11

  # Pairwise comparisons:
    # Must subset 8/11:
ghg_ppm_filtered_8.11 <- ghg_ppm_filtered %>% filter(date_short == "8/11")

SDCFlig_CH4_8.11 <- pSDCFlig(x = ghg_ppm_filtered_8.11$ch4_ppm, g = ghg_ppm_filtered_8.11$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_8.11

# CO2
kruskal_CO2_6.28 <- kruskal.test(data = ghg_ppm_filtered, co2_ppm ~ treatment, subset = date_short == "6/28")
kruskal_CO2_6.28

kruskal_CO2_7.11 <- kruskal.test(data = ghg_ppm_filtered, co2_ppm ~ treatment, subset = date_short == "7/11")
kruskal_CO2_7.11

  # Pairwise comparisons:
SDCFlig_CO2_7.11 <- pSDCFlig(x = ghg_ppm_filtered_7.11$co2_ppm, g = ghg_ppm_filtered_7.11$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CO2_7.11

kruskal_CO2_7.25 <- kruskal.test(data = ghg_ppm_filtered, co2_ppm ~ treatment, subset = date_short == "7/25")
kruskal_CO2_7.25

kruskal_CO2_8.11 <- kruskal.test(data = ghg_ppm_filtered, co2_ppm ~ treatment, subset = date_short == "8/11")
kruskal_CO2_8.11

  # Pairwise comparisons:
SDCFlig_CO2_8.11 <- pSDCFlig(x = ghg_ppm_filtered_8.11$co2_ppm, g = ghg_ppm_filtered_8.11$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CO2_8.11

# N2O
kruskal_N2O_6.28 <- kruskal.test(data = ghg_ppm_filtered, n2o_ppm ~ treatment, subset = date_short == "6/28")
kruskal_N2O_6.28

  # Pairwise comparisons:
SDCFlig_N2O_6.28 <- pSDCFlig(x = ghg_ppm_filtered_6.28$n2o_ppm, g = ghg_ppm_filtered_6.28$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_N2O_6.28

kruskal_N2O_7.11 <- kruskal.test(data = ghg_ppm_filtered, n2o_ppm ~ treatment, subset = date_short == "7/11")
kruskal_N2O_7.11

kruskal_N2O_7.25 <- kruskal.test(data = ghg_ppm_filtered, n2o_ppm ~ treatment, subset = date_short == "7/25")
kruskal_N2O_7.25

kruskal_N2O_8.11 <- kruskal.test(data = ghg_ppm_filtered, n2o_ppm ~ treatment, subset = date_short == "8/11")
kruskal_N2O_8.11
```


## GHG plots: T30 mg m-2 (Gina's method)

```{r CH4 plot}
plot.CH4.noplant <- ggplot(ghg.noplant, aes(x=date_short, y=mg.CH4.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4],"-C m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0,175))
 
plot.CH4.noplant


plot.CH4.plant <- ggplot(ghg.plant, aes(x=date_short, y=mg.CH4.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4],"-C m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 175))

 
plot.CH4.plant

#Panel together using ggarrange{ggpubr}:
#No legend version, for combining figures later:
CH4.Combined <- ggarrange(plot.CH4.noplant, plot.CH4.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CH4.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.CH4.plant.autoScale <- ggplot(ghg.plant, aes(x=date_short, y=mg.CH4.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.CH4.plant.autoScale

#ggsave("../figures/pub/Fig3/GinaMethod/CH4_Inset.tiff", plot=plot.CH4.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset. Two different versions for combined figure with or without legend:
#subvp_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#CH4.Combined
#print(plot.CH4.plant.autoScale, vp=subvp_NO_legend)
```

```{r CO2 plot}
plot.CO2.noplant <- ggplot(ghg.noplant, aes(x=date_short, y=mg.CO2.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2],"-C m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0,180))
 
plot.CO2.noplant


plot.CO2.plant <- ggplot(ghg.plant, aes(x=date_short, y=mg.CO2.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2],"-C m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 180))

 
plot.CO2.plant

#Panel together using ggarrange{ggpubr}:
#No legend version, for combining figures later:
CO2.Combined <- ggarrange(plot.CO2.noplant, plot.CO2.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CO2.Combined

```

```{r N2O Boxplot}
plot.N2O.noplant <- ggplot(ghg.noplant, aes(x=date_short, y=mg.N2O.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O-N m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 1.6))
 
plot.N2O.noplant


plot.N2O.plant <- ggplot(ghg.plant, aes(x=date_short, y=mg.N2O.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O-N m"^{-2})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 1.6))

plot.N2O.plant


#Panel together using ggarrange{ggpubr}:
# Version without legend for multi-panel figure:
N2O.Combined <- ggarrange(plot.N2O.noplant, plot.N2O.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

N2O.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.N2O.plant.autoScale <- ggplot(ghg.plant, aes(x=date_short, y=mg.N2O.N.m.2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.N2O.plant.autoScale

#ggsave("../figures/pub/Fig3/GinaMethod/N2O_Inset.tiff", plot=plot.N2O.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset:
#subvp2_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp2_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#N2O.Combined
#print(plot.N2O.plant.autoScale, vp=subvp2_NO_legend)
```

```{r Combined GHG Boxplot}
comb_GHG_BoxPlot <- ggarrange(CH4.Combined, CO2.Combined, N2O.Combined, labels = c("A", "B", "C"), ncol = 3, nrow = 1, legend = "none")

comb_GHG_BoxPlot

#ggsave("../figures/pub/Fig3/GinaMethod/BasePlots_ggpubr.tiff", plot=comb_GHG_BoxPlot, device="tiff", path=NULL, scale=1, width=15, height=6, dpi=600, limitsize=TRUE, bg="white")
```


## GHG plots: T30 ppm

```{r subset plant no plant for ppms}
ghg_ppm_filtered.noplant <- ghg_ppm_filtered %>%
  filter(plant=="N")

ghg_ppm_filtered.plant <- ghg_ppm_filtered %>%
  filter(plant=="P")
```

```{r CH4 T30 ppm plot}
plot.CH4.T30ppm.noplant <- ggplot(ghg_ppm_filtered.noplant, aes(x=date_short, y=ch4_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CH"[4]," (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 2000))
 
plot.CH4.T30ppm.noplant


plot.CH4.T30ppm.plant <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=ch4_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CH"[4]," (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 2000))

 
plot.CH4.T30ppm.plant

#Panel together using ggarrange{ggpubr}:
#No legend version, for combining figures later:
CH4.T30ppm.Combined <- ggarrange(plot.CH4.T30ppm.noplant, plot.CH4.T30ppm.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CH4.T30ppm.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.CH4.T30ppm.plant.autoScale <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=ch4_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.CH4.T30ppm.plant.autoScale

#ggsave("../figures/pub/Fig3/T30ppm/CH4_Inset_T30ppm.tiff", plot=plot.CH4.T30ppm.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset. Two different versions for combined figure with or without legend:
#subvp_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#CH4.Combined
#print(plot.CH4.plant.autoScale, vp=subvp_NO_legend)
```

```{r CO2 T30 ppm plot}
plot.CO2.T30ppm.noplant <- ggplot(ghg_ppm_filtered.noplant, aes(x=date_short, y=co2_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CO"[2]," (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 2200))
 
plot.CO2.T30ppm.noplant


plot.CO2.T30ppm.plant <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=co2_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CO"[2]," (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 2200))
 
plot.CO2.T30ppm.plant

#Panel together using ggarrange{ggpubr}:
#No legend version, for combining figures later:
CO2.T30ppm.Combined <- ggarrange(plot.CO2.T30ppm.noplant, plot.CO2.T30ppm.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CO2.T30ppm.Combined

```

```{r N2O T30 ppm Boxplot}
plot.N2O.T30ppm.noplant <- ggplot(ghg_ppm_filtered.noplant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("N"[2],"O (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 8))
 
plot.N2O.T30ppm.noplant


plot.N2O.T30ppm.plant <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("N"[2],"O (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 8))
 
plot.N2O.T30ppm.plant


#Panel together using ggarrange{ggpubr}:
# Version without legend for multi-panel figure:
N2O.T30ppm.Combined <- ggarrange(plot.N2O.T30ppm.noplant, plot.N2O.T30ppm.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

N2O.T30ppm.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.N2O.T30ppm.plant.autoScale <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.N2O.T30ppm.plant.autoScale

#ggsave("../figures/pub/Fig3/T30ppm/N2O_Inset_T30ppm.tiff", plot=plot.N2O.T30ppm.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset:
#subvp2_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp2_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#N2O.Combined
#print(plot.N2O.plant.autoScale, vp=subvp2_NO_legend)
```

```{r Extract and save GHG boxplot legend}
leg <- get_legend(plot.N2O.T30ppm.noplant, position = "right") #Extract legend as a legend Grob
legend_plot <- as_ggplot(leg) #Asign the grob as a ggplot
legend_plot #View the plot

#ggsave("../figures/pub/Fig3/Legend.tiff", plot=legend_plot, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```

```{r Combined GHG Boxplot}
comb_T30ppm_BoxPlot <- ggarrange(CH4.T30ppm.Combined, CO2.T30ppm.Combined, N2O.T30ppm.Combined, labels = c("A", "B", "C"), ncol = 3, nrow = 1, legend = "none")

comb_T30ppm_BoxPlot

#ggsave("../figures/pub/Fig3/T30ppm/BasePlots_T30ppm.tiff", plot=comb_T30ppm_BoxPlot, device="tiff", path=NULL, scale=1, width=15, height=6, dpi=600, limitsize=TRUE, bg="white")
```

## T30 ppm: No Plant, Plant, side-by-side

```{r CH4 T30 ppm NP vs. P}
plot.CH4.T30ppm.NPvP <- ggplot(ghg_ppm_filtered, aes(x=date_short, y=ch4_ppm)) +
                     geom_boxplot(aes(color=plant), outlier.shape = NA) +
                     geom_point(aes(color=plant), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#00571A","#BABABA"), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CH"[4]," (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.x.npc = "center", label.y.npc = "center", size = 7)
 
plot.CH4.T30ppm.NPvP
```

```{r CO2 T30 ppm NP vs P}
# Wilcox Test, nicer format, can use to verify stats on plot:
CO2_wilcox<- compare_means(co2_ppm~plant, ghg_ppm_filtered, group.by="date_short", method = "wilcox.test", paired = T)

# Plotting:
plot.CO2.T30ppm.NPvP <- ggplot(ghg_ppm_filtered, aes(x=plant, y=co2_ppm)) +
  facet_wrap(~date_short, nrow = 1, ncol = 4)+
                     geom_boxplot(aes(color=plant), outlier.shape = NA) +
                     geom_point(aes(color=plant), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("CO"[2]," (PPM)")), title="") +
  scale_x_discrete(labels=c("No Plant", "Plant"))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.y.npc = "top", label.x.npc = "center",size = 7)
 
plot.CO2.T30ppm.NPvP
```

```{r DELETE IF ABOVE WORKS}
CO2_NPvP_boxplot <- ggboxplot(ghg_ppm_filtered, x = "plant", y = "co2_ppm",
                              color = "plant",
                              add = "jitter",
                              facet.by = "date_short")

CO2_NPvP_boxplot + stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.x.npc = "center", label.y.npc = "center", size = 7)
```


```{r N2O T30 ppm Boxplot}
plot.N2O.T30ppm.noplant <- ggplot(ghg_ppm_filtered.noplant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("N"[2],"O (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 8))
 
plot.N2O.T30ppm.noplant


plot.N2O.T30ppm.plant <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("N"[2],"O (PPM)")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(0, 8))
 
plot.N2O.T30ppm.plant


#Panel together using ggarrange{ggpubr}:
# Version without legend for multi-panel figure:
N2O.T30ppm.Combined <- ggarrange(plot.N2O.T30ppm.noplant, plot.N2O.T30ppm.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

N2O.T30ppm.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.N2O.T30ppm.plant.autoScale <- ggplot(ghg_ppm_filtered.plant, aes(x=date_short, y=n2o_ppm)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.N2O.T30ppm.plant.autoScale

#ggsave("../figures/pub/Fig3/T30ppm/N2O_Inset_T30ppm.tiff", plot=plot.N2O.T30ppm.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset:
#subvp2_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp2_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#N2O.Combined
#print(plot.N2O.plant.autoScale, vp=subvp2_NO_legend)
```

```{r Extract and save GHG boxplot legend}
leg <- get_legend(plot.N2O.T30ppm.noplant, position = "right") #Extract legend as a legend Grob
legend_plot <- as_ggplot(leg) #Asign the grob as a ggplot
legend_plot #View the plot

#ggsave("../figures/pub/Fig3/Legend.tiff", plot=legend_plot, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```

```{r Combined GHG Boxplot}
comb_T30ppm_BoxPlot <- ggarrange(CH4.T30ppm.Combined, CO2.T30ppm.Combined, N2O.T30ppm.Combined, labels = c("A", "B", "C"), ncol = 3, nrow = 1, legend = "none")

comb_T30ppm_BoxPlot

#ggsave("../figures/pub/Fig3/T30ppm/BasePlots_T30ppm.tiff", plot=comb_T30ppm_BoxPlot, device="tiff", path=NULL, scale=1, width=15, height=6, dpi=600, limitsize=TRUE, bg="white")
```


## GHG plots flux

```{r subset plant and no plant flux data}
fluxes.noplant <- select_fluxes_combined %>%
  filter(plant=="N")

fluxes.plant <- select_fluxes_combined %>%
  filter(plant=="P")
```

```{r CH4 flux plot}
plot.CH4.flux.noplant <- ggplot(fluxes.noplant, aes(x=date_short, y=f0_CH4)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4]," m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-100, 20))
 
plot.CH4.flux.noplant


plot.CH4.flux.plant <- ggplot(fluxes.plant, aes(x=date_short, y=f0_CH4)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4]," m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-100, 20))
 
plot.CH4.flux.plant

#Panel together using ggarrange{ggpubr}:
#No legend version, for combining figures later:
CH4.flux.Combined <- ggarrange(plot.CH4.flux.noplant, plot.CH4.flux.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CH4.flux.Combined

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.CH4.flux.plant.autoScale <- ggplot(fluxes.plant, aes(x=date_short, y=f0_CH4)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.CH4.flux.plant.autoScale

#ggsave("../figures/pub/Fig3/Flux/CH4_Inset_flux.tiff", plot=plot.CH4.flux.plant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")
```

```{r CO2 flux plot}
plot.CO2.flux.noplant <- ggplot(fluxes.noplant, aes(x=date_short, y=f0_CO2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2]," m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-450, 200))
 
plot.CO2.flux.noplant


plot.CO2.flux.plant <- ggplot(fluxes.plant, aes(x=date_short, y=f0_CO2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2]," m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-450, 200))
 
plot.CO2.flux.plant

#Panel together using ggarrange{ggpubr}:

#No legend version, for combining figures later:
CO2.flux.Combined <- ggarrange(plot.CO2.flux.noplant, plot.CO2.flux.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

CO2.flux.Combined

# No plant has large negative outlier, remove and autoscale axis to show outlier:
plot.CO2.flux.noplant.autoScale <- ggplot(fluxes.noplant, aes(x=date_short, y=f0_CO2)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

 
plot.CO2.flux.noplant.autoScale

#ggsave("../figures/pub/Fig3/Flux/CO2_NoPlant_Inset.tiff", plot=plot.CO2.flux.noplant.autoScale, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")
```

```{r N2O T30 ppm Boxplot}
plot.N2O.flux.noplant <- ggplot(fluxes.noplant, aes(x=date_short, y=f0_N2O)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))
  #scale_y_continuous(limits = c(-4e+61, 1))
 
plot.N2O.flux.noplant


plot.N2O.flux.plant <- ggplot(fluxes.plant, aes(x=date_short, y=f0_N2O)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment), size = 1.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))
  #scale_y_continuous(limits = c(-1, 1))
 
plot.N2O.flux.plant

#Panel together using ggarrange{ggpubr}:

#No legend version, for combining figures later:
N2O.flux.Combined <- ggarrange(plot.N2O.flux.noplant, plot.N2O.flux.plant, labels = c("No Plant", "Plant"), ncol = 1, nrow = 2, legend = "none")

N2O.flux.Combined

# Add an inset to show small values for the 'no plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.N2O.flux.noplant.zoom <- ggplot(fluxes.noplant, aes(x=date_short, y=f0_N2O)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-1, 1))
 
plot.N2O.flux.noplant.zoom

#ggsave("../figures/pub/Fig3/Flux/N2O_NoPlant_inset.tiff", plot=plot.N2O.flux.noplant.zoom, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Add an inset to show small values for the 'plant' treatment:
# First, render plot with no axes labels and smaller axes text:
plot.N2O.flux.plant.zoom <- ggplot(fluxes.plant, aes(x=date_short, y=f0_N2O)) +
                     geom_boxplot(aes(color=treatment), outlier.shape = NA, lwd = 2) +
                     geom_point(aes(color=treatment), size = 3.5, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20, face = "bold"),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=30),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=2),strip.text = element_text(size = 20), legend.position="none", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = NULL, y = NULL, title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  scale_y_continuous(limits = c(-1, 1))
 
plot.N2O.flux.plant.zoom

#ggsave("../figures/pub/Fig3/Flux/N2O_plant_inset.tiff", plot=plot.N2O.flux.plant.zoom, device="tiff", path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE, bg="white")

# Create a viewport (vp) with appropriate dimensions for inset:
#subvp2_legend <- viewport(width = 0.35, height = 0.3, x = 0.67, y = 0.34)

#subvp2_NO_legend <- viewport(width = 0.35, height = 0.3, x = 0.81, y = 0.34)

# This viewport approach works, but only if ran in console, one line followed by the next. To save the image, must manually click through "Export" on "Plots" tab.
#N2O.Combined
#print(plot.N2O.plant.autoScale, vp=subvp2_NO_legend)
```

```{r Combined GHG Boxplot}
comb_flux_BoxPlot <- ggarrange(CH4.flux.Combined, CO2.flux.Combined, N2O.flux.Combined, labels = c("A", "B", "C"), ncol = 3, nrow = 1, legend = "none")

comb_flux_BoxPlot

#ggsave("../figures/pub/Fig3/Flux/BasePlots_flux.tiff", plot=comb_flux_BoxPlot, device="tiff", path=NULL, scale=1, width=15, height=6, dpi=600, limitsize=TRUE, bg="white")
```

## GHG summary

```{r GHG summary}
#Quick Data Summaries:

# History:
ghg.summary.history <- ghg %>%
  group_by(history)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.history

# Treatment:
ghg.summary.treatment <- ghg %>%
  group_by(treatment)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.treatment

# Plant:
ghg.summary.plant <- ghg %>%
  group_by(plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.plant

# Treatment and plant (tp):
ghg.summary.tp <- ghg %>%
  group_by(treatment, plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.tp

# Inlcude all factors: history, treatment, and plant (htp):
ghg.summary.htp <- ghg %>%
  group_by(history, treatment, plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.htp



# Use to make summary stats table (per reviewer comments):
#selfesteem %>%
  #group_by(time) %>%
  #get_summary_stats(score, type = "mean_sd")
```

# Mixed-effects modeling

```{r CH4 Do we have normaility?}
#First check the distribution of residuals --- not so great
ch4.lm <- lmer(mg.CH4.N.m.2~ (1|date) +(1|chamber), data=ghg, REML=TRUE)
summary(ch4.lm)
confint(ch4.lm)
p=profile(ch4.lm)
xyplot(p)
densityplot(p)
qqPlot(resid(ch4.lm), main="QQ Plot")
shapiro.test(resid(ch4.lm))

ghg$log.ch4 <- log(ghg$mg.CH4.N.m.2) # Log transforming (all GHGs) for better normality

ch4.lm <- lmer(log.ch4~ (1|date) + (1|chamber), data=ghg, REML=TRUE)
summary(ch4.lm)
confint(ch4.lm)
p=profile(ch4.lm)
xyplot(p)
densityplot(p)
qqPlot(resid(ch4.lm), main="QQ Plot")
shapiro.test(resid(ch4.lm))
```

```{r CH4 hyposthesis testing}
ch4.lm1 <- lmer(log.ch4 ~ (1|date) + (1|chamber) , data=ghg, REML=FALSE)

ch4.lm2 <- lmer(log.ch4~treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm3 <- lmer(log.ch4~plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm4 <- lmer(log.ch4~treatment * plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm5 <- lmer(log.ch4~ history + plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

#ch4.lm6 <- lmer(log.ch4 ~ (1|chamber) , data=ghg, REML=FALSE)

#ch4.lm7 <- lmer(log.ch4 ~ (1|date), data=ghg, REML=FALSE)

ch4.lm8 <- lmer(log.ch4~ plant + treatment + history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm9 <- lmer(log.ch4~ history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm10 <- lmer(log.ch4~ plant + treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm11 <- lmer(log.ch4~ plant * history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

ch4.lm12 <- lmer(log.ch4~ treatment + history + (1|date) + (1|chamber), data=ghg, REML=FALSE)
#This is a likelihood ratio test to get Chi squared value and p-value
##a <- anova(ch4.lm1, ch4.lm2, ch4.lm3,ch4.lm4,ch4.lm5, ch4.lm6, ch4.lm7, ch4.lm8, ch4.lm9, ch4.lm4a, ch4.lm7a, ch4.lm8a)
##a
##anova(ch4.lm6)

#AIC weights
c.mod <- list(ch4.lm1, ch4.lm2, ch4.lm3,ch4.lm4,ch4.lm5, ch4.lm8, ch4.lm9, ch4.lm10, ch4.lm11, ch4.lm12)
Modnames <- c("null", "treatment", "plant", "t*p", "h + p", "P+T+H", "H", "p+T", "p*h", "h+t")
aictab(cand.set = c.mod, modnames = Modnames, second.ord = TRUE)

#Variance explained by fixed effects and the entire model
r.squaredGLMM(ch4.lm1) #null
r.squaredGLMM(ch4.lm2) #treatment
r.squaredGLMM(ch4.lm3) #plant
r.squaredGLMM(ch4.lm4) #t * p
r.squaredGLMM(ch4.lm5) #h + p
r.squaredGLMM(ch4.lm8) #p+t+h
r.squaredGLMM(ch4.lm9) #h
r.squaredGLMM(ch4.lm10)#p+t
r.squaredGLMM(ch4.lm11)#p*h
r.squaredGLMM(ch4.lm12)#h+t

summary(ch4.lm8, ddf="Kenward-Roger")
confint.merMod(ch4.lm8, oldNames = FALSE)
p=profile(ch4.lm8)
xyplot(p)
densityplot(p)
qqPlot(resid(ch4.lm8), main="QQ Plot")
shapiro.test(resid(ch4.lm8))

#Interaction not significant so did not run post hoc
#Post hoc Tests
#Estimated marginal means - similar to lsmeans, uses t distribution
#https://cran.rstudio.com/web/packages/emmeans/vignettes/interactions.html

 emm <- emmeans(ch4.lm8, ~ treatment , adjust = "tukey")
 pairs(emm)
 multcomp::cld(emm, Letters=letters)
 
 emm <- emmeans(ch4.lm8, ~ history , adjust = "tukey")
 pairs(emm)
 multcomp::cld(emm, Letters=letters)
 
 emm <- emmeans(ch4.lm8, ~ plant , adjust = "tukey")
 pairs(emm)
 multcomp::cld(emm, Letters=letters)
  #p <-contrast(emm, "consec", simple="each", combine =TRUE, adjust="mvt")
  #p
```

```{r co2 Do we have normaility?}
#First check the distibution of residuals --- not so great
co2.lm <- lmer(mg.CO2.N.m.2~(1|chamber), data=ghg, REML=TRUE)
summary(co2.lm, ddf="Kenward-Roger")
confint(co2.lm)
p=profile(co2.lm)
xyplot(p)
densityplot(p)
qqPlot(resid(co2.lm), main="QQ Plot")
shapiro.test(resid(co2.lm))

ghg$log.co2 <- log(ghg$mg.CO2.N.m.2)

co2.lm <- lmer(log.co2~(1|chamber), data=ghg, REML=TRUE)
summary(co2.lm, ddf="Kenward-Roger")
confint(co2.lm)
p=profile(co2.lm)
xyplot(p)
densityplot(p)
qqPlot(resid(co2.lm), main="QQ Plot")
shapiro.test(resid(co2.lm))
```

```{r co2 hyposthesis testing}
co2.lm1 <- lmer(log.co2 ~ (1|date) + (1|chamber) , data=ghg, REML=FALSE)

co2.lm2 <- lmer(log.co2~treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm3 <- lmer(log.co2~plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm4 <- lmer(log.co2~treatment * plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm5 <- lmer(log.co2~ history + plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

#co2.lm6 <- lmer(log.co2 ~ (1|chamber) , data=ghg, REML=FALSE)

#co2.lm7 <- lmer(log.co2 ~ (1|date), data=ghg, REML=FALSE)

co2.lm8 <- lmer(log.co2~ treatment + history + plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm9 <- lmer(log.co2~ history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm10 <- lmer(log.co2~ plant+ treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm11 <- lmer(log.co2~ plant* history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

co2.lm12 <- lmer(log.co2~ treatment +history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

#This is a likelihood ratio test to get Chi squared value and p-value
##a <- anova(co2.lm1, co2.lm2, co2.lm3,co2.lm4,co2.lm5, co2.lm6, co2.lm7, co2.lm8, co2.lm9, co2.lm4a, co2.lm7a, co2.lm8a)
##a
##anova(co2.lm6)

#AIC weights
c.mod <- list(co2.lm1, co2.lm2, co2.lm3,co2.lm4,co2.lm5, co2.lm8, co2.lm9, co2.lm10, co2.lm11, co2.lm12)
Modnames <- c("null", "treatment", "plant", "t * p", "h + p","h+t+p", "H", "p+T", "p*h", "h+t")
aictab(cand.set = c.mod, modnames = Modnames, second.ord = TRUE)

#Variance explained by fixed effects and the entire model
r.squaredGLMM(co2.lm1) #null
r.squaredGLMM(co2.lm2) #treatment
r.squaredGLMM(co2.lm3) #plant
r.squaredGLMM(co2.lm4) #t * p
r.squaredGLMM(co2.lm5) #h + p
#r.squaredGLMM(co2.lm6) #chamber
#r.squaredGLMM(co2.lm7) #date
r.squaredGLMM(co2.lm8) #p+t+h
r.squaredGLMM(co2.lm9) #h
r.squaredGLMM(co2.lm10)#p+t
r.squaredGLMM(co2.lm11) #p*h
r.squaredGLMM(co2.lm12) #h+t

summary(co2.lm4, ddf="Kenward-Roger")
confint.merMod(co2.lm4, oldNames = FALSE)
p=profile(co2.lm4)
xyplot(p)
densityplot(p)
qqPlot(resid(co2.lm4), main="QQ Plot")
shapiro.test(resid(co2.lm4))

#Post hoc Tests
#Estimated marginal means - similar to lsmeans, uses t distribution
#https://cran.rstudio.com/web/packages/emmeans/vignettes/interactions.html

emm <- emmeans(co2.lm4, ~ plant*treatment , adjust = "tukey")
pairs(emm)
multcomp::cld(emm, Letters=letters)
 
  #p <-contrast(emm, "consec", simple="each", combine =TRUE, adjust="mvt")
  #p
```

```{r n2o Do we have normaility?}
#First check the distibution of residuals --- not so great
n2o.lm <- lmer(mg.N2O.N.m.2~(1|chamber), data=ghg, REML=TRUE)
summary(n2o.lm, ddf="Kenward-Roger")
confint(n2o.lm)
p=profile(n2o.lm)
xyplot(p)
#densityplot(p)
qqPlot(resid(n2o.lm), main="QQ Plot")
shapiro.test(resid(n2o.lm))

ghg$log.n2o <- log(ghg$mg.N2O.N.m.2)

#ghg.n2o <- ghg[-c(45,73, 111, 139),]
n2o.lm <- lmer(log.n2o~(1|chamber), data=ghg, REML=TRUE)
summary(n2o.lm, ddf="Kenward-Roger")
confint(n2o.lm)
p=profile(n2o.lm)
xyplot(p)
densityplot(p)
qqPlot(resid(n2o.lm), main="QQ Plot")
shapiro.test(resid(n2o.lm))
```

```{r n2o hyposthesis testing}
n2o.lm1 <- lmer(log.n2o ~ (1|date) + (1|chamber) , data=ghg, REML=FALSE)

n2o.lm2 <- lmer(log.n2o~treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm3 <- lmer(log.n2o~plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm4 <- lmer(log.n2o~treatment * plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm5 <- lmer(log.n2o~ history + plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

#n2o.lm6 <- lmer(log.n2o ~ (1|chamber) , data=ghg, REML=FALSE)

#n2o.lm7 <- lmer(log.n2o ~ (1|date), data=ghg, REML=FALSE)

n2o.lm8 <- lmer(log.n2o~ treatment + history + plant + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm9 <- lmer(log.n2o~ history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm10 <- lmer(log.n2o~ plant + treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm11 <- lmer(log.n2o~ plant * history + (1|date) + (1|chamber), data=ghg, REML=FALSE)

n2o.lm12 <- lmer(log.n2o~ history + treatment + (1|date) + (1|chamber), data=ghg, REML=FALSE)
#This is a liklihood ratio test to get Chi squared value and p-value
##a <- anova(n2o.lm1, n2o.lm2, n2o.lm3,n2o.lm4,n2o.lm5, n2o.lm6, n2o.lm7, n2o.lm8, n2o.lm9, n2o.lm4a, n2o.lm7a, n2o.lm8a)
##a
##anova(n2o.lm6)

#AIC weights
c.mod <- list(n2o.lm1, n2o.lm2, n2o.lm3,n2o.lm4,n2o.lm5, n2o.lm8, n2o.lm9, n2o.lm10, n2o.lm11, n2o.lm12)
Modnames <- c("null", "treatment", "plant", "t * p", "h + p","h+t+p", "H", "p+T", "p*h", "h+t")
aictab(cand.set = c.mod, modnames = Modnames, second.ord = TRUE)

#Variance explained by fixed effects and the entire model
r.squaredGLMM(n2o.lm1) #null
r.squaredGLMM(n2o.lm2) #treatment
r.squaredGLMM(n2o.lm3) #plant
r.squaredGLMM(n2o.lm4) #t * p
r.squaredGLMM(n2o.lm5) #h + p
#r.squaredGLMM(n2o.lm6) #chamber
#r.squaredGLMM(n2o.lm7) #date
r.squaredGLMM(n2o.lm8) #p+t+h
r.squaredGLMM(n2o.lm9) #h
r.squaredGLMM(n2o.lm10) #p+t
r.squaredGLMM(n2o.lm11) #p*h
r.squaredGLMM(n2o.lm12)#h+t

summary(n2o.lm10, ddf="Kenward-Roger")
confint.merMod(n2o.lm10, oldNames = FALSE)
p=profile(n2o.lm10)
xyplot(p)
densityplot(p)
qqPlot(resid(n2o.lm10), main="QQ Plot")
shapiro.test(resid(n2o.lm10))

emm <- emmeans(co2.lm10, ~ plant + treatment , adjust = "tukey")
 pairs(emm)
 multcomp::cld(emm, Letters=letters)
```

# Distance-based partial least squares regression (dbplsr)

```{r only ending samples}
#Using end samples for GHG because microbial community analyses was on end samples
#make df of design, soil, and ghg
ghg.filter.soil <- ghg %>%
  dplyr::select('date', 'chamber', 'timepoint', 'mg.CH4.N.m.2', 'mg.CO2.N.m.2', 'mg.N2O.N.m.2')%>%
  filter(timepoint=='T1' & date == '8/11/2016') 

soil.ghg <- soil %>% 
dplyr::select('chamber', 'time', 'plant', 'treatment', 'moisture.percent', 'pH', 'nh4.mg.l', 'c.percent', 'n.percent', 'P.ppm', 'K.ppm', 'Mg.ppm', 'S.ppm', 'Fe.ppm', 'Mn.ppm', 'HM', 'redox.percent.paint.removed') %>%
  filter(time=="finish") %>%
  bind_cols(ghg.filter.soil)
  
rownames(soil.ghg) <- rownames(otu.2.rel)

# Drop the two 'Chamber' columns (because Chambers are now row names)
soil.ghg <- soil.ghg[,-c(1,19)]

# Calculate euclidean distances for soil parameters only, soil parameters and GHGs, and GHGs only:
soilonly.dist <- vegdist(x = scale(soil.ghg[, -c(1:3, 17:21)]), "euclid")
soilghg.dist <- vegdist(x = scale(soil.ghg[, -c(1:3,17:18)]), "euclid")
ghg.dist <- vegdist(x = scale(soil.ghg[, -c(1:18)]), "euclid")
```

```{r dbplsr analysis}
#distance based partial least squares regression
#Compares distance matrices. can use mixture of continuous and qualitative variables. 
#Method good when factors are may and highly colinear. A PLS model find the multidimensional direction in the Z space that explains the variance in the Y space. 

# GHGs compared to Bray-Curtis Distance matrix of rarified OTUs: 
dbplsr.ch4 <- dbplsr(mg.CH4.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.ch4)

dbplsr.co2 <- dbplsr(mg.CO2.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.co2)

dbplsr.n2o <- dbplsr(mg.N2O.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.n2o)

# GHGs compared to euclidean distance matrix of soil parameters (sp): 
dbplsr.sp.ch4 <- dbplsr(mg.CH4.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.ch4)

dbplsr.sp.co2 <- dbplsr(mg.CO2.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.co2)

dbplsr.sp.n2o <- dbplsr(mg.N2O.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.n2o)
```

# Mantel test: soil and bacterial community

```{r mantel soil and bac comm }
# For reproducibility, confirm that random seed is set:
set.seed(42)

#soil only
bac.soil.mantel.1 <- vegan::mantel(as.matrix(soilonly.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.1

#soil and ghg
bac.soil.mantel.2 <- vegan::mantel(as.matrix(soilghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.2

#ghg only 
bac.soil.mantel.3 <- vegan::mantel(as.matrix(ghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.3
```

# Greenhouse gas and redox regressions

```{r CH4 redox regressions}
ch4.lm <- lm(log(mg.CH4.N.m.2) ~ redox.percent.paint.removed, data=soil.ghg)
anova(ch4.lm)
summary(ch4.lm)

#Outlier excluded, improves R2 slightly:
ch4.lm2 <- lm(log(mg.CH4.N.m.2) ~ redox.percent.paint.removed, data=soil.ghg[-c(17),])
anova(ch4.lm2)
summary(ch4.lm2)


ch4.redox.regress.plot <- ggplot(soil.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=mg.CH4.N.m.2, color=plant), size=5)+ 
  scale_color_manual(values=c("gray", "darkgreen"), labels=c("No Plant", "Plant"), name = "") +
  scale_fill_manual(values=c("gray", "darkgreen")) +
  geom_smooth(aes(x=(redox.percent.paint.removed*100), y=mg.CH4.N.m.2), method=lm, se=T, color="black")+
  #annotate("text",x=4.0,y=7, size =7, label="C")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = expression(paste("mg CH"[4],"-C m"^{-2}))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "right")

ch4.redox.regress.plot

#ggsave("../figures/pub/ch4_redox.tiff", plot=ch4.redox.regress.plot, device="tiff", path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)

#ggsave("../figures/pub/ch4_redox.jpg", plot=ch4.redox.regress.plot, device=NULL, path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)

ggsave("../figures/Thesis/ch4_redox.png", plot=ch4.redox.regress.plot, device="png", path=NULL, scale=1, width=6.3, height=4.4, dpi=300, limitsize=TRUE)
```

```{r CO2 redox regression}
co2.lm <- lm(mg.CO2.N.m.2 ~ redox.percent.paint.removed, data=soil.ghg)
anova(co2.lm)
summary(co2.lm)

co2.redox.regress.plot <- ggplot(soil.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=mg.CO2.N.m.2, color=plant), size=5)+ 
  scale_color_manual(values=c("gray", "darkgreen"), labels=c("No Plant", "Plant"), name = "") +
  scale_fill_manual(values=c("gray", "darkgreen")) +
  geom_smooth(aes(x=(redox.percent.paint.removed *100), y=mg.CO2.N.m.2), method=lm, se=T, color="black")+
  #annotate("text",x=4.0,y=7, size =7, label="C")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = expression(paste("mg CO"[2],"-C m"^{-2}))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "right")

co2.redox.regress.plot

#ggsave("../figures/pub/co2_redox.tiff", plot=co2.redox.regress.plot, device="tiff", path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)

#ggsave("../figures/pub/co2_redox.jpg", plot=co2.redox.regress.plot, device=NULL, path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)

ggsave("../figures/Thesis/co2_redox.png", plot=co2.redox.regress.plot, device="png", path=NULL, scale=1, width=6.3, height=4.4, dpi=300, limitsize=TRUE)
```

```{r N2O redox regression}
n2o.lm <- lm(mg.N2O.N.m.2 ~ redox.percent.paint.removed, data=soil.ghg)
anova(n2o.lm)
summary(n2o.lm)

n2o.redox.regress.plot <- ggplot(soil.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=mg.N2O.N.m.2, color=plant), size=5)+ 
  scale_color_manual(values=c("gray", "darkgreen"), labels=c("No Plant", "Plant"), name = "") +
  scale_fill_manual(values=c("gray", "darkgreen")) +
  geom_smooth(aes(x=(redox.percent.paint.removed *100), y=mg.N2O.N.m.2), method=lm, se=T, color="black")+
  #annotate("text",x=4.0,y=7, size =7, label="C")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "IRIS Percent", y = expression(paste("mg N"[2],"O-N m"^{-2}))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "right")

n2o.redox.regress.plot

#ggsave("../figures/pub/n2o_redox.tiff", plot=n2o.redox.regress.plot, device="tiff", path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)

#ggsave("../figures/pub/n2o_redox.jpg", plot=n2o.redox.regress.plot, device=NULL, path=NULL, scale=1, width=6.3, height=4.4, dpi=900, limitsize=TRUE)
```

```{r Panel together GHG ~ redox regressions}
GHG_redox_3panel <- ggarrange(ch4.redox.regress.plot, co2.redox.regress.plot, n2o.redox.regress.plot, labels = c("A", "B", "C"), ncol = 1, nrow = 3, legend = "right", common.legend = T)

GHG_redox_3panel

ggsave("../figures/pub/redox_regres.tiff", plot=GHG_redox_3panel, device="tiff", path=NULL, scale=1, width=7, height=10, dpi=900, limitsize=TRUE, bg="white")
```

