---
title: "TOWeR GHG microcosm experiment - Summer 2016"
author: "Regina B. Bledsoe, Colin G. Finlay, Ariane L. Peralta"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Setup

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#use to set working directory 
knitr::opts_knit$set(root.dir="~/GitHub/WetlandMesocosm_GHG_Timberlake/analyses")
```

# Load packages and functions

```{r load packages and functions, include FALSE }
#Set source R tools
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")

#load required packages
require("vegan")
require("tidyverse")
require("reshape2")
require("ecodist")
require("MASS")
require("MuMIn")
require("AICcmodavg")
require("emmeans")
require("car")
require("dbstats")
require("plyr")
require("ggpubr")
require("labdsv")
require("grid")
require("multcomp")
require("multcompView")
require("gasfluxes")
require("rstatix")
require("NSM3")
```

```{r Phyloseq install and load}
# If phyloseq not installed, un-comment and run:
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install(version = "3.20")

#BiocManager::install("phyloseq")

# Compatibility issue with lme4. Attempting to run without loading phyloseq.
require("phyloseq")
```

```{r Standard error and confidence intervals}
#Set Std Err and Conf Int
se <- function(x, ...) {
  sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
  1.96 * sd(x, na.rm = TRUE)
}
```
# Load files

```{r load files paths}
#Load data files
#Assign file names to variables
sharedfile = "../data/tmg2016.with.archaeaopti_mcc.shared"
taxfile = "../data/tmg2016.with.archaea.opti_mcc.0.03.cons.taxonomy"
metafile = "../data/tmg_design.csv"
```

```{r read in otu file and trim}
#Read in OTU file
otu <- read.otu(sharedfile)
rownames(otu)

#remove rogue sample... looks like there was a naming error at some point resulting in the additional sample below
otu.2 <- otu[!(rownames(otu) %in% "13PN"),]

#Removing microbial samples from "box" soils and keeping samples collected from within the chamber area since these should more directly relate to GHG flux data
c <- grep("*B", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#REMOVING STARTING SAMPLES
c <- grep("*S", rownames(otu.2), value=T)
otu.2 <- otu.2[!(rownames(otu.2) %in% c),]

#Remove Field samples. 
otu.2.nofield <- otu.2[!(rownames(otu.2) %in% "TL1"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL2"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL3"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL4"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL5"),]
otu.2.nofield <- otu.2.nofield[!(rownames(otu.2.nofield) %in% "TL6"),]

otu.2 <- otu.2.nofield

#Remove OTUs with less than 10 occurrences across all sites
b <- ncol(otu.2) #67355 OTUs,

# !=0 35,897
# =0 31,458
# =1 15,582
# =2  5,536
# =3  2,580
# =4  1,613
# =5  1,029
# =6    853
# =7    669
# =8    535
# =9    474
#=10    402

s<- sum(colSums(otu.2)) #1570135
otu.2 <- otu.2[, which(colSums(otu.2) >= 10)]
a <- ncol(otu.2) #7026 OTUs
sf<- sum(colSums(otu.2)) #1505797

#What sample has lowest read count? 
otu2<- colSums(t(otu.2))
otu2[which.min(colSums(t(otu.2)))]
otu2[which.max(colSums(t(otu.2)))]

#Lowest sample is 8NS 20370, highest 16NS 154034
#Most sample fall between 20370 and ~60000 with only a few higher than that so not removing the lowest sample 

#graph of read counts
d<-as.data.frame(otu.2)
d$colsum <- colSums(t(d))
d$group <- rownames(d)
p<-ggplot(d, aes(x=group,y=colsum))+geom_bar(stat="identity")+coord_flip()
p

#Rarefy
#Set min sample number
min.N <- min(rowSums(otu.2))
#rrarefy returns a randomly rarefied community data frame or vector of selected size
#This is the otu set that should be used for beta diversity metrics
otu.2 <- rrarefy(otu.2, min.N)
```

```{r read in meta data}
#Read in design file
meta <- read.csv(metafile)

#Remove mock, field, and box samples. 
rownames(meta) <- meta$sample
meta <- meta[!(rownames(meta) %in% "MOCK"),]
meta <- meta[!(rownames(meta) %in% "TL1"),]
meta <- meta[!(rownames(meta) %in% "TL2"),]
meta <- meta[!(rownames(meta) %in% "TL3"),]
meta <- meta[!(rownames(meta) %in% "TL4"),]
meta <- meta[!(rownames(meta) %in% "TL5"),]
meta <- meta[!(rownames(meta) %in% "TL6"),]
meta <- meta[!(meta$source %in% "b"),]
meta <- meta[!(meta$source %in% "s"),]
```

```{r soil}
soil <- read.csv("../data/Design_and_soil.csv",  header=T)

soil <- soil %>%
  filter(time!="field")%>%
  filter(time!="start")

by_time_plant_treatment <- soil %>%
  column_to_rownames("chamber")%>%
  dplyr::select(-history, -time) %>%
  group_by(plant, treatment) %>%
  dplyr::summarise_all(list(mean,sd, min, max), na.rm=T)

#write.csv(x = by_time_plant_treatment, file = "../figures/pub/soil_chem.csv")
```

# PERMANOVA

```{r PERMANOVA with just end samples}
#transforms read counts to relative abundance
otu.2.rel <- otu.2
for(i in 1:dim(otu.2)[1]){
  otu.2.rel[i,] <- otu.2[i,]/sum(otu.2[i,])
}

rownames(meta) <- rownames(otu.2.rel)
all.equal(rownames(otu.2.rel), rownames(meta))
metaotu <- cbind(meta, otu.2.rel)

# PERMANOVA:
metaotu.ad <- adonis2(metaotu[,-c(1:9)] ~ plant * history + plant * treatment, method = "bray", data = metaotu, perm=10000, set.seed=42)
metaotu.ad 
```
# PCoA

```{r Calculate distance matrix and PCoA components without field, end samples only}
otu.2.rel.dist<-vegdist(otu.2.rel, method="bray")

# Principal Coordinates Analysis
otu.2.rel.dist.pcoa <- cmdscale(otu.2.rel.dist, k=3, eig=TRUE, add=FALSE)
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1a <- round(otu.2.rel.dist.pcoa$eig[1] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
explainvar2a <- round(otu.2.rel.dist.pcoa$eig[2] / sum(otu.2.rel.dist.pcoa$eig), 3) * 100
sum.eiga <- sum(explainvar1a, explainvar2a)

explainvar1a #25% (rounding)
explainvar2a #11% (rounding)
```

```{r plot PCoA without field, end samples only}
all.equal(rownames(meta), rownames(otu.2.rel))
# Set treatments
treat1 <- as.factor(meta$history)
droplevels(treat1)
treat2 <- as.factor(meta$treatment)
droplevels(treat2)
#levels(treat2) <- c("I","W","D")
treat3 <- as.factor(meta$plant)
droplevels(treat3)

pcoa.groups <- paste(meta$history, meta$treatment, meta$plant, sep = "_")

pcoa.points <- data.frame(otu.2.rel.dist.pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)

pcoa.plant <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 3))  #plant
pcoa.trt <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 2)) #treatment
pcoa.history <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 1))  #history 

pcoa.trt <- ordered(pcoa.trt, levels=c("D", "I", "W"))
pcoa.trt <-revalue(pcoa.trt, c("D"="Dry", "I"="Interim","W"="Wet"))

pcoa.history <- ordered(pcoa.history, levels=c("D", "I", "W"))
pcoa.history <-revalue(pcoa.history, c("D"="Dry", "I"="Interim","W"="Wet"))

pcoa.plant <- ordered(pcoa.plant, levels=c("n", "p"))
pcoa.plant <-revalue(pcoa.plant, c("n"="No Plant","p"="Plant"))

soil.finish <- soil %>% 
dplyr::select(chamber, time, treatment, moisture.percent, pH, nh4.mg.l, no2.no3.mg.l, c.percent, n.percent, P.ppm, K.ppm, Mg.ppm, S.ppm, Fe.ppm, Mn.ppm, HM, CEC)

c<-cor(soil.finish[, -c(1:3,7)]) #design fields and no3 because many missing values
c<-cor(soil.finish[, -c(1:3,7,9,12,14,17)]) # %N Mg Fe CEC

# Add points data frame (seems to be newer issue, used to be able to use $ on pcoa_denit_FINAL$points)
pcoa_otu_DF <- as.data.frame(otu.2.rel.dist.pcoa$points)

# envfit:
meta.env <- soil.finish[, -c(1:3,7,9,12,14,17)]
fit<-envfit(pcoa_otu_DF, meta.env, perm=1000,na.rm=T, set.seed(42))
A <-as.list(fit$vectors)
vec<-as.data.frame(fit$vectors$arrows*sqrt(fit$vectors$r)*0.25) #Scale vector by correlation
p<- as.data.frame(A$pvals)

vec <- cbind(vec, p) 
vec <-  subset(vec, A$pvals<=0.05)

vec$parm<-rownames(vec)
colnames(vec)
rownames(vec)
vec$parm<-c("pH", "%C", "S", "Mn") # Check that these readable labels match the originals

# To use open and closed shapes, must create a combined category for History and Plant:
history_plant_combined <- c(paste(pcoa.history, pcoa.plant, sep = "_"))

#Make new PCoA data frame, then append experimental factors (plant, treatment, history) as new columns:
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$history <- as.factor(pcoa.history)
pcoa.cent.dataframe.trts$trt <- as.factor(pcoa.trt)
pcoa.cent.dataframe.trts$plant <- as.factor(pcoa.plant)
pcoa.cent.dataframe.trts$history_plant <- as.factor(history_plant_combined)

#Plotting:
plot16S_PCoA <- ggplot(pcoa.cent.dataframe.trts, aes(x=V1, y=V2, shape=history_plant, colour=trt)) + geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +   
  geom_point(aes(shape = history_plant, color=pcoa.trt), size=5, stroke=2) +  
  theme_bw() +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  scale_shape_manual(labels = c("Dry_No Plant", "Dry_Plant", "Interim_No Plant", "Interim_Plant", "Wet_No Plant", "Wet_Plant"), values = c(0, 15, 2, 17, 1, 19), name="History_Plant") +
  xlab("PCoA 1 (25%)") + ylab("PCoA 2 (11%)") + 
  geom_segment(data=vec, aes(x=0,xend=V1,y=0,yend=V2), size=1, arrow = arrow(length = unit(0.2, "cm")),colour="black",inherit.aes=F)+ 
  #Label soil vectors 
  geom_text(data=vec,aes(x=V1-0.014,y=V2+0.011,label=parm),size=5, inherit.aes=F)+
  ggtitle("") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=18), axis.text=element_text(size=18), 
          axis.text.x = element_text(size=18),  axis.text.y = element_text(size=18),
          panel.border = element_rect(colour = "black", size=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=18)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(fill = guide_legend(override.aes=list(shape=21)))

plot16S_PCoA

#ggsave("../figures/pub/16S_PCoA.tiff", plot=plot16S_PCoA, device="tiff", path=NULL, scale=1, width=9.5, height=5, dpi=900, limitsize=TRUE)

#ggsave("../figures/pub/16S_PCoA.png", plot=plot16S_PCoA, device="png", path=NULL, scale=1, width=9.5, height=5, dpi=300, limitsize=TRUE)
```
# Indicator species

```{r read and clean taxonomy file}
tax <- import_mothur(mothur_constaxonomy_file =  taxfile)
tax.df <- as.data.frame(tax)

tax.df <- tax.df %>% 
  rownames_to_column(var = "otu")

colnames(tax.df)=c("otu","domain","Phylum","Class", "Order","Family","Genus")

#Clean up taxonomy file. Remove underscores and unclassified
tax.df <- tax.df %>%
    mutate(Family = str_replace(Family, "_family_incertae_sedis", ""), Genus = str_replace(Genus, "_genera_incertae_sedis", "")) %>%
  mutate(Phylum = str_replace_all(Phylum,"_unclassified", ""), Order = str_replace_all(Order,"_unclassified", ""),Class = str_replace_all(Class,"_unclassified", ""), Family = str_replace_all(Family,"_unclassified", ""), Genus = str_replace_all(Genus,"_unclassified", "")) %>%
  mutate(Phylum = str_replace_all(Phylum,"_", " "), Order = str_replace_all(Order,"_", " "),Class = str_replace_all(Class,"_", " "), Family = str_replace_all(Family,"_", " "), Genus = str_replace_all(Genus,"_", " ")) 
```

```{r Indicator Species}
#Final sampling, no field
new.data <-metaotu
new.data$history <- as.factor(new.data$history)
new.data$treatment <- as.factor(new.data$treatment)

#group = interaction(new.data$plant, new.data$treatment, new.data$history) #Setup to analyze all three factors
# Only History and Treatment had significant results in PERMANOVA (i.e., adonis) (see above)
# Therefore, only History and Treatment will be used to group bacterial indicator species analysis:
group.hist <- new.data$history
group.trt <- new.data$treatment

dataREL <- new.data[,-c(1:9)] # Only OTU columns (i.e., relative [REL] abundance data), no metadata columns

dataREL <- dataREL[, colSums(dataREL) > 0.025] # Drop very rare OTUs from the analysis; only analyze most abundant 2.5%

bac.ind_hist <- indval(dataREL, group.hist, numitr = 10000, set.seed(42))
bac.ind_trt <- indval(dataREL, group.trt, numitr = 10000, set.seed(42))

summary(bac.ind_hist)
summary(bac.ind_trt)

# p-value cutoff:
inds_hist <- which(bac.ind_hist$pval <= 0.01)
inds_trt <- which(bac.ind_trt$pval <= 0.01)

# Make data frame skeleton:
bac.indicators_hist <- as.data.frame(matrix(NA, nrow = length(inds_hist), ncol = 4))
bac.indicators_trt <- as.data.frame(matrix(NA, nrow = length(inds_trt), ncol = 4))

# Set column names:
colnames(bac.indicators_hist) <- c("OTU", "Cluster", "IndVal", "Prob")
colnames(bac.indicators_trt) <- c("OTU", "Cluster", "IndVal", "Prob")

# Poplulate OTU column:
bac.indicators_hist$OTU <- names(inds_hist)
bac.indicators_trt$OTU <- names(inds_trt)

# Populate Cluster column:
bac.indicators_hist$Cluster <- bac.ind_hist$maxcls[inds_hist]
bac.indicators_trt$Cluster <- bac.ind_trt$maxcls[inds_trt]

# Populate IndVal column:
bac.indicators_hist$IndVal <- bac.ind_hist$indcls[inds_hist]
bac.indicators_trt$IndVal <- bac.ind_trt$indcls[inds_trt]

# Populate p-value column:
bac.indicators_hist$Prob <- bac.ind_hist$pval[inds_hist]
bac.indicators_trt$Prob <- bac.ind_trt$pval[inds_trt]

otu.tax <- as.data.frame(tax.df)

ind.tax_hist <- otu.tax[which(as.character(otu.tax$otu) %in% bac.indicators_hist$OTU), ]
ind.tax_trt <- otu.tax[which(as.character(otu.tax$otu) %in% bac.indicators_trt$OTU), ]

ind.tax_hist <- ind.tax_hist[match(ind.tax_hist$otu, bac.indicators_hist$OTU), ]
ind.tax_trt <- ind.tax_trt[match(ind.tax_trt$otu, bac.indicators_trt$OTU), ]

# Add taxonomic information:
indicator.bac_hist <- cbind(bac.indicators_hist, ind.tax_hist[, -c(1)])
indicator.bac_trt <- cbind(bac.indicators_trt, ind.tax_trt[, -c(1)])

# Order rows by Cluster:
indicator.bac_hist <- indicator.bac_hist[order(as.numeric(indicator.bac_hist$Cluster)), ]
indicator.bac_trt <- indicator.bac_trt[order(as.numeric(indicator.bac_trt$Cluster)), ]

# Print information from tables:
table(indicator.bac_hist$Cluster)
table(indicator.bac_trt$Cluster)

table(indicator.bac_hist$Phylum)
table(indicator.bac_trt$Phylum)

# Identify what the Cluster numbers mean:
levels(group.hist)
levels(group.trt)

#Replace cluster numbers with factor names:
indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 1] <- "Dry"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 1] <- "Dry"

indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 2] <- "Interim"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 2] <- "Interim"

indicator.bac_hist["Cluster"][indicator.bac_hist["Cluster"] == 3] <- "Wet"
indicator.bac_trt["Cluster"][indicator.bac_trt["Cluster"] == 3] <- "Wet"

#Find overlapping indicator tax in hist and trt:
common_indicators <- generics::intersect(indicator.bac_hist$OTU, indicator.bac_trt$OTU)
common_indicators <- as.data.frame(common_indicators) #Make it a data frame
colnames(common_indicators) <- "OTU" #Rename column header to "OTU" to match with indicator.bac_hist and indicator.bac_trt

common_indicators <- left_join(common_indicators, indicator.bac_hist, by = "OTU") #Add data from indicator.bac_hist
colnames(common_indicators)[2:4] <- c("Cluster_hist", "IndVal_hist", "Prob_hist") #Rename to not confuse with treatment indicators, added next:
  #Add data from indicator.bac_trt:
common_indicators <- left_join(common_indicators, indicator.bac_trt, by = c("OTU", "domain", "Phylum", "Class", "Order", "Family", "Genus")) 
colnames(common_indicators)[11:13] <- c("Cluster_treatment", "IndVal_treatment", "Prob_treatment")

# Export Bacteria Indicator Tables:
#write.table(indicator.bac_hist, "../figures/pub/BacterialIndicators_History.txt",
           #sep="\t", row.names = F, quote = F)

#write.table(indicator.bac_trt, "../figures/pub/BacterialIndicators_Treatment.txt",
            #sep="\t", row.names = F, quote = F)
```

# Soil parameters

```{r soil parameters}
soil <- read.csv("../data/Design_and_soil.csv",  header=T)

# Remove field samples:
soil_nofield <- soil %>%
  filter(soil$time!="field")

# For axes titles, change abbreviations (D, I, W) to full words (Dry, Interim, Wet):
soil_nofield$treatment <-revalue(soil_nofield$treatment, c("D" = "Dry", "I" = "Interim","W" = "Wet"))

# For axes titles, change abbreviations (N, P) to full words (No Plant, Plant):
soil_nofield$plant <- revalue(soil_nofield$plant, c("N" = "No Plant", "P" = "Plant"))

# Make soil_nofield$time uppercase, first letter only:
soil_nofield$time <- str_to_title(soil_nofield$time)

#write.csv(soil_nofield, file="../figures/soil.csv")

#Subset Start and Finish for plotting:
soil_nofield.Start <- subset(soil_nofield, time == "Start")
soil_nofield.Finish <- subset(soil_nofield, time == "Finish")

#Plotting:
#Start:
iris_tubes.Start <- ggplot(soil_nofield.Start, aes(x=treatment, y=(redox.percent.paint.removed*100))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(~plant) +
  geom_point(aes(color=factor(treatment)), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  theme_bw() +
  #Remove plot grid lines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = "IRIS Percent", title="") +
  scale_x_discrete(labels=c("", "", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

iris_tubes.Start

#Finish:
iris_tubes.Finish <- ggplot(soil_nofield.Finish, aes(x=treatment, y=(redox.percent.paint.removed*100))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(~plant) +
  geom_point(aes(color=factor(treatment)), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), name="Treatment") +
  theme_bw() +
  #Remove plot grid lines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(text=element_text(size=14),axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right", legend.title = element_text(size=14), legend.text=element_text(size=14)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "Treatment", y = "IRIS Percent", title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

iris_tubes.Finish

#Panel together using ggarrange{ggpubr}:
iris_tubes.Combined <- ggarrange(iris_tubes.Start, iris_tubes.Finish, labels = c("Start (Week 0)", "Finish (Week 8)"), ncol = 1, nrow = 2, legend = "right", common.legend = T)

iris_tubes.Combined

#ggsave("../figures/pub/iris_tubes.tiff", plot=iris_tubes.Combined, device="tiff", path=NULL, scale=1, width=6, height=5, dpi=900, limitsize=TRUE, bg="white")

#ggsave("../figures/pub/iris_tubes.png", plot=iris_tubes.Combined, device=NULL, path=NULL, scale=1, width=6, height=5, dpi=600, limitsize=TRUE ,bg="white")

#ggsave("../figures/pub/iris_tubes.pdf", plot=iris_tubes.Combined, device=NULL, path=NULL, scale=1, width=6, height=5, dpi=900, limitsize=TRUE ,bg="transparent")
```

```{r IRIS stats}
# For FeS paper
# Test parametric assumptions:
# Outliers:
Outliers_IRIS <- soil_nofield %>%
  identify_outliers(redox.percent.paint.removed)
Outliers_IRIS
  # There are no extreme outliers

# Normality:
soil_nofield %>%
  shapiro_test(redox.percent.paint.removed)
  # Not normally distributed

  # QQ plot:
ggqqplot(soil_nofield, "redox.percent.paint.removed", facet.by = "time")
```



```{r soil moisture boxplot}
soil_moisture <- ggplot(soil_nofield, aes(x=treatment, y=moisture.percent, fill=plant)) + 
  geom_boxplot() +
  facet_grid(~ fct_rev(time)) +
 scale_fill_manual(values=c("gray", "darkgreen"), labels=c("No Plant", "Plant")) +
#Set plot elements
theme_bw()  +
#Remove plot gridlines
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#Set axis title and text properties, tick marks, and labels
theme(axis.title=element_text(size=14,face="bold"),
          axis.text=element_text(size=14),  
          axis.title.y=element_text(margin=margin(r=10)),
          panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 14), legend.position="right") + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = "Moisture Percent", title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))

soil_moisture

#ggsave("../figure/pub/soil_moisture.tiff", plot=soil_moisture, device="tiff", path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE, bg="transparent")

#ggsave("../figures/pub/soil_moisture.jpg", plot=soil_moisture, device=NULL, path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE , bg="transparent")

#ggsave("../figures/pub/soil_moisture.pdf", plot=soil_moisture, device=NULL, path=NULL, scale=1, width=6, height=6.3, dpi=900, limitsize=TRUE , bg="transparent")
```

```{r soil plots}
ammonium <- ggplot(soil_nofield, aes(x=treatment, y=nh4.mg.l)) + 
  geom_boxplot(aes(x=treatment, y=nh4.mg.l, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment")+
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Ammonium (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

ammonium

nitrate <- ggplot(soil_nofield, aes(x=treatment, y=no2.no3.mg.l)) + 
  geom_boxplot(aes(x=treatment, y=no2.no3.mg.l, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment")+
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Nitrate (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

nitrate

soil_Fe <- ggplot(soil_nofield, aes(x=treatment, y=Fe.ppm)) + 
  geom_boxplot(aes(x=treatment, y=Fe.ppm, fill=treatment)) +
  scale_fill_manual(values=c("#997700","#6699CC","#004488"), name="Treatment", labels=c("Dry", "Interim", "Wet"))+
  scale_x_discrete(labels=c("Dry", "Interim", "Wet"), name="") +
  facet_wrap(. ~plant) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor =element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(colour = "black",size=1)) + 
  #stat_summary(fun.data=mean_cl_boot,size=1.25, position=position_dodge(width=0.75)) + 
#scale_color_manual(values=cbbPalette, name="Treatment")+
labs(x = "Treatment", y = "Iron (mg/L)") + theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14))

soil_Fe
```

# Greenhouse gases (GHGs)

```{r GHG T1 concentration analysis}
# Version of ghg with labels for T1 through T4 analysis:
ghg2 <- read.csv("../data/TMG_GHG-Flux_Pub_3.csv", header=TRUE)

# FIX c2o TYPO IN ghg2:
colnames(ghg2)[18] <- "co2_ppm"

# ghg2, select T1 through T4, and all ppm measurements:
ghg_ppm <- ghg2 %>%
  dplyr::select('date', 'chamber', 'timepoint', 'plant', 'ch4_ppm', 'co2_ppm', 'n2o_ppm')%>%
  dplyr::filter(date !='7/24/2016')

# Set factors:
ghg_ppm$chamber <- as.factor(ghg_ppm$chamber)
ghg_ppm$date <- as.factor(ghg_ppm$date)

# Change design$box to design$chamber in order to join with ghg$chamber
colnames(meta)[2] <- "chamber"

# Add on meta data
ghg_ppm <- ghg_ppm %>%
  left_join(meta[,-c(3, 4, 7, 8, 9)], by="chamber")

# Add new date column, to shorten:
ghg_ppm$date_short <- ghg_ppm$date
# remove "/2016" from date_short:
ghg_ppm$date_short <- str_remove(ghg_ppm$date_short, "/2016")

# Separate data frames for "No Plant" and "Plant", for plotting:
ghg_ppm.noplant <- ghg_ppm %>%
  filter(plant=="N")

ghg_ppm.plant <- ghg_ppm %>%
  filter(plant=="P")
```

```{r Calculate T0s from WRC static chamber monitoring}
# will attempt to use results from monthly WRC static chamber to calculate average T0. Then calculate fluxes from TMG mesocosms.

# read in compiled T0 data:
WRC_T0s <- read.csv("../data/WRC_T0s_for_TMG_ambient_calculation.csv")

# Set Sample_Date as factor:
WRC_T0s$Sample_Date <- as.factor(WRC_T0s$Sample_Date)

# calculate averages within each month:
CO2_T0_monthly_avg <- aggregate(T0_CO2_ppm ~ Sample_Date, data = WRC_T0s, FUN = mean, na.rm = TRUE)
CH4_T0_monthly_avg <- aggregate(T0_CH4_ppm ~ Sample_Date, data = WRC_T0s, FUN = mean, na.rm = TRUE)
N2O_T0_monthly_avg <- aggregate(T0_N2O_ppm ~ Sample_Date, data = WRC_T0s, FUN = mean, na.rm = TRUE)

# Average the two years of August data:
CO2_T0_monthly_avg[nrow(CO2_T0_monthly_avg)+1, "T0_CO2_ppm"] <- mean(c(CO2_T0_monthly_avg[1, "T0_CO2_ppm"], CO2_T0_monthly_avg[2, "T0_CO2_ppm"]), na.rm = TRUE)

CH4_T0_monthly_avg[nrow(CH4_T0_monthly_avg)+1, "T0_CH4_ppm"] <- mean(c(CH4_T0_monthly_avg[1, "T0_CH4_ppm"], CH4_T0_monthly_avg[2, "T0_CH4_ppm"]), na.rm = TRUE)

N2O_T0_monthly_avg[nrow(N2O_T0_monthly_avg)+1, "T0_N2O_ppm"] <- mean(c(N2O_T0_monthly_avg[1, "T0_N2O_ppm"], N2O_T0_monthly_avg[2, "T0_N2O_ppm"]), na.rm = TRUE)

# Add in the name:
CO2_T0_monthly_avg$Sample_Date <- factor(c("Aug_2022", "August_2021", "Jul_2022", "June_2022", "August_21_22"))
CH4_T0_monthly_avg$Sample_Date <- factor(c("Aug_2022", "August_2021", "Jul_2022", "June_2022", "August_21_22"))
N2O_T0_monthly_avg$Sample_Date <- factor(c("Aug_2022", "August_2021", "Jul_2022", "June_2022", "August_21_22"))
```

```{r wrangle T0 averages for flux calculation}
# bind together the three GHGs:
T0_averages_combined <- CH4_T0_monthly_avg %>%
  cbind(CO2_T0_monthly_avg$T0_CO2_ppm) %>%
  cbind(N2O_T0_monthly_avg$T0_N2O_ppm)

# drop the two August monthly:
T0_averages_combined <- T0_averages_combined[3:5,]

# Rename 'Sample_Date' in T0_averages_combined to 'date' to match the 'ghg_flux' column
colnames(T0_averages_combined)[1] <- "date"

# Add chamber as NAs:
T0_averages_combined$chamber <- NA

# Rename GHG columns to match ghg_flux:
colnames(T0_averages_combined)[2:4] <- c("ch4_ppm", "co2_ppm", "n2o_ppm")

# Add time columns to match ghg_flux:
T0_averages_combined$time_min <- 0
T0_averages_combined$time <- 0

# Add month column:
T0_averages_combined$month <- c("July", "June", "August")
```

```{r GHG flux wrangling}
# Format a data frame for 'gasfluxes' package:
ghg_flux <- ghg2[,c(1:2, 12, 15:19)]

# Remove 7/24/2016 from data frame, as per Gina's advice
ghg_flux <- ghg_flux %>%
  dplyr::filter(date !='7/24/2016')

# Combine 'date' and 'chamber' strings to create a unique ID:
ghg_flux$ID <- paste(ghg_flux$date, ghg_flux$chamber)

# convert Volume (cm^3) to Volume (m^3):
ghg_flux$V <- (ghg_flux$Vol..cm3. * (1/1e+06))

# Copy 'Area..m2.' to a new column named 'A':
ghg_flux$A <- ghg_flux$Area..m2.

# Convert time in minutes to time in hours:
ghg_flux$time <- (ghg_flux$time_min * (1/60))

# Data frame to bind T0s from WRC:
ghg_flux_w_T0 <- ghg_flux

# Add in month column that takes date strings and matches them to "July", "June", "August":
# Create a vector of month names
month_names <- c("January", "February", "March", "April", "May", "June",
                 "July", "August", "September", "October", "November", "December")

# Extract the month as a number and convert it to a month name
ghg_flux_w_T0 <- ghg_flux_w_T0 %>%
  mutate(
    month = month_names[as.numeric(substr(date, 1, 1))]
  )

# Copy July and June rows because there are two July and June sampling dates:
T0_averages_combined <- rbind(T0_averages_combined, T0_averages_combined[1,])
T0_averages_combined <- rbind(T0_averages_combined, T0_averages_combined[2,])

# Add TMG-style dates to match with ghg_flux:
T0_averages_combined$date_TMG <- c("7/11/2016", "6/13/2016", "8/11/2016", "7/25/2016", "6/28/2016")

# For loop:
for (unique_date in unique(ghg_flux_w_T0$date)) {
  for (unique_chamber in unique(ghg_flux_w_T0$chamber)) {
    
    # Filter the monthly average corresponding to the current date
    month_avg <- T0_averages_combined %>% filter(date_TMG == unique_date)
    
    # Skip if there is no corresponding monthly average for the date
    if (nrow(month_avg) == 0) next
    
    # Create a new row with time = 0 and the averages for the current date and chamber
    new_row <- data.frame(
      date = unique_date,
      chamber = unique_chamber,
      time = month_avg$time,
      time_min = month_avg$time_min,
      ch4_ppm = month_avg$ch4_ppm,
      co2_ppm = month_avg$co2_ppm,
      n2o_ppm = month_avg$n2o_ppm,
      month = month_avg$month
    )
    
    # Bind the new row to ghg_flux
    ghg_flux_w_T0 <- bind_rows(ghg_flux_w_T0, new_row)
  }
}

# Reorder rows by date, chamber, and time:
ghg_flux_w_T0 <- ghg_flux_w_T0 %>%
  arrange(date, chamber, time)

# Copy some of rows following T0s onto T0s (ID, V, A):
# Define the columns with missing values to be filled from the row below
columns_to_fill <- c("ID", "V", "A")

# Loop through the rows from the bottom to the top to fill missing values from the row below
for (i in (nrow(ghg_flux_w_T0) - 1):1) {
  # Check for missing values in the specified columns
  missing_values <- is.na(ghg_flux_w_T0[i, columns_to_fill])
  
  # If there are missing values, copy the values from the row below
  if (any(missing_values)) {
    ghg_flux_w_T0[i, columns_to_fill][missing_values] <- ghg_flux_w_T0[i + 1, columns_to_fill][missing_values]
  }
}



# Convert ppm to mg/m^3, based on https://www.cdc.gov/niosh/docs/2004-101/calc.html (based on 25°C and 1 atm):
# y mg/m3 = (x ppm)(molar mass)/24.45
  # 24.45 is the volume (liters) of a mole (gram molecular weight) of a gas or vapor when the pressure is at 1 atmosphere (760 torr or 760 mm Hg) and at 25°C

# CH4 mg/m3
# C is the package 'gasfluxes' abbreviation for concentration. 16.04 g/mol = molar mass of CH4
ghg_flux_w_T0$C_ch4 <- ((ghg_flux_w_T0$ch4_ppm * 16.04)/24.45)

# CO2 mg/m3
# molar mass of CO2 = 44.01 g/mol
ghg_flux_w_T0$C_co2 <- ((ghg_flux_w_T0$co2_ppm * 44.01)/24.45)

# N2O mg/m3
# molar mass of N2O = 44.013 g/mol
ghg_flux_w_T0$C_n2o <-  ((ghg_flux_w_T0$n2o_ppm * 44.013)/24.45)

# Split off data frames for each greenhouse gas:
ch4_flux_input_w_T0 <- ghg_flux_w_T0[,c(9:12, 14)]

co2_flux_input_w_T0 <- ghg_flux_w_T0[,c(9:12, 15)]

n2o_flux_input_w_T0 <- ghg_flux_w_T0[,c(9:12, 16)]

# Rename C_ghg columns to 'C':
colnames(ch4_flux_input_w_T0)[5] <- "C"

colnames(co2_flux_input_w_T0)[5] <- "C"

colnames(n2o_flux_input_w_T0)[5] <- "C"

# Must omit NAs before using gasfluxes():
ch4_flux_input_w_T0 <- na.omit(ch4_flux_input_w_T0)

co2_flux_input_w_T0 <- na.omit(co2_flux_input_w_T0)

# n2o_flux_input has negative values, will need to be converted to NAs, then all NAs omitted
n2o_flux_input_w_T0$C[n2o_flux_input_w_T0$C<0] <- NA
n2o_flux_input_w_T0 <- na.omit(n2o_flux_input_w_T0)
```

```{r Make T0 flux dataframes' V and A numeric}
ch4_flux_input_w_T0$V <- as.numeric(ch4_flux_input_w_T0$V)
ch4_flux_input_w_T0$A <- as.numeric(ch4_flux_input_w_T0$A)

co2_flux_input_w_T0$V <- as.numeric(co2_flux_input_w_T0$V)
co2_flux_input_w_T0$A <- as.numeric(co2_flux_input_w_T0$A)

n2o_flux_input_w_T0$V <- as.numeric(n2o_flux_input_w_T0$V)
n2o_flux_input_w_T0$A <- as.numeric(n2o_flux_input_w_T0$A)
```

```{r GHG flux calculation}
ch4_flux.results_w_T0 <- gasfluxes(ch4_flux_input_w_T0, method = c("linear","robust linear", "HMR"), plot = T)
co2_flux.results_w_T0 <- gasfluxes(co2_flux_input_w_T0, method = c("linear","robust linear", "HMR"), plot = T)
n2o_flux.results_w_T0 <- gasfluxes(n2o_flux_input_w_T0, method = c("linear","robust linear", "HMR"), plot = T)
```

```{r GHG flux results wrangling}
# If HMR is available, use HMR
# If HMR is not available, use robust linear
# If robust linear is not available use linear

# New DF with all three f0s:
ch4_fluxes_only_w_T0 <- ch4_flux.results_w_T0[,c(1:2, 11, 17)]
co2_fluxes_only_w_T0 <- co2_flux.results_w_T0[,c(1:2, 11, 17)]
n2o_fluxes_only_w_T0 <- n2o_flux.results_w_T0[,c(1:2, 11, 17)]

# New column for final flux selection:
ch4_fluxes_only_w_T0$f0_selected <- ch4_fluxes_only_w_T0$HMR.f0
co2_fluxes_only_w_T0$f0_selected <- co2_fluxes_only_w_T0$HMR.f0
n2o_fluxes_only_w_T0$f0_selected <- n2o_flux.results_w_T0$HMR.f0

# Where HMR = NA, replace with the robust linear value:
ch4_fluxes_only_w_T0$f0_selected[na.action(na.omit(ch4_fluxes_only_w_T0$HMR.f0))] <- ch4_fluxes_only_w_T0$robust.linear.f0[na.action(na.omit(ch4_fluxes_only_w_T0$HMR.f0))]

co2_fluxes_only_w_T0$f0_selected[na.action(na.omit(co2_fluxes_only_w_T0$HMR.f0))] <- co2_fluxes_only_w_T0$robust.linear.f0[na.action(na.omit(co2_fluxes_only_w_T0$HMR.f0))]

n2o_fluxes_only_w_T0$f0_selected[na.action(na.omit(n2o_fluxes_only_w_T0$HMR.f0))] <- n2o_fluxes_only_w_T0$robust.linear.f0[na.action(na.omit(n2o_fluxes_only_w_T0$HMR.f0))]

# Where f0_selected still has NAs, replace with the linear.f0:
ch4_fluxes_only_w_T0$f0_selected[na.action(na.omit(ch4_fluxes_only_w_T0$f0_selected))] <- ch4_fluxes_only_w_T0$linear.f0[na.action(na.omit(ch4_fluxes_only_w_T0$f0_selected))]

co2_fluxes_only_w_T0$f0_selected[na.action(na.omit(co2_fluxes_only_w_T0$f0_selected))] <- co2_fluxes_only_w_T0$linear.f0[na.action(na.omit(co2_fluxes_only_w_T0$f0_selected))]

n2o_fluxes_only_w_T0$f0_selected[na.action(na.omit(n2o_fluxes_only_w_T0$f0_selected))] <- n2o_fluxes_only_w_T0$linear.f0[na.action(na.omit(n2o_fluxes_only_w_T0$f0_selected))]

# Combine separate DFs back into one DF for statistics
# start with CH4 IDs and f0 for a base dataframe
select_fluxes_combined_w_T0 <- ch4_fluxes_only_w_T0[,c(1,5)]

# rename to distinguish between GHGs
colnames(select_fluxes_combined_w_T0)[2] <- "f0_CH4"

# add on CO2 fluxes
select_fluxes_combined_w_T0$f0_CO2 <- co2_fluxes_only_w_T0$f0_selected

# add on N2O fluxes
select_fluxes_combined_w_T0$f0_N2O <- n2o_fluxes_only_w_T0$f0_selected

# Character vector with correct sample ID order:
Order_GHG_flux <-   c("6/13/2016 1N", "6/13/2016 1P", "6/13/2016 2N", "6/13/2016 2P", "6/13/2016 3N", "6/13/2016 3P", "6/13/2016 4N", "6/13/2016 4P", "6/13/2016 5N", "6/13/2016 5P", "6/13/2016 6N", "6/13/2016 6P", "6/13/2016 7N", "6/13/2016 7P", "6/13/2016 8N", "6/13/2016 8P", "6/13/2016 9N", "6/13/2016 9P", "6/13/2016 10N", "6/13/2016 10P", "6/13/2016 11N", "6/13/2016 11P", "6/13/2016 12N", "6/13/2016 12P", "6/13/2016 13N", "6/13/2016 13P", "6/13/2016 14N", "6/13/2016 14P", "6/13/2016 15N", "6/13/2016 15P", "6/13/2016 16N", "6/13/2016 16P", "6/13/2016 17N", "6/13/2016 17P", "6/13/2016 18N", "6/13/2016 18P", "6/28/2016 1N", "6/28/2016 1P", "6/28/2016 2N", "6/28/2016 2P", "6/28/2016 3N", "6/28/2016 3P", "6/28/2016 4N", "6/28/2016 4P", "6/28/2016 5N", "6/28/2016 5P", "6/28/2016 6N", "6/28/2016 6P", "6/28/2016 7N", "6/28/2016 7P", "6/28/2016 8N", "6/28/2016 8P", "6/28/2016 9N", "6/28/2016 9P", "6/28/2016 10N", "6/28/2016 10P", "6/28/2016 11N", "6/28/2016 11P", "6/28/2016 12N", "6/28/2016 12P", "6/28/2016 13N", "6/28/2016 13P", "6/28/2016 14N", "6/28/2016 14P", "6/28/2016 15N", "6/28/2016 15P", "6/28/2016 16N", "6/28/2016 16P", "6/28/2016 17N", "6/28/2016 17P", "6/28/2016 18N", "6/28/2016 18P", "7/11/2016 1N", "7/11/2016 1P", "7/11/2016 2N", "7/11/2016 2P", "7/11/2016 3N", "7/11/2016 3P", "7/11/2016 4N", "7/11/2016 4P", "7/11/2016 5N", "7/11/2016 5P", "7/11/2016 6N", "7/11/2016 6P", "7/11/2016 7N", "7/11/2016 7P", "7/11/2016 8N", "7/11/2016 8P", "7/11/2016 9N", "7/11/2016 9P", "7/11/2016 10N", "7/11/2016 10P", "7/11/2016 11N", "7/11/2016 11P", "7/11/2016 12N", "7/11/2016 12P", "7/11/2016 13N", "7/11/2016 13P", "7/11/2016 14N", "7/11/2016 14P", "7/11/2016 15N", "7/11/2016 15P", "7/11/2016 16N", "7/11/2016 16P", "7/11/2016 17N", "7/11/2016 17P", "7/11/2016 18N", "7/11/2016 18P", "7/25/2016 1N", "7/25/2016 1P", "7/25/2016 2N", "7/25/2016 2P", "7/25/2016 3N", "7/25/2016 3P", "7/25/2016 4N", "7/25/2016 4P", "7/25/2016 5N", "7/25/2016 5P", "7/25/2016 6N", "7/25/2016 6P", "7/25/2016 7N", "7/25/2016 7P", "7/25/2016 8N", "7/25/2016 8P", "7/25/2016 9N", "7/25/2016 9P", "7/25/2016 10N", "7/25/2016 10P", "7/25/2016 11N", "7/25/2016 11P", "7/25/2016 12N", "7/25/2016 12P", "7/25/2016 13N", "7/25/2016 13P", "7/25/2016 14N", "7/25/2016 14P", "7/25/2016 15N", "7/25/2016 15P", "7/25/2016 16N", "7/25/2016 16P", "7/25/2016 17N", "7/25/2016 17P", "7/25/2016 18N", "7/25/2016 18P", "8/11/2016 1N", "8/11/2016 1P", "8/11/2016 2N", "8/11/2016 2P", "8/11/2016 3N", "8/11/2016 3P", "8/11/2016 4N", "8/11/2016 4P", "8/11/2016 5N", "8/11/2016 5P", "8/11/2016 6N", "8/11/2016 6P", "8/11/2016 7N", "8/11/2016 7P", "8/11/2016 8N", "8/11/2016 8P", "8/11/2016 9N", "8/11/2016 9P", "8/11/2016 10N", "8/11/2016 10P", "8/11/2016 11N", "8/11/2016 11P", "8/11/2016 12N", "8/11/2016 12P", "8/11/2016 13N", "8/11/2016 13P", "8/11/2016 14N", "8/11/2016 14P", "8/11/2016 15N", "8/11/2016 15P", "8/11/2016 16N", "8/11/2016 16P", "8/11/2016 17N", "8/11/2016 17P", "8/11/2016 18N", "8/11/2016 18P")

# Change order of select_fluxes_combined_w_T0 to match the other dataframes
select_fluxes_combined_w_T0 <- select_fluxes_combined_w_T0[match(Order_GHG_flux, select_fluxes_combined_w_T0$ID), ]

# CHECK ROW ORDERS NOW

# DF that contains meta info and same row number as select_fluxes_combined_w_T0
ghg_ppm_filtered <- ghg_ppm %>%
  dplyr::filter(timepoint == "T1")

# Add date_short back onto DFs for plotting
select_fluxes_combined_w_T0$date_short <- ghg_ppm_filtered$date_short

# Add on 'plant'
select_fluxes_combined_w_T0$plant <- ghg_ppm_filtered$plant

# Add on 'treatment'
select_fluxes_combined_w_T0$treatment <- ghg_ppm_filtered$treatment

# Add on 'history'
select_fluxes_combined_w_T0$history <- ghg_ppm_filtered$history
```

```{r ANOVAS prep}
# https://www.datanovia.com/en/lessons/repeated-measures-anova-in-r/

# Make sure factors are factors: 

  # ghg_ppm_filtered:
ghg_ppm_filtered$chamber <- as.factor(ghg_ppm_filtered$chamber)

select_fluxes_combined_w_T0$chamber <- ghg_ppm_filtered$chamber
select_fluxes_combined_w_T0$date_short <- as.factor(select_fluxes_combined_w_T0$date_short)
select_fluxes_combined_w_T0$plant <- as.factor(select_fluxes_combined_w_T0$plant)
select_fluxes_combined_w_T0$treatment <- as.factor(select_fluxes_combined_w_T0$treatment)
```

```{r ANOVA assumption testing: outliers}
# CH4
Outliers_f0_CH4_w_T0 <- select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  identify_outliers(f0_CH4)
Outliers_f0_CH4_w_T0
  # There are extreme outliers.

# CO2
Outliers_f0_CO2_w_T0 <- select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  identify_outliers(f0_CO2)
Outliers_f0_CO2_w_T0
  # There are extreme outliers.

# N2O
Outliers_f0_N2O_w_T0 <- select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  identify_outliers(f0_N2O)
Outliers_f0_N2O_w_T0
  # There are extreme outliers.
```
  
```{r ANOVA assumption testing: normality}
# CH4
select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  shapiro_test(f0_CH4)
  # Not normally distributed

  # QQ plot:
ggqqplot(select_fluxes_combined_w_T0, "f0_CH4", facet.by = "date_short")


# CO2
select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  shapiro_test(f0_CO2)
  # Not normally distributed

  # QQ plot:
ggqqplot(select_fluxes_combined_w_T0, "f0_CO2", facet.by = "date_short")

# N2O
select_fluxes_combined_w_T0 %>%
  group_by(date_short) %>%
  shapiro_test(f0_N2O)
  # Not normally distributed

  # QQ plot:
ggqqplot(select_fluxes_combined_w_T0, "f0_N2O", facet.by = "date_short")
```

```{r Kruskal-Wallis on flux with T0, treatment}
# Will use DWASS-Steel-Critchlow-Fligner test for multiple comparisons, in package NSM3

# CH4
kruskal_CH4_6.13_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CH4 ~ treatment, subset = date_short == "6/13")
kruskal_CH4_6.13_w_T0

kruskal_CH4_6.28_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CH4 ~ treatment, subset = date_short == "6/28")
kruskal_CH4_6.28_w_T0

  # Pairwise comparisons:
    # Must subset 6/28:
select_fluxes_combined_w_T0_6.28 <- select_fluxes_combined_w_T0 %>% filter(date_short == "7/25")

SDCFlig_CH4_6.28_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0_6.28$f0_CH4, g = select_fluxes_combined_w_T0_6.28$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_6.28_w_T0

kruskal_CH4_7.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CH4 ~ treatment, subset = date_short == "7/11")
kruskal_CH4_7.11_w_T0

  # Pairwise comparisons:
    # Must subset 7/11:
select_fluxes_combined_w_T0_7.11 <- select_fluxes_combined_w_T0 %>% filter(date_short == "7/11")

SDCFlig_CH4_7.11_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0_7.11$f0_CH4, g = select_fluxes_combined_w_T0_7.11$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_7.11_w_T0

kruskal_CH4_7.25_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CH4 ~ treatment, subset = date_short == "7/25")
kruskal_CH4_7.25_w_T0

  # Pairwise comparisons:
    # Must subset 7/25:
select_fluxes_combined_w_T0_7.25 <- select_fluxes_combined_w_T0 %>% filter(date_short == "7/25")

SDCFlig_CH4_7.25_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0_7.25$f0_CH4, g = select_fluxes_combined_w_T0_7.25$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CH4_7.25_w_T0

kruskal_CH4_8.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CH4 ~ treatment, subset = date_short == "8/11")
kruskal_CH4_8.11_w_T0


# CO2
kruskal_CO2_6.13_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CO2 ~ treatment, subset = date_short == "6/13")
kruskal_CO2_6.13_w_T0

kruskal_CO2_6.28_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CO2 ~ treatment, subset = date_short == "6/28")
kruskal_CO2_6.28_w_T0

  # Pairwise comparisons:
SDCFlig_CO2_6.28_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0$f0_CO2, g = select_fluxes_combined_w_T0$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CO2_6.28_w_T0

kruskal_CO2_7.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CO2 ~ treatment, subset = date_short == "7/11")
kruskal_CO2_7.11_w_T0

kruskal_CO2_7.25_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CO2 ~ treatment, subset = date_short == "7/25")
kruskal_CO2_7.25_w_T0

  # Pairwise comparisons:
SDCFlig_CO2_7.25_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0$f0_CO2, g = select_fluxes_combined_w_T0$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_CO2_7.25_w_T0

kruskal_CO2_8.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_CO2 ~ treatment, subset = date_short == "8/11")
kruskal_CO2_8.11_w_T0

# N2O
kruskal_N2O_6.13_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_N2O ~ treatment, subset = date_short == "6/13")
kruskal_N2O_6.13_w_T0

  # Pairwise comparisons:
    # Must subset 6/13:
select_fluxes_combined_w_T0_6.13 <- select_fluxes_combined_w_T0 %>% filter(date_short == "6/13")

SDCFlig_N2O_6.13_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0_6.13$f0_N2O, g = select_fluxes_combined_w_T0_6.13$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_N2O_6.13_w_T0

kruskal_N2O_6.28_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_N2O ~ treatment, subset = date_short == "6/28")
kruskal_N2O_6.28_w_T0

SDCFlig_N2O_6.28_w_T0 <- pSDCFlig(x = select_fluxes_combined_w_T0_6.28$f0_N2O, g = select_fluxes_combined_w_T0_6.28$treatment, method = "Monte Carlo", n.mc = 10000)
SDCFlig_N2O_6.28_w_T0

kruskal_N2O_7.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_N2O ~ treatment, subset = date_short == "7/11")
kruskal_N2O_7.11_w_T0

kruskal_N2O_7.25_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_N2O ~ treatment, subset = date_short == "7/25")
kruskal_N2O_7.25_w_T0

kruskal_N2O_8.11_w_T0 <- kruskal.test(data = select_fluxes_combined_w_T0, f0_N2O ~ treatment, subset = date_short == "8/11")
kruskal_N2O_8.11_w_T0
```


```{r format dates for figures}
select_fluxes_combined_w_T0$date_format <- factor(select_fluxes_combined_w_T0$date_short, levels = levels(select_fluxes_combined_w_T0$date_short), labels = c("13-Jun", "28-Jun", "11-Jul", "25-Jul","11-Aug"))
```

## Flux with T0: Plant vs. No Plant
```{r CH4 flux w T0 NP vs. P}
# Wilcox Test, nicer format, can use to verify stats on plot:
CH4_wilcox<- compare_means(f0_CH4~plant, select_fluxes_combined_w_T0, group.by="date_short", method = "wilcox.test", paired = T)

# Turn into data frame:
CH4_wilcox.DF <- as.data.frame(CH4_wilcox)

# Export the data frame to a .csv file
#write.csv(CH4_wilcox, file = "../figures/pub/tables/Wilcox_NPvP/CH4_wilcox.csv", row.names = T)

# Plotting:
plot.CH4.flux.NPvP <- ggplot(select_fluxes_combined_w_T0, aes(x=plant, y=f0_CH4)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=plant), outlier.shape = NA) +
                     geom_point(aes(color=plant, shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.y.npc = "center", label.x.npc = "center",size = 10, hide.ns = T)
 
plot.CH4.flux.NPvP
```

```{r CO2 flux NP vs P}
# Wilcox Test, nicer format, can use to verify stats on plot:
CO2_wilcox<- compare_means(f0_CO2~plant, select_fluxes_combined_w_T0, group.by="date_short", method = "wilcox.test", paired = T)

# Turn into data frame:
CO2_wilcox.DF <- as.data.frame(CO2_wilcox)

# Export the data frame to a .csv file
#write.csv(CO2_wilcox, file = "../figures/pub/tables/Wilcox_NPvP/CO2_wilcox.csv", row.names = T)

# Plotting:
plot.CO2.flux.NPvP <- ggplot(select_fluxes_combined_w_T0, aes(x=plant, y=f0_CO2)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=plant), outlier.shape = NA) +
                     geom_point(aes(color=plant, shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.y.npc = "center", label.x.npc = "center",size = 10, hide.ns = T)
 
plot.CO2.flux.NPvP
```

```{r N2O flux NP vs. P}
# Wilcox Test, nicer format, can use to verify stats on plot:
N2O_wilcox<- compare_means(f0_N2O~plant, select_fluxes_combined_w_T0, group.by="date_short", method = "wilcox.test", paired = T)

# Turn into data frame:
N2O_wilcox.DF <- as.data.frame(N2O_wilcox)

# Export the data frame to a .csv file
#write.csv(N2O_wilcox, file = "../figures/pub/tables/Wilcox_NPvP/N2O_wilcox.csv", row.names = T)

# Plotting:
plot.N2O.flux.NPvP <- ggplot(select_fluxes_combined_w_T0, aes(x=plant, y=f0_N2O)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=plant), outlier.shape = NA) +
                     geom_point(aes(color=plant, shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="bottom", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("No Plant", "Plant"))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_compare_means(method = "wilcox.test", paired = T, label = "p.signif", label.y.npc = 0.75, label.x.npc = "center",size = 10, hide.ns = T)
 
plot.N2O.flux.NPvP
```

```{r Extract and save GHG boxplot legend}
leg_NPvP <- get_legend(plot.N2O.flux.NPvP, position = "bottom") #Extract legend as a legend Grob
legend_plot_NPvP <- as_ggplot(leg_NPvP) #Asign the grob as a ggplot
legend_plot_NPvP #View the plot

#ggsave("../figures/pub/GHG_NPvP/legend_NPvP.tiff", plot=legend_plot_NPvP, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```

```{r Combined GHG Boxplot}
CombPlot_flux_NPvP <- ggarrange(plot.CH4.flux.NPvP, plot.CO2.flux.NPvP, plot.N2O.flux.NPvP, labels = c("A", "B", "C"), ncol = 1, nrow = 3, legend = "none", align = "hv")

CombPlot_flux_NPvP

#ggsave("../figures/pub/GHG_NPvP/Flux_NPvP.tiff", plot=CombPlot_flux_NPvP, device="tiff", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")

#ggsave("../figures/pub/GHG_NPvP/Flux_NPvP.png", plot=CombPlot_flux_NPvP, device="png", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")
```

## Flux w T0: Treatment and Kruskal boxplots

```{r CH4 flux Treatment}
# Kruskal-Wallis test on Treatment, and multiple comparisons:
CH4_kruskal<- compare_means(f0_CH4~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "kruskal.test")
CH4_multipleComparisons<- compare_means(f0_CH4~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "wilcox.test", p.adjust.method = "fdr")

# Adjustments for plotting:
CH4_multipleComparisons <- add_significance(CH4_multipleComparisons, p.col = "p.adj", output.col = "p.adj.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Turn into data frame:
CH4_MultComp.DF <- as.data.frame(CH4_multipleComparisons)

# Add on SDCFlig p-values:
CH4_MultComp.DF$SDCF_p <- c(1, 1, 1, SDCFlig_CH4_6.28_w_T0$p.val, SDCFlig_CH4_7.11_w_T0$p.val, SDCFlig_CH4_7.25_w_T0$p.val, 1, 1, 1)

# Symbols for plotting:
CH4_MultComp.DF <- add_significance(CH4_MultComp.DF, p.col = "SDCF_p", output.col = "p.SDCF.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Export the data frames to a .csv file
#write.csv(CH4_MultComp.DF, file = "../figures/pub/tables/Kruskal_Treatment/multiple_comparisons/CH4_MultComp.csv", row.names = T)

# Plotting:
plot.CH4.flux.Trt <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_CH4)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment,  shape=treatment), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(19, 17, 15), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", "", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_pvalue_manual(data = CH4_MultComp.DF, comparisons = list(c("D", "I"), c("D", "W"), c("I", "W")), label = "p.SDCF.signif", size = 10, vjust = 1, hide.ns = T, y.position = (max(select_fluxes_combined_w_T0$f0_CH4)+20), step.group.by = "date_format", step.increase = 0.1)
  #scale_y_continuous(limits = c(NA, 50)) # Useful zoom to exclude extreme outlier
 
plot.CH4.flux.Trt
```

```{r CO2 T30 ppm Treatment}
# Kruskal-Wallis test on Treatment, and multiple comparisons:
CO2_kruskal<- compare_means(f0_CO2~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "kruskal.test")
CO2_multipleComparisons<- compare_means(f0_CO2~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "wilcox.test", p.adjust.method = "fdr")

# Adjustments for plotting:
CO2_multipleComparisons <- add_significance(CO2_multipleComparisons, p.col = "p.adj", output.col = "p.adj.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Turn into data frame:
CO2_MultComp.DF <- as.data.frame(CO2_multipleComparisons)

# Add on SDCFlig p-values:
CO2_MultComp.DF$SDCF_p <- c(1, 1, 1, SDCFlig_CO2_6.28_w_T0$p.val, 1,1,1, SDCFlig_CO2_7.25_w_T0$p.val, 1, 1, 1)

# Symbols for plotting:
CO2_MultComp.DF <- add_significance(CO2_MultComp.DF, p.col = "SDCF_p", output.col = "p.SDCF.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Export the data frames to a .csv file
#write.csv(CO2_MultComp.DF, file = "../figures/pub/tables/Kruskal_Treatment/multiple_comparisons/CO2_MultComp.csv", row.names = T)

# Plotting:
plot.CO2.flux.Trt <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_CO2)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment,  shape=treatment), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(19, 17, 15), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", "", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_pvalue_manual(data = CO2_MultComp.DF, comparisons = list(c("D", "I"), c("D", "W"), c("I", "W")), label = "p.SDCF.signif", size = 10, vjust = 1, hide.ns = "p", y.position = max(select_fluxes_combined_w_T0$f0_CO2), step.group.by = "date_format", step.increase = 0.1)
 
plot.CO2.flux.Trt
```

```{r N2O T30 ppm Treatment}
# Kruskal-Wallis test on Treatment, and multiple comparisons:
N2O_kruskal<- compare_means(f0_N2O~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "kruskal.test")
N2O_multipleComparisons<- compare_means(f0_N2O~treatment, select_fluxes_combined_w_T0, group.by="date_format", method = "wilcox.test", p.adjust.method = "fdr")

# Adjustments for plotting:
N2O_multipleComparisons <- add_significance(N2O_multipleComparisons, p.col = "p.adj", output.col = "p.adj.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Turn into data frame:
N2O_MultComp.DF <- as.data.frame(N2O_multipleComparisons)

# Add on SDCFlig p-values:
N2O_MultComp.DF$SDCF_p <- c(SDCFlig_N2O_6.13_w_T0$p.val, 1,1,1, 1,1,1, 1,1,1, 1, 1, 1)

# Symbols for plotting:
N2O_MultComp.DF <- add_significance(N2O_MultComp.DF, p.col = "SDCF_p", output.col = "p.SDCF.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Export the data frames to a .csv file
#write.csv(N2O_MultComp.DF, file = "../figures/pub/tables/Kruskal_Treatment/multiple_comparisons/N2O_MultComp.csv", row.names = T)

# Weird 'hide.ns' behavior patch:
N2O_MultComp.DF$p[4] <- 1.0

# Plotting:
plot.N2O.flux.Trt <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_N2O)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                    geom_point(aes(color=treatment,  shape=treatment), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(19, 17, 15), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="bottom", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("Dry", "Interim", "Wet"))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  stat_pvalue_manual(data = N2O_MultComp.DF, comparisons = list(c("D", "I"), c("D", "W"), c("I", "W")), label = "p.SDCF.signif", size = 10, vjust = 1, hide.ns = "p", y.position = max(select_fluxes_combined_w_T0$f0_N2O), step.group.by = "date_format", step.increase = 0.1)
 
plot.N2O.flux.Trt
```

```{r Extract and save GHG boxplot legend}
leg_treat <- get_legend(plot.N2O.T30ppm.Trt, position = "bottom") #Extract legend as a legend Grob
legend_plot_treat <- as_ggplot(leg_treat) #Asign the grob as a ggplot
legend_plot_treat #View the plot

#ggsave("../figures/pub/GHG_Treatment/legend_treatment.tiff", plot=legend_plot_treat, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```


```{r Combined GHG Boxplot}
CombPlot_flux_treat <- ggarrange(plot.CH4.flux.Trt, plot.CO2.flux.Trt, plot.N2O.flux.Trt, labels = c("A", "B", "C"), ncol = 1, nrow = 3, legend = "none", align = "hv")

CombPlot_flux_treat

#ggsave("../figures/pub/GHG_Treatment/Flux_treatment.tiff", plot=CombPlot_flux_treat, device="tiff", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")

#ggsave("../figures/pub/GHG_Treatment/Flux_treatment.png", plot=CombPlot_flux_treat, device="png", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")
```

```{r Kruskal-Wallis table}
# Rbind the 3 Kruskal-Wallis tables:
Kruskal_comb <- rbind(CH4_kruskal, CO2_kruskal, N2O_kruskal)

# Write .csv: 
#write.csv(Kruskal_comb, file = "../figures/pub/tables/Kruskal_Treatment/Kruskal_combined.csv", row.names = T)
```



## GHG flux plots for FeS manuscript (new)
```{r CH4}
plot.CH4.flux <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_CH4)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment,  shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(1,19), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="bottom", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CH"[4]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", "", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))
  #scale_y_continuous(limits = c(NA, 50)) # Useful zoom to exclude extreme outlier
 
plot.CH4.flux
```

```{r CO2}
plot.CO2.flux <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_CO2)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment,  shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(1,19), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg CO"[2]," m"^{-2}, " h"^{-1})), title="") +
  scale_x_discrete(labels=c("", "", ""))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))
 
plot.CO2.flux
```

```{r N2O}
plot.N2O.flux <- ggplot(select_fluxes_combined_w_T0, aes(x=treatment, y=f0_N2O)) +
  facet_wrap(~date_format, nrow = 1, ncol = 5)+
                     geom_boxplot(aes(color=treatment), outlier.shape = NA) +
                     geom_point(aes(color=treatment,  shape=plant), size = 2, position = position_jitterdodge()) +
  scale_color_manual(values=c("#997700","#6699CC","#004488"), labels = c("Dry", "Interim", "Wet"), name="Treatment")+
  scale_shape_manual(values=c(1,19), labels = c("No Plant", "Plant"), name="")+
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 0),panel.border = element_rect(colour = "black",size=1),strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("mg N"[2],"O m"^{-2}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))
 
plot.N2O.flux
```

```{r extract and save legend}
leg_Flux <- get_legend(plot.CH4.flux, position = "bottom") #Extract legend as a legend Grob
legend_plot_Flux <- as_ggplot(leg_Flux) #Asign the grob as a ggplot
legend_plot_Flux #View the plot

#ggsave("../figures/Fe_S_manuscript/GHG_flux/legend_flux.tiff", plot=legend_plot_Flux, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```



```{r Combined GHG Boxplot}
CombPlot_flux <- ggarrange(plot.CH4.flux, plot.CO2.flux, plot.N2O.flux, labels = c("A", "B", "C"), ncol = 1, nrow = 3, legend = "top", align = "hv")

CombPlot_flux

#ggsave("../figures/Fe_S_manuscript/GHG_flux/Flux.tiff", plot=CombPlot_flux, device="tiff", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")

#ggsave("../figures/Fe_S_manuscript/GHG_flux/Flux.png", plot=CombPlot_flux, device="png", path=NULL, scale=1, width=12, height=10, dpi=600, limitsize=TRUE, bg="white")
```

## GHG summary

```{r GHG summary}
#Quick Data Summaries:

# History:
ghg.summary.history <- ghg %>%
  group_by(history)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.history

# Treatment:
ghg.summary.treatment <- ghg %>%
  group_by(treatment)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.treatment

# Plant:
ghg.summary.plant <- ghg %>%
  group_by(plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.plant

# Treatment and plant (tp):
ghg.summary.tp <- ghg %>%
  group_by(treatment, plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.tp

# Include all factors: history, treatment, and plant (htp):
ghg.summary.htp <- ghg %>%
  group_by(history, treatment, plant)%>%
  dplyr::summarise(mean_ch4=mean(mg.CH4.N.m.2), sd_ch4=sd(mg.CH4.N.m.2), mean_co2=mean(mg.CO2.N.m.2), sd_co2=sd(mg.CO2.N.m.2) , mean_n2o=mean(mg.N2O.N.m.2), sd_n2o=sd(mg.N2O.N.m.2))

ghg.summary.htp
```

```{r GHG ppm summary ppm}
# Treatment and plant (tp):
ghg.ppm.summary.tp <- ghg_ppm_filtered %>%
  group_by(treatment, plant)%>%
  dplyr::summarise(mean_ch4=mean(ch4_ppm), sd_ch4=sd(ch4_ppm), mean_co2=mean(co2_ppm), sd_co2=sd(co2_ppm) , mean_n2o=mean(n2o_ppm), sd_n2o=sd(n2o_ppm))

ghg.ppm.summary.tp

# format decimal places, 2 decimal places:
ghg.ppm.summary.tp$mean_ch4 <- format(round(ghg.ppm.summary.tp$mean_ch4, 2), nsmall = 2)
ghg.ppm.summary.tp$mean_ch4 <- as.numeric(ghg.ppm.summary.tp$mean_ch4)
ghg.ppm.summary.tp$sd_ch4 <- format(round(ghg.ppm.summary.tp$sd_ch4, 2), nsmall = 2)
ghg.ppm.summary.tp$sd_ch4 <- as.numeric(ghg.ppm.summary.tp$sd_ch4)

ghg.ppm.summary.tp$mean_co2 <- format(round(ghg.ppm.summary.tp$mean_co2, 2), nsmall = 2)
ghg.ppm.summary.tp$mean_co2 <- as.numeric(ghg.ppm.summary.tp$mean_co2)
ghg.ppm.summary.tp$sd_co2 <- format(round(ghg.ppm.summary.tp$sd_co2, 2), nsmall = 2)
ghg.ppm.summary.tp$sd_co2 <- as.numeric(ghg.ppm.summary.tp$sd_co2)

ghg.ppm.summary.tp$mean_n2o <- format(round(ghg.ppm.summary.tp$mean_n2o, 2), nsmall = 2)
ghg.ppm.summary.tp$mean_n2o <- as.numeric(ghg.ppm.summary.tp$mean_n2o)
ghg.ppm.summary.tp$sd_n2o <- format(round(ghg.ppm.summary.tp$sd_n2o, 2), nsmall = 2)
ghg.ppm.summary.tp$sd_n2o <- as.numeric(ghg.ppm.summary.tp$sd_n2o)

# Add plus or minus symbol to divide mean and SD:
ghg.ppm.summary.tp$mean_sd_ch4 <- paste(ghg.ppm.summary.tp$mean_ch4, "±", ghg.ppm.summary.tp$sd_ch4, sep = " ")
ghg.ppm.summary.tp$mean_sd_co2 <- paste(ghg.ppm.summary.tp$mean_co2, "±", ghg.ppm.summary.tp$sd_co2, sep = " ")
ghg.ppm.summary.tp$mean_sd_n2o <- paste(ghg.ppm.summary.tp$mean_n2o, "±", ghg.ppm.summary.tp$sd_n2o, sep = " ")

# Write .csv:
#write.csv(ghg.ppm.summary.tp, file = "../figures/pub/tables/ghg_ppm_summary.csv", row.names = T)
```



# Distance-based partial least squares regression (dbplsr)

```{r only ending samples}
#Using end samples for GHG because microbial community analyses was on end samples
#make df of design, soil, and ghg
ghg.filter.soil <- ghg %>%
  dplyr::select('date', 'chamber', 'timepoint', 'mg.CH4.N.m.2', 'mg.CO2.N.m.2', 'mg.N2O.N.m.2')%>%
  filter(timepoint=='T1' & date == '8/11/2016') 

soil.ghg <- soil %>% 
dplyr::select('chamber', 'time', 'plant', 'treatment', 'moisture.percent', 'pH', 'nh4.mg.l', 'c.percent', 'n.percent', 'P.ppm', 'K.ppm', 'Mg.ppm', 'S.ppm', 'Fe.ppm', 'Mn.ppm', 'HM', 'redox.percent.paint.removed') %>%
  filter(time=="finish") %>%
  bind_cols(ghg.filter.soil)
  
rownames(soil.ghg) <- rownames(otu.2.rel)

# Drop the two 'Chamber' columns (because Chambers are now row names)
soil.ghg <- soil.ghg[,-c(1,19)]

# Calculate euclidean distances for soil parameters only, soil parameters and GHGs, and GHGs only:
soilonly.dist <- vegdist(x = scale(soil.ghg[, -c(1:3, 17:21)]), "euclid")
soilghg.dist <- vegdist(x = scale(soil.ghg[, -c(1:3,17:18)]), "euclid")
ghg.dist <- vegdist(x = scale(soil.ghg[, -c(1:18)]), "euclid")
```

```{r only ending samples ppm}
#Using end samples for GHG because microbial community analyses was on end samples
#make df of design, soil, and ghg
  # Use ghg_ppm_filtered_8.11
ghg_ppm_filtered_8.11_MERGE <- ghg_ppm_filtered_8.11[,c(5:7, 11:12)]

soil.PPM.END.ghg <- soil %>% 
dplyr::select('chamber', 'time', 'plant', 'treatment', 'moisture.percent', 'pH', 'nh4.mg.l', 'c.percent', 'n.percent', 'P.ppm', 'K.ppm', 'Mg.ppm', 'S.ppm', 'Fe.ppm', 'Mn.ppm', 'HM', 'redox.percent.paint.removed') %>%
  filter(time=="finish") %>%
  bind_cols(ghg_ppm_filtered_8.11_MERGE)
  
rownames(soil.PPM.END.ghg) <- rownames(otu.2.rel)

# Drop the two 'Chamber' columns (because Chambers are now row names)
soil.PPM.END.ghg <- soil.PPM.END.ghg[,-1]

# Calculate euclidean distances for soil parameters only, soil parameters and GHGs, and GHGs only:
PPM.soilonly.dist <- vegdist(x = scale(soil.PPM.END.ghg[, -c(1:3, 17:21)]), "euclid")
PPM.soilghg.dist <- vegdist(x = scale(soil.PPM.END.ghg[, -c(1:3, 20:21)]), "euclid")
PPM.ghg.dist <- vegdist(x = scale(soil.PPM.END.ghg[, c(17:19)]), "euclid")
```

```{r dbplsr analysis}
#distance based partial least squares regression
#Compares distance matrices. can use mixture of continuous and qualitative variables. 
#Method good when factors are may and highly colinear. A PLS model find the multidimensional direction in the Z space that explains the variance in the Y space. 

# GHGs compared to Bray-Curtis Distance matrix of rarified OTUs: 
dbplsr.ch4 <- dbplsr(mg.CH4.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.ch4)

dbplsr.co2 <- dbplsr(mg.CO2.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.co2)

dbplsr.n2o <- dbplsr(mg.N2O.N.m.2 ~ as.matrix(otu.2.rel.dist), data=soil.ghg , ncomp=30, method="GCV")
summary(dbplsr.n2o)

# GHGs compared to euclidean distance matrix of soil parameters (sp): 
dbplsr.sp.ch4 <- dbplsr(mg.CH4.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.ch4)

dbplsr.sp.co2 <- dbplsr(mg.CO2.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.co2)

dbplsr.sp.n2o <- dbplsr(mg.N2O.N.m.2 ~ as.matrix(soilonly.dist), data=soil.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.n2o)
```

```{r dbplsr analysis ppm}
#distance based partial least squares regression
#Compares distance matrices. can use mixture of continuous and qualitative variables. 
#Method good when factors are may and highly colinear. A PLS model find the multidimensional direction in the Z space that explains the variance in the Y space. 

# GHGs compared to Bray-Curtis Distance matrix of rarified OTUs: 
dbplsr.PPM.ch4 <- dbplsr(ch4_ppm ~ as.matrix(otu.2.rel.dist), data=soil.PPM.END.ghg , ncomp=30, method="GCV")
summary(dbplsr.PPM.ch4)

dbplsr.PPM.co2 <- dbplsr(co2_ppm ~ as.matrix(otu.2.rel.dist), data=soil.PPM.END.ghg , ncomp=30, method="GCV")
summary(dbplsr.PPM.co2)

dbplsr.PPM.n2o <- dbplsr(n2o_ppm ~ as.matrix(otu.2.rel.dist), data=soil.PPM.END.ghg , ncomp=30, method="GCV")
summary(dbplsr.PPM.n2o)

# GHGs compared to euclidean distance matrix of soil parameters (sp): 
dbplsr.sp.PPM.ch4 <- dbplsr(ch4_ppm ~ as.matrix(soilonly.dist), data=soil.PPM.END.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.PPM.ch4)

dbplsr.sp.PPM.co2 <- dbplsr(co2_ppm ~ as.matrix(soilonly.dist), data=soil.PPM.END.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.PPM.co2)

dbplsr.sp.PPM.n2o <- dbplsr(n2o_ppm ~ as.matrix(soilonly.dist), data=soil.PPM.END.ghg , ncomp=35, method="GCV")
summary(dbplsr.sp.PPM.n2o)
```

# Mantel test: soil and bacterial community

```{r mantel soil and bac comm }
# For reproducibility, confirm that random seed is set:
set.seed(42)

#soil only
bac.soil.mantel.1 <- vegan::mantel(as.matrix(soilonly.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.1

#soil and ghg
bac.soil.mantel.2 <- vegan::mantel(as.matrix(soilghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.2

#ghg only 
bac.soil.mantel.3 <- vegan::mantel(as.matrix(ghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
bac.soil.mantel.3
```

```{r mantel soil and bac comm ppm}
# For reproducibility, confirm that random seed is set:
set.seed(42)

#soil only
PPM.bac.soil.mantel.1 <- vegan::mantel(as.matrix(PPM.soilonly.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
PPM.bac.soil.mantel.1

#soil and ghg
PPM.bac.soil.mantel.2 <- vegan::mantel(as.matrix(PPM.soilghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
PPM.bac.soil.mantel.2

#ghg only 
PPM.bac.soil.mantel.3 <- vegan::mantel(as.matrix(PPM.ghg.dist), as.matrix(otu.2.rel.dist), method="pearson", permutations=10000)
PPM.bac.soil.mantel.3
```


# Greenhouse gas and redox regressions

```{r CH4 redox regression ppm}
require(ggpmisc)

PPM.ch4.lm <- lm(log(ch4_ppm) ~ redox.percent.paint.removed, data=soil.PPM.END.ghg)
anova(PPM.ch4.lm)
summary(PPM.ch4.lm)

PPM.ch4.redox.regress.plot <- ggplot(soil.PPM.END.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=log(ch4_ppm), color=plant, shape = plant), size=5)+ 
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_fill_manual(values=c("#E69F00","#009E73")) +
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  geom_smooth(aes(x=(redox.percent.paint.removed*100), y=log(ch4_ppm)), formula = y ~ x, method=lm, se=T, color="black")+
  stat_poly_eq(aes(x=(redox.percent.paint.removed*100), y=log(ch4_ppm), label = paste(after_stat(adj.rr.label), after_stat(p.value.label), sep = "*\", \"*")), data = soil.PPM.END.ghg, formula = y ~ x, method = "lm", label.y = "top",label.x = "right")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = expression(paste("log(CH"[4],") (PPM)"))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "right")

PPM.ch4.redox.regress.plot
```


```{r CO2 redox regression ppm}
PPM.co2.lm <- lm(log(co2_ppm) ~ redox.percent.paint.removed, data=soil.PPM.END.ghg)
anova(PPM.co2.lm)
summary(PPM.co2.lm)

PPM.co2.redox.regress.plot <- ggplot(soil.PPM.END.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=log(co2_ppm), color=plant, shape = plant), size=5)+ 
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_fill_manual(values=c("#E69F00","#009E73")) +
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  geom_smooth(aes(x=(redox.percent.paint.removed*100), y=log(co2_ppm)), formula = y ~ x, method=lm, se=T, color="black")+
  stat_poly_eq(aes(x=(redox.percent.paint.removed*100), y=log(co2_ppm), label = paste(after_stat(adj.rr.label), after_stat(p.value.label), sep = "*\", \"*")), data = soil.PPM.END.ghg, formula = y ~ x, method = "lm", label.y = "top",label.x = "right")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = expression(paste("log(CO"[2],") (PPM)"))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "right")

PPM.co2.redox.regress.plot
```



```{r N2O redox regression ppm}
PPM.n2o.lm <- lm(log(n2o_ppm) ~ redox.percent.paint.removed, data=soil.PPM.END.ghg)
anova(PPM.n2o.lm)
summary(PPM.n2o.lm)

PPM.n2o.redox.regress.plot <- ggplot(soil.PPM.END.ghg)  +
  geom_point(aes(x=(redox.percent.paint.removed *100), y=log(n2o_ppm), color=plant, shape = plant), size=5)+ 
  scale_color_manual(values=c("#E69F00","#009E73"), labels = c("No Plant", "Plant"), name="")+
  scale_fill_manual(values=c("#E69F00","#009E73")) +
  scale_shape_manual(values=c(19, 17), labels = c("No Plant", "Plant"), name="")+
  geom_smooth(aes(x=(redox.percent.paint.removed*100), y=log(n2o_ppm)), formula = y ~ x, method=lm, se=T, color="black")+
  stat_poly_eq(aes(x=(redox.percent.paint.removed*100), y=log(n2o_ppm), label = paste(after_stat(adj.rr.label), after_stat(p.value.label), sep = "*\", \"*")), data = soil.PPM.END.ghg, formula = y ~ x, method = "lm", label.y = "top",label.x = "right")+
  theme_bw()  +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          axis.title.y=element_text(margin=margin(r=10)), 
          axis.title.x=element_text(margin=margin(t=10))) +
    theme(axis.title=element_text(vjust=1,size=18),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14), panel.border =
          element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "IRIS Percent", y = expression(paste("log(N"[2],"O) (PPM)"))) + 
    theme(legend.text=element_text(size=14), legend.title = element_text(size=14), 
          legend.position = "bottom")

PPM.n2o.redox.regress.plot
```

```{r extract legend and save}
leg_NPvP_reg <- get_legend(PPM.n2o.redox.regress.plot, position = "bottom") #Extract legend as a legend Grob
legend_plot_NPvP_reg <- as_ggplot(leg_NPvP_reg) #Asign the grob as a ggplot
legend_plot_NPvP_reg #View the plot

#ggsave("../figures/pub/Redox_Reg_legend.tiff", plot=legend_plot_NPvP_reg, device="tiff", path=NULL, scale=1, dpi=300, limitsize=TRUE, bg="white")
```


```{r Panel together GHG ~ redox regressions}
PPM.GHG_redox_3panel <- ggarrange(PPM.ch4.redox.regress.plot, PPM.co2.redox.regress.plot, PPM.n2o.redox.regress.plot, labels = c("A", "B", "C"), ncol = 1, nrow = 3, legend = "none", common.legend = T)

PPM.GHG_redox_3panel

#ggsave("../figures/pub/PPM_redox_regres.tiff", plot=PPM.GHG_redox_3panel, device="tiff", path=NULL, scale=1, width=10, height=10, dpi=900, limitsize=TRUE, bg="white")
```

